### 2010-03-01 tex/ms.tex
groups-simulations: abstract

\documentclass[12pt,preprint]{aastex}
\usepackage{amssymb,amsmath}
%\usepackage{color,hyperref}
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\figurenames}{\figurename s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\tvector}[1]{\boldsymbol{\vec{#1}}}
\newcommand{\vx}{\tvector{x}}
\newcommand{\vv}{\tvector{v}}

\begin{document}

\title{Probing Galactic structure with moving groups and radial velocity surveys}
\author{Jo~Bovy}%\altaffilmark{1,2}}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\email{jb2777@nyu.edu}
%\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
%  New York University, 4 Washington Place, New York, NY 10003, USA}
%\altaffiltext{2}{Correspondence should be addressed to jo.bovy@nyu.edu~.}

\begin{abstract}
Various dynamical scenarios related to the Milky Way's bar or spiral
structure have been proposed to explain the existence of moving
groups---clumps of co-moving stars---in the Solar neighborhood. These
models are hard to distinguish with observations of the local Galactic
neighborhood. We show that future programs that survey the Milky Way
disk's kinematics on a global scale can distinguish between these
models. By integrating orbits of stars in various steady-state and
transient bar and spiral structure potentials that produce the pattern
of moving groups observed in the Solar neighborhood, we predict the
velocity distribution as a function of location in the Galactic
disk. We identify the locations that distinguish most clearly between
the various scenarios. Future infrared radial velocity surveys can
probe the velocity distribution at these locations; We predict the
distribution of radial velocities that can be directly compared to
observations to establish the origin of the moving groups and to
constrain the structural and dynamical parameters of the
non-axisymetric Milky Way potential.
\end{abstract}

\keywords{
  Galaxy: bulge
  ---  
  Galaxy: disk
  ---
  Galaxy: evolution
  ---
  Galaxy: fundamental parameters
  ---
  Galaxy: kinematics and dynamics
  ---
  Galaxy: structure
}

\section{Introduction}

\begin{thebibliography}{}

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
{Binney},~J. \& {Tremaine},~S., 2008, {Galactic Dynamics: Second Edition}
  (Princeton University Press)

\bibitem[Bovy, Hogg, \& Roweis(2009)]{Bovy09a} Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T., 2009,
  \apj, 700, 1794

\end{thebibliography}

\end{document}

### 2010-03-01 tex/ms.tex
groups-simulations: outline of methodology section
groups-simulations: abstract tweaks

@@
 structure have been proposed to explain the existence of moving
-groups---clumps of co-moving stars---in the Solar neighborhood. These
+groups---clumps of co-moving stars---in the Solar neighborhood. Since
+the observed pattern of moving groups is very sensitive to the
+properties of the non-axisymmetric feature that creates them, these
+moving groups can be a powerful tool in constraining the dynamical
+structure of the inner Milky Way. However, at present the various
 models are hard to distinguish with observations of the local Galactic
 neighborhood. We show that future programs that survey the Milky Way
-disk's kinematics on a global scale can distinguish between these
+disk's kinematics on a more global scale can distinguish between these
 models. By integrating orbits of stars in various steady-state and
@@
 the various scenarios. Future infrared radial velocity surveys can
-probe the velocity distribution at these locations; We predict the
+probe the velocity distribution at these locations. We predict the
 distribution of radial velocities that can be directly compared to
@@
 constrain the structural and dynamical parameters of the
-non-axisymetric Milky Way potential.
+non-axisymmetric Milky Way potential.
 \end{abstract}
@@
 
+\section{Methodology}
+
+We follow Dehnen's approach \citep{dehnen00a}.
+
+\subsection{Bar potential}
+
+\subsection{Steady-state spiral structure potential}
+
+\subsection{Transient spiral structure potential}
+
 \begin{thebibliography}{}
@@
 
+\bibitem[Dehnen(2000)]{dehnen00a}
+  Dehnen,~W., 2000, \aj, 119, 800
+
 \end{thebibliography}

### 2010-03-02 py/integrate_orbits.py
groups-simulations: integrate orbits in the bar potential

@@ -1,0 +1,121 @@
+###############################################################################
+#   integrate_orbits.py: module that integrates orbits in a given 
+#                        non-axisymmetric, time-dependent potential
+###############################################################################
+import scipy as sc
+import scipy.integrate as integrate
+_degtorad= sc.pi/180.
+def uvToELz(UV=(0.,0.),Rtheta=(1.,0.),t=-4.,pot='bar',
+            potparams=(0.9,0.01,25.*_degtorad,.8,0.,2.)):
+    """
+    NAME:
+       uvToELz
+    PURPOSE:
+       calculate initial (E.Lz) for final (u,v)
+    INPUT:
+       (u,v) - final radial and tangential velocity, divided by vcirc
+       (R,theta) - final R and theta
+       t - time to integrate backwards for 
+           (interpretation depends on potential)
+       pot - type of non-axisymmetric, time-depedent potential
+       potparams - parameters for this potential
+    OUTPUT:
+       final (E,Lz)
+    HISTORY:
+       2010-03-01 - Written - Bovy (NYU)
+    """
+    #For now assume pot == 'bar'
+    (Rolr,alpha,phi,chi,beta,t1) = potparams
+    OmegaoOmegab= Rolr**(1.-beta)/(1.+sc.sqrt((1.+beta)/2.))
+    #Final conditions
+    u,v= UV
+    v+= 1. #Add circular velocity
+    R,theta = Rtheta
+    x= R*sc.cos(theta)
+    y= R*sc.sin(theta)
+    vR= (u*x+v*y)/R * OmegaoOmegab
+    vT= (v*x-u*y)/R * OmegaoOmegab
+    (vR,vT,R) = integrate_orbit((vR,vT,R),t=t,pot=pot,potparams=potparams)
+    return (axipotential(R,beta)+0.5*vR**2.+0.5*vT**2.,vT*R) #Assumes no perturbation at this time
+
+def axipotential(R,beta=0.):
+    """
+    NAME:
+       axipotential
+    PURPOSE:
+       return the axisymmetric potential at R/Ro
+    INPUT:
+       R - Galactocentric radius
+       beta - rotation curve power-law
+    OUTPUT:
+       Pot(R)/vo**2.
+    HISTORY:
+       2010-03-01 - Written - Bovy (NYU)
+    """
[…]

### 2010-03-02 py/integrate_orbits.py
groups-simulations: calculate energy on grid + time it (it's fast!!)

@@ -5,6 +5,38 @@
 import scipy as sc
 import scipy.integrate as integrate
 _degtorad= sc.pi/180.
+def uvToELz_grid(ulinspace,vlinspace,Rtheta=(1.,0.),t=-4.,pot='bar',
+            potparams=(0.9,0.01,25.*_degtorad,.8,0.,2.)):
+    """
+    NAME:
+       uvToELz_grid
+    PURPOSE:
+       calculate uvToLz on a grid in (u,v)
+    INPUT:
+       ulinspace, vlinspace - build the grid using scipy's linspace with
+                              these arguments
+       (R,theta) - final R and theta
+       t - time to integrate backwards for 
+           (interpretation depends on potential)
+       pot - type of non-axisymmetric, time-depedent potential
+       potparams - parameters for this potential
+    OUTPUT:
+       final (E,Lz) on grid [nus,nvs,2]
+    HISTORY:
+       2010-03-01 - Written - Bovy (NYU)
+    """
+    us= sc.linspace(*ulinspace)
+    vs= sc.linspace(*vlinspace)
+    nus= len(us)
+    nvs= len(vs)
+    out= sc.zeros((nus,nvs,2))
+    for ii in range(nus):
+        for jj in range(nvs):
+            tmp_out= uvToELz(UV=(us[ii],vs[jj]),Rtheta=Rtheta,t=t,pot=pot,potparams=potparams)
+            out[ii,jj,0]= tmp_out[0]
+            out[ii,jj,1]= tmp_out[1]
+    return out
+
 def uvToELz(UV=(0.,0.),Rtheta=(1.,0.),t=-4.,pot='bar',
             potparams=(0.9,0.01,25.*_degtorad,.8,0.,2.)):
     """
@@ -118,4 +150,15 @@ def barEOM(y,t,*args):
             
 
 if __name__ == '__main__':
+    #Various tests
     print uvToELz()
+
+    import timeit
+    
+    s="""r=uvToELz((numpy.random.random()-.5,numpy.random.random()-.5))"""
+    t = timeit.Timer(stmt=s,setup="from __main__ import uvToELz\nimport numpy")
+    print "uvToELz: %.4f sec/pass" % (t.timeit(number=500)/500)
+
+    if False:
+        testN= 101
+        print uvToELz_grid((-.5,.5,testN),(-.5,.5,testN))[testN/2,testN/2,:]

### 2010-03-08 py/integrate_orbits.py
groups-simulations: change to vR,vT initial conditions, get rid of theta

@@ -5,7 +5,7 @@
 import scipy as sc
 import scipy.integrate as integrate
 _degtorad= sc.pi/180.
-def uvToELz_grid(ulinspace,vlinspace,Rtheta=(1.,0.),t=-4.,pot='bar',
+def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',
             potparams=(0.9,0.01,25.*_degtorad,.8,0.,2.)):
     """
     NAME:
@@ -15,7 +15,7 @@ def uvToELz_grid(ulinspace,vlinspace,Rtheta=(1.,0.),t=-4.,pot='bar',
     INPUT:
        ulinspace, vlinspace - build the grid using scipy's linspace with
                               these arguments
-       (R,theta) - final R and theta
+       R - Galactoccentric Radius
        t - time to integrate backwards for 
            (interpretation depends on potential)
        pot - type of non-axisymmetric, time-depedent potential
@@ -32,12 +32,12 @@ def uvToELz_grid(ulinspace,vlinspace,Rtheta=(1.,0.),t=-4.,pot='bar',
     out= sc.zeros((nus,nvs,2))
     for ii in range(nus):
         for jj in range(nvs):
-            tmp_out= uvToELz(UV=(us[ii],vs[jj]),Rtheta=Rtheta,t=t,pot=pot,potparams=potparams)
+            tmp_out= uvToELz(UV=(us[ii],vs[jj]),R=R,t=t,pot=pot,potparams=potparams)
             out[ii,jj,0]= tmp_out[0]
             out[ii,jj,1]= tmp_out[1]
     return out
 
-def uvToELz(UV=(0.,0.),Rtheta=(1.,0.),t=-4.,pot='bar',
+def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',
             potparams=(0.9,0.01,25.*_degtorad,.8,0.,2.)):
     """
     NAME:
@@ -46,7 +46,7 @@ def uvToELz(UV=(0.,0.),Rtheta=(1.,0.),t=-4.,pot='bar',
        calculate initial (E.Lz) for final (u,v)
     INPUT:
        (u,v) - final radial and tangential velocity, divided by vcirc
-       (R,theta) - final R and theta
+       R = Galactocentric radius
        t - time to integrate backwards for 
            (interpretation depends on potential)
[…]

### 2010-03-09 py/integrate_orbits.py
groups-simulations: interpretation in terms of a df

@@ -22,6 +22,7 @@ def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',
        potparams - parameters for this potential
     OUTPUT:
        final (E,Lz) on grid [nus,nvs,2]
+       E=E/vo^2; Lz= Lz/Ro/vo
     HISTORY:
        2010-03-01 - Written - Bovy (NYU)
     """
@@ -53,6 +54,7 @@ def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',
        potparams - parameters for this potential
     OUTPUT:
        final (E,Lz)
+       E=E/vo^2; Lz= Lz/Ro/vo
     HISTORY:
        2010-03-01 - Written - Bovy (NYU)
     """
@@ -65,6 +67,8 @@ def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',
     vR= u * OmegaoOmegab
     vT= v * OmegaoOmegab
     (vR,vT,R) = integrate_orbit((vR,vT,R),t=t,pot=pot,potparams=potparams)
+    vR/= OmegaoOmegab
+    vT/= OmegaoOmegab
     return (axipotential(R,beta)+0.5*vR**2.+0.5*vT**2.,vT*R) #Assumes no perturbation at this time
 
 def axipotential(R,beta=0.):

### 2010-03-09 py/interpret_as_df.py

@@ -1,0 +1,42 @@
+###############################################################################
+#   interpret_as_df.py: module that interprets (E,Lz) pairs in terms of a 
+#                        distribution function
+###############################################################################
+import scipy as sc
+def interpret_as_df(EL,beta=0.,xD=0.33,xS=1.,Sro=0.2,type='dehnen'):
+    """
+    NAME:
+       interpret_as_df
+    PURPOSE:
+       interpret (E,L) as a DF (probability)
+    INPUT:
+       EL - (E,L) - Energy and angular momentum (E/vo^2,L/Ro/vo)
+       beta - rotation curve power-law
+       xD - disk surface mass scalelength / Ro
+       xS - disk velocity dispersion scalelength / Ro
+       Sro - disk velocity dispersion at Ro (/vo)
+       type - type of DF ('dehnen' or 'shu')
+    OUTPUT:
+       DF(E,L)
+    HISTORY:
+       2010-03-08 - Written - Bovy (NYU)
+    """
+    if type == 'dehnen':
+        return _interpret_as_df_dehnen(EL,beta,xD,xS,Sro)
+
+def _interpret_as_df_dehnen(EL,beta,xD,xS,Sro):
+    E, L= EL
+    #Calculate Re,LE, OmegaE
+    if beta == 0.:
+        xE= sc.exp(2.*E-1.)
+        LE= xE
+        OmegaE= 1./xE
+    else: #non-flat rotation curve
+        xE= (2.*E/(1.+1./2./beta))**(1./2./beta)
+        LE= xE**(beta+1.)
+        xE= xE*(beta-1.)
+    SRE2= Sro**2.*sc.exp(-2.*(xE-1.)/xS)
+    return sc.exp(-xE/xD)/SRE2*sc.exp(OmegaE*(L-LE)/SRE2)
+
+if __name__ == '__main__':
+    print interpret_as_df((0.54636764432720319, 0.99999999999999989),type='dehnen')

### 2010-03-09 py/integrate_orbits.py
groups-simulations: full pipeline for bar / not working

@@ -5,8 +5,8 @@
 import scipy as sc
 import scipy.integrate as integrate
 _degtorad= sc.pi/180.
-def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',
-            potparams=(0.9,0.01,25.*_degtorad,.8,0.,2.)):
+def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
+            potparams=(0.9,0.01,25.*_degtorad,.8,2.)):
     """
     NAME:
        uvToELz_grid
@@ -15,10 +15,11 @@ def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',
     INPUT:
        ulinspace, vlinspace - build the grid using scipy's linspace with
                               these arguments
-       R - Galactoccentric Radius
+       R - Galactocentric Radius
        t - time to integrate backwards for 
            (interpretation depends on potential)
-       pot - type of non-axisymmetric, time-depedent potential
+       pot - type of non-axisymmetric, time-dependent potential
+       beta - power-law index of rotation curve
        potparams - parameters for this potential
     OUTPUT:
        final (E,Lz) on grid [nus,nvs,2]
@@ -33,13 +34,13 @@ def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',
     out= sc.zeros((nus,nvs,2))
     for ii in range(nus):
         for jj in range(nvs):
-            tmp_out= uvToELz(UV=(us[ii],vs[jj]),R=R,t=t,pot=pot,potparams=potparams)
+            tmp_out= uvToELz(UV=(us[ii],vs[jj]),R=R,t=t,pot=pot,beta=beta,potparams=potparams)
             out[ii,jj,0]= tmp_out[0]
             out[ii,jj,1]= tmp_out[1]
     return out
 
-def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',
-            potparams=(0.9,0.01,25.*_degtorad,.8,0.,2.)):
+def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',beta=0.,
+            potparams=(0.9,0.01,25.*_degtorad,.8,2.)):
     """
     NAME:
        uvToELz
@@ -50,7 +51,8 @@ def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',
        R = Galactocentric radius
[…]

### 2010-03-09 py/calc_veldist_2d.py
groups-simulations: forgot to ci calc_veldist_2d

@@ -1,0 +1,44 @@
+import scipy as sc
+from integrate_orbits import uvToELz
+from interpret_as_df import interpret_as_df
+_degtorad= sc.pi/180.
+def calc_veldist_2d(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
+                    potparams=(0.9,0.01,25.*_degtorad,.8,2.),
+                    dfparams=(0.33,1.,0.2),dftype='dehnen'):
+    """
+    NAME:
+       calc_veldist_2d
+    PURPOSE:
+       calculate the velocity distribution at a given R on a grid in (u,v)
+    INPUT:
+       ulinspace, vlinspace - build the grid using scipy's linspace with
+                              these arguments
+       R - Galactocentric Radius
+       t - time to integrate backwards for 
+           (interpretation depends on potential)
+       pot - type of non-axisymmetric, time-dependent potential
+       beta - power-law index of rotation curve
+       potparams - parameters for this potential
+       dfparams - parameters of the DF (xD,xS,Sro)
+       dftype - type of DF ('dehnen' or 'shu')
+    OUTPUT:
+       f(u,v) on the grid
+    HISTORY:
+       2010-03-08 - Written - Bovy (NYU)
+    """
+    us= sc.linspace(*ulinspace)
+    vs= sc.linspace(*vlinspace)
+    nus= len(us)
+    nvs= len(vs)
+    out= sc.zeros((nus,nvs))
+    xD, xS, Sro= dfparams
+    for ii in range(nus):
+        for jj in range(nvs):
+            tmp_out= uvToELz(UV=(us[ii],vs[jj]),R=R,t=t,pot=pot,potparams=potparams,beta=beta)
+            out[ii,jj]= interpret_as_df(tmp_out,beta=beta,xD=xD,xS=xS,Sro=Sro,type=dftype)
+    return out
+
+if __name__ == '__main__':
+    testN= 51
+    #print calc_veldist_2d((-.5,.5,testN),(-.5,.5,testN))[testN/2,testN/2]
+    print calc_veldist_2d((-.5,.5,testN),(-.5,.5,testN))

### 2010-03-09 py/calc_veldist_2d.py
groups-simulations: I seem able to reproduce Dehnen's results!

@@ -3,7 +3,7 @@ from integrate_orbits import uvToELz
 from interpret_as_df import interpret_as_df
 _degtorad= sc.pi/180.
 def calc_veldist_2d(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
-                    potparams=(0.9,0.01,25.*_degtorad,.8,2.),
+                    potparams=(0.9,0.01,25.*_degtorad,.8,None),
                     dfparams=(0.33,1.,0.2),dftype='dehnen'):
     """
     NAME:
@@ -34,8 +34,8 @@ def calc_veldist_2d(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
     xD, xS, Sro= dfparams
     for ii in range(nus):
         for jj in range(nvs):
-            tmp_out= uvToELz(UV=(us[ii],vs[jj]),R=R,t=t,pot=pot,potparams=potparams,beta=beta)
-            out[ii,jj]= interpret_as_df(tmp_out,beta=beta,xD=xD,xS=xS,Sro=Sro,type=dftype)
+            EL= uvToELz(UV=(-us[ii],vs[jj]),R=R,t=t,pot=pot,potparams=potparams,beta=beta)
+            out[ii,jj]= interpret_as_df(EL,beta=beta,xD=xD,xS=xS,Sro=Sro,type=dftype)
     return out
 
 if __name__ == '__main__':

### 2010-03-09 py/integrate_orbits.py

@@ -6,7 +6,7 @@ import scipy as sc
 import scipy.integrate as integrate
 _degtorad= sc.pi/180.
 def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
-            potparams=(0.9,0.01,25.*_degtorad,.8,2.)):
+            potparams=(0.9,0.01,25.*_degtorad,.8,None)):
     """
     NAME:
        uvToELz_grid
@@ -40,7 +40,7 @@ def uvToELz_grid(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
     return out
 
 def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',beta=0.,
-            potparams=(0.9,0.01,25.*_degtorad,.8,2.)):
+            potparams=(0.9,0.01,25.*_degtorad,.8,None)):
     """
     NAME:
        uvToELz
@@ -48,12 +48,14 @@ def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',beta=0.,
        calculate initial (E.Lz) for final (u,v)
     INPUT:
        (u,v) - final radial and tangential velocity, divided by vcirc
+               ACTUALLY -U!!
        R = Galactocentric radius
        t - time to integrate backwards for 
            (interpretation depends on potential)
        pot - type of non-axisymmetric, time-dependent potential
        beta - power-law index of rotation curve
        potparams - parameters for this potential
+                   (Rolr,alpha,phi,chi,t1)
     OUTPUT:
        final (E,Lz)
        E=E/vo^2; Lz= Lz/Ro/vo
@@ -93,7 +95,7 @@ def axipotential(R,beta=0.):
         return R**(2.*beta)/2./beta
 
 def integrate_orbit(vRvTR= (0.,1.,1.),t=-4.,pot='bar',beta=0.,
-                    potparams=(0.9,0.01,25.*_degtorad,.8,2.)):
+                    potparams=(0.9,0.01,25.*_degtorad,.8,None)):
     """
     NAME:
        integrate_orbit
@@ -114,15 +116,19 @@ def integrate_orbit(vRvTR= (0.,1.,1.),t=-4.,pot='bar',beta=0.,
     """
[…]

### 2010-03-09 py/interpret_as_df.py

@@ -28,15 +28,35 @@ def _interpret_as_df_dehnen(EL,beta,xD,xS,Sro):
     E, L= EL
     #Calculate Re,LE, OmegaE
     if beta == 0.:
-        xE= sc.exp(2.*E-1.)
+        xE= sc.exp(E-.5)
         LE= xE
         OmegaE= 1./xE
     else: #non-flat rotation curve
-        xE= (2.*E/(1.+1./2./beta))**(1./2./beta)
+        xE= (2.*E/(1.+1./beta))**(1./2./beta)
         LE= xE**(beta+1.)
-        xE= xE*(beta-1.)
+        OmegaE= xE*(beta-1.)
     SRE2= Sro**2.*sc.exp(-2.*(xE-1.)/xS)
     return sc.exp(-xE/xD)/SRE2*sc.exp(OmegaE*(L-LE)/SRE2)
 
 if __name__ == '__main__':
     print interpret_as_df((0.54636764432720319, 0.99999999999999989),type='dehnen')
+
+    from integrate_orbits import uvToELz
+    step=0.1
+    EL= uvToELz((0.,0.))
+    print "step= ", step
+    print "(0,0):",EL, interpret_as_df(EL)
+
+    EL= uvToELz((step,0.))
+    print "(1,0):",EL, interpret_as_df(EL)
+
+    EL= uvToELz((0.,step))
+    print "(0,1):",EL, interpret_as_df(EL)
+
+    EL= uvToELz((-step,0.))
+    print "(-1,0):",EL, interpret_as_df(EL)
+
+    EL= uvToELz((0.,-step))
+    print "(0,-1):",EL, interpret_as_df(EL)
+
+

### 2010-03-10 py/integrate_orbits.py
groups-simulations: bar-potential inside bar is wrong!!

@@ -128,7 +128,8 @@ def integrate_orbit(vRvTR= (0.,1.,1.),t=-4.,pot='bar',beta=0.,
     args= (h,OmegaoOmegab,beta,alpha,phi,Rb,t1)
     time= [0.,2.*sc.pi*t]
     initCond= [R,vR]
-    intOut= integrate.odeint(barEOM,initCond,time,args=args,rtol=10.**-15.,mxstep=100000000)
+    intOut= integrate.odeint(barEOM,initCond,time,args=args,
+                             rtol=10.**-15.,mxstep=100000000)
     vRF= intOut[1,1]
     RF= intOut[1,0]
     vTF= h/RF
@@ -157,15 +158,15 @@ def barEOM(y,t,*args):
     else: #bar is fully on
         smooth= 1.
     if x <= Rb:
-        barsign= 1.
+        #BOVY: IMPLEMENT
+        return [y[1],10000]
     else: #outside of bar
-        barsign= -1.
-    return [y[1],h**2./x**3.-OmegaoOmegab**2.*x**(2.*beta-1.)+barsign*alpha*OmegaoOmegab**2.*sc.cos(2.*(phi-t))*smooth/x**4.]
+        return [y[1],h**2./x**3.-OmegaoOmegab**2.*x**(2.*beta-1.)-alpha*OmegaoOmegab**2.*sc.cos(2.*(phi-t))*smooth/x**4.]
             
 
 if __name__ == '__main__':
     #Various tests
-    print uvToELz()
+    print uvToELz((-.9,-.8))
 
     import timeit
     

### 2010-03-10 py/integrate_orbits.py
groups-simulations: inside-bar potential

@@ -159,7 +159,7 @@ def barEOM(y,t,*args):
         smooth= 1.
     if x <= Rb:
         #BOVY: IMPLEMENT
-        return [y[1],10000]
+        return [y[1],h**2./x**3.-OmegaoOmegab**2.*x**(2.*beta-1.)-alpha*OmegaoOmegab**2./Rb**6.*sc.cos(2.*(phi-t))*smooth*x**2.]
     else: #outside of bar
         return [y[1],h**2./x**3.-OmegaoOmegab**2.*x**(2.*beta-1.)-alpha*OmegaoOmegab**2.*sc.cos(2.*(phi-t))*smooth/x**4.]
             

### 2010-03-10 py/calc_veldist_2d.py
groups-simulations: experiments with integration

@@ -3,7 +3,7 @@ from integrate_orbits import uvToELz
 from interpret_as_df import interpret_as_df
 _degtorad= sc.pi/180.
 def calc_veldist_2d(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
-                    potparams=(0.9,0.01,25.*_degtorad,.8,None),
+                    potparams=(0.9,0.01,20.*_degtorad,.8,None),
                     dfparams=(0.33,1.,0.2),dftype='dehnen'):
     """
     NAME:

### 2010-03-10 py/integrate_orbits.py

@@ -64,10 +64,10 @@ def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',beta=0.,
     """
     #For now assume pot == 'bar'
     (Rolr,alpha,phi,chi,t1) = potparams
-    OmegaoOmegab= Rolr**(1.-beta)/(1.+sc.sqrt((1.+beta)/2.))
+    OmegaoOmegab= (Rolr**(1.-beta))/(1.+sc.sqrt((1.+beta)/2.))
     #Final conditions
     u,v= UV
-    v+= 1. #Add circular velocity
+    v+= R**beta #Add circular velocity
     vR= u * OmegaoOmegab
     vT= v * OmegaoOmegab
     (vR,vT,R) = integrate_orbit((vR,vT,R),t=t,pot=pot,beta=beta,potparams=potparams)
@@ -122,11 +122,12 @@ def integrate_orbit(vRvTR= (0.,1.,1.),t=-4.,pot='bar',beta=0.,
     #Integrate the orbit
     Rb= chi*OmegaoOmegab**(1./(1.-beta))
     if t1 == None:
-        t1= t/2.
+        t1= sc.fabs(t)/2.
     t1*= 2.*sc.pi
     phi+= t*2.*sc.pi
-    args= (h,OmegaoOmegab,beta,alpha,phi,Rb,t1)
-    time= [0.,2.*sc.pi*t]
+    tend= 2.*sc.pi*t
+    args= (h,OmegaoOmegab,beta,alpha,phi,Rb,t1,tend)
+    time= [0.,tend]
     initCond= [R,vR]
     intOut= integrate.odeint(barEOM,initCond,time,args=args,
                              rtol=10.**-15.,mxstep=100000000)
@@ -150,23 +151,31 @@ def barEOM(y,t,*args):
     HISTORY:
        2010-03-01 - Written - Bovy (NYU)
     """
-    h, OmegaoOmegab,beta,alpha,phi,Rb, t1= args
+    h, OmegaoOmegab,beta,alpha,phi,Rb, t1, tend= args
     x= y[0]
-    if sc.fabs(t) < t1:
-        xi= 2.*sc.fabs(t)/t1-1.
+    deltat= sc.fabs(t-tend)
+    if deltat < t1:
+        xi= 2.*deltat/t1-1.
         smooth= (3./16.*xi**5.-5./8*xi**3.+15./16.*xi+.5)
     else: #bar is fully on
         smooth= 1.
[…]

### 2010-03-10 py/interpret_as_df.py

@@ -42,6 +42,7 @@ if __name__ == '__main__':
     print interpret_as_df((0.54636764432720319, 0.99999999999999989),type='dehnen')
 
     from integrate_orbits import uvToELz
+    """
     step=0.1
     EL= uvToELz((0.,0.))
     print "step= ", step
@@ -58,5 +59,10 @@ if __name__ == '__main__':
 
     EL= uvToELz((0.,-step))
     print "(0,-1):",EL, interpret_as_df(EL)
+    """
+
+    _degtorad= sc.pi/180.
+    print interpret_as_df(uvToELz((0,0),potparams=(1.,0.01,25.*_degtorad,.8,None)))
+    print interpret_as_df(uvToELz((-.1,-.1),potparams=(1.,0.01,25.*_degtorad,.8,None)))
 
 

### 2010-03-11 py/interpret_as_df.py
groups-simulations: started working on a better DF

@@ -2,6 +2,8 @@
 #   interpret_as_df.py: module that interprets (E,Lz) pairs in terms of a 
 #                        distribution function
 ###############################################################################
+import os, os.path
+import cPickle as pickle
 import scipy as sc
 def interpret_as_df(EL,beta=0.,xD=0.33,xS=1.,Sro=0.2,type='dehnen'):
     """
@@ -37,6 +39,132 @@ def _interpret_as_df_dehnen(EL,beta,xD,xS,Sro):
         OmegaE= xE*(beta-1.)
     SRE2= Sro**2.*sc.exp(-2.*(xE-1.)/xS)
     return sc.exp(-xE/xD)/SRE2*sc.exp(OmegaE*(L-LE)/SRE2)
+
+class DFcorrection:
+    """Class that contains the corrections necessary to reach
+    exponential profiles"""
+    def __init__(self,**kwargs):
+        """
+        NAME:
+           __init__
+        PURPOSE:
+           initialize the corrections: set them, load them, or calculate
+           and save them
+        OPTIONAL INPUTS:
+           corrections - if Set, these are the corrections and they should
+                         be used as such
+           npoints - number of points from 0 to Rmax
+           rmax - correct up to this radius (/ro)
+           savedir - save the corrections in this directory
+           dfparams - parameters of the df: xD, xS, Sro
+           beta - power-law index of the rotation curve (when calculating)
+           dftype - 'dehnen'
+           niter - number of iterations to perform to calculate the corrections
+        OUTPUT:
+        HISTORY:
+           2010-03-10 - Written - Bovy (NYU)
+        """
+        if not kwargs.has_key('dfparams'):
+            self._dfparams= (0.33,1.,0.2)
+        else:
+            self._dfparams= kwargs['dfparams']
+        if not kwargs.has_key('rmax'):
+            self._rmax= 10.*self._dfparams[0]
+        else:
+            self._rmax= kwargs['rmax']
+        if not kwargs.has_key('niter'):
+            self._niter= 4
+        else:
+            self._niter= kwargs['niter']
+        if not kwargs.has_key('npoints'):
+            if kwargs.has_key('corrections'):
+                self._npoints= kwargs['corrections'].shape[0]
+            else:
+                self._npoints= 101
+        else:
+            self._npoints= kwargs['npoints']
+        if kwargs.has_key('dftype'):
+            self._dftype= kwargs['dftype']
+        else:
+            self._dftype= 'dehnen'
+        if kwargs.has_key('corrections'):
[…]

### 2010-03-11 py/calc_veldist_2d.py
groups: some progress on calculating the surface mass density

@@ -1,6 +1,6 @@
 import scipy as sc
 from integrate_orbits import uvToELz
-from interpret_as_df import interpret_as_df
+from interpret_as_df import interpret_as_df, distF
 _degtorad= sc.pi/180.
 def calc_veldist_2d(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
                     potparams=(0.9,0.01,20.*_degtorad,.8,None),
@@ -31,11 +31,11 @@ def calc_veldist_2d(ulinspace,vlinspace,R=1.,t=-4.,pot='bar',beta=0.,
     nus= len(us)
     nvs= len(vs)
     out= sc.zeros((nus,nvs))
-    xD, xS, Sro= dfparams
+    df= distF(dftype=dftype,dfparams=dfparams,beta=beta)
     for ii in range(nus):
         for jj in range(nvs):
-            EL= uvToELz(UV=(-us[ii],vs[jj]),R=R,t=t,pot=pot,potparams=potparams,beta=beta)
-            out[ii,jj]= interpret_as_df(EL,beta=beta,xD=xD,xS=xS,Sro=Sro,type=dftype)
+            E,L= uvToELz(UV=(-us[ii],vs[jj]),R=R,t=t,pot=pot,potparams=potparams,beta=beta)
+            out[ii,jj]= df.eval(E,L)
     return out
 
 if __name__ == '__main__':

### 2010-03-11 py/integrate_orbits.py

@@ -73,7 +73,23 @@ def uvToELz(UV=(0.,0.),R=1.,t=-4.,pot='bar',beta=0.,
     (vR,vT,R) = integrate_orbit((vR,vT,R),t=t,pot=pot,beta=beta,potparams=potparams)
     vR/= OmegaoOmegab
     vT/= OmegaoOmegab
-    return (axipotential(R,beta)+0.5*vR**2.+0.5*vT**2.,vT*R) #Assumes no perturbation at this time
+    return vRvTRToEL(vR,vT,R)
+
+def vRvTRToEL(vR,vT,R,beta):
+    """
+    NAME:
+       vRvTRToEL
+    PURPOSE:
+       calculate the energy and angular momentum
+    INPUT:
+       vR - radial velocity
+       vT - tangential velocity
+       R - Galactocentric radius
+    OUTPUT:
+    HISTORY:
+       2010-03-10 - Written - Bovy (NYU)
+    """
+    return (axipotential(R,beta)+0.5*vR**2.+0.5*vT**2.,vT*R)
 
 def axipotential(R,beta=0.):
     """

### 2010-03-11 py/interpret_as_df.py

@@ -5,40 +5,98 @@
 import os, os.path
 import cPickle as pickle
 import scipy as sc
-def interpret_as_df(EL,beta=0.,xD=0.33,xS=1.,Sro=0.2,type='dehnen'):
-    """
-    NAME:
-       interpret_as_df
-    PURPOSE:
-       interpret (E,L) as a DF (probability)
-    INPUT:
-       EL - (E,L) - Energy and angular momentum (E/vo^2,L/Ro/vo)
-       beta - rotation curve power-law
-       xD - disk surface mass scalelength / Ro
-       xS - disk velocity dispersion scalelength / Ro
-       Sro - disk velocity dispersion at Ro (/vo)
-       type - type of DF ('dehnen' or 'shu')
-    OUTPUT:
-       DF(E,L)
-    HISTORY:
-       2010-03-08 - Written - Bovy (NYU)
-    """
-    if type == 'dehnen':
-        return _interpret_as_df_dehnen(EL,beta,xD,xS,Sro)
+import scipy.integrate as integrate
+from integrate_orbits import vRvTRToEL
+class distF:
+    """Class that represents a DF"""
+    def __init__(self,dftype='dehnen',dfparams=(0.33,1.0,0.2),beta=0.,**kwargs):
+        """
+        NAME:
+           __init__
+        PURPOSE:
+           Initialize a DF
+        INPUT:
+           dftype= 'dehnen' or 'corrected-dehnen'
+           dfparams - parameters of the df (xD,xS,Sro)
+                      xD - disk surface mass scalelength / Ro
+                      xS - disk velocity dispersion scalelength / Ro
+                      Sro - disk velocity dispersion at Ro (/vo)
+           beta - power-law index of the rotation curve
+           + DFcorrection kwargs
+        OUTPUT:
+        HISTORY:
+            2010-03-10 - Written - Bovy (NYU)
+        """
+        self._dftype= dftype
+        self._dfparams= dfparams
+        self._beta= beta
+        if dftype == 'corrected-dehnen':
+            #Load corrections
+            corr= DFcorrection(**kwargs)
+        return None
 
-def _interpret_as_df_dehnen(EL,beta,xD,xS,Sro):
-    E, L= EL
-    #Calculate Re,LE, OmegaE
-    if beta == 0.:
-        xE= sc.exp(E-.5)
-        LE= xE
-        OmegaE= 1./xE
-    else: #non-flat rotation curve
-        xE= (2.*E/(1.+1./beta))**(1./2./beta)
[…]

### 2010-03-11 py/interpret_as_df.py
groups-simulations: do kind of a lognorm for the integral for the surface mass density

@@ -35,7 +35,7 @@ class distF:
             corr= DFcorrection(**kwargs)
         return None
 
-    def eval(self,E,L):
+    def eval(self,E,L,log=False):
         """
         NAME:
            eval
@@ -44,17 +44,18 @@ class distF:
         INPUT:
            E - energy (/vo^2)
            L - angular momentun (/ro/vo)
+           log - return the log (default: false)
         OUTPUT:
            DF(E,L)
         HISTORY:
            2010-03-10 - Written - Bovy (NYU)
         """
         if self._dftype == 'dehnen':
-            return self._eval_dehnen(E,L)
+            return self._eval_dehnen(E,L,log)
         elif self._dftype == 'corrected-dehnen':
-            return self._eval_corrected_dehnen(E,L)
+            return self._eval_corrected_dehnen(E,L,log)
     
-    def _eval_dehnen(self,E,L):
+    def _eval_dehnen(self,E,L,log):
         """Internal function that evaluates the uncorrected Dehnen DF"""
         #Calculate Re,LE, OmegaE
         if self._beta == 0.:
@@ -66,9 +67,12 @@ class distF:
             LE= xE**(self._beta+1.)
             OmegaE= xE**(self._beta-1.)
         SRE2= self._dfparams[2]**2.*sc.exp(-2.*(xE-1.)/self._dfparams[1])
-        return 1./2./sc.pi*sc.sqrt(2./(1.+self._beta))*sc.exp(-xE/self._dfparams[0])/SRE2*sc.exp(OmegaE*(L-LE)/SRE2)
+        if log:#BOVY: IS THIS CORRECT?
+            return -sc.log(2.*sc.pi*SRE2)+.5*sc.log(2./(1.+self._beta))-xE/self._dfparams[0]+OmegaE*(L-LE)/SRE2
+        else:
+            return 1./2./sc.pi*sc.sqrt(2./(1.+self._beta))*sc.exp(-xE/self._dfparams[0])/SRE2*sc.exp(OmegaE*(L-LE)/SRE2)
         
-    def _eval_corrected_dehnen(self,E,L):
+    def _eval_corrected_dehnen(self,E,L,log):
         """Internal function that evaluates the corrected Dehnen DF"""
         #Calculate Re,LE, OmegaE
         if self._beta == 0.:
@@ -86,16 +90,16 @@ class distF:
     def _calc_surfacemass(self,R):
[…]

### 2010-03-11 py/interpret_as_df.py
groups-simulations: calculating the surface mass density now works

@@ -2,6 +2,7 @@
 #   interpret_as_df.py: module that interprets (E,Lz) pairs in terms of a 
 #                        distribution function
 ###############################################################################
+import copy
 import os, os.path
 import cPickle as pickle
 import scipy as sc
@@ -32,7 +33,7 @@ class distF:
         self._beta= beta
         if dftype == 'corrected-dehnen':
             #Load corrections
-            corr= DFcorrection(**kwargs)
+            self._corr= DFcorrection(**kwargs)
         return None
 
     def eval(self,E,L,log=False):
@@ -66,8 +67,9 @@ class distF:
             xE= (2.*E/(1.+1./self._beta))**(1./2./self._beta)
             LE= xE**(self._beta+1.)
             OmegaE= xE**(self._beta-1.)
-        SRE2= self._dfparams[2]**2.*sc.exp(-2.*(xE-1.)/self._dfparams[1])
-        if log:#BOVY: IS THIS CORRECT?
+        SRE2= self._eval_SR2(xE)
+        #BOVY: REPLACE SURFACE MASS WITH FUNCTION
+        if log:
             return -sc.log(2.*sc.pi*SRE2)+.5*sc.log(2./(1.+self._beta))-xE/self._dfparams[0]+OmegaE*(L-LE)/SRE2
         else:
             return 1./2./sc.pi*sc.sqrt(2./(1.+self._beta))*sc.exp(-xE/self._dfparams[0])/SRE2*sc.exp(OmegaE*(L-LE)/SRE2)
@@ -83,23 +85,100 @@ class distF:
             xE= (2.*E/(1.+1./self._beta))**(1./2./self._beta)
             LE= xE**(self._beta+1.)
             OmegaE= xE*(self._beta-1.)
-        correction= corr.correct(xE)
-        SRE2= self._dfparams[2]**2.*sc.exp(-2.*(xE-1.)/self._dfparams[1])*correction[1]
+        correction= self._corr.correct(xE)
+        SRE2= self._eval_SR2(xE)*correction[1]
+        #BOVY: REPLACE SURFACE MASS WITH FUNCTION
         return 1./2./sc.pi*sc.sqrt(2./(1.+self._beta))*sc.exp(-xE/self._dfparams[0])/SRE2*sc.exp(OmegaE*(L-LE)/SRE2)*correction[0]
         
-    def _calc_surfacemass(self,R):
-        """Internal function that calculates the surface mass for a given DF at R"""
-        bound= 1.
[…]

### 2010-05-16 py/bar_figures.py
groups-simulations: start on figure 4 analyzing a single location

@@ -16,8 +16,8 @@ _degtorad= m.pi/180.
 _radtodeg= 180./m.pi
 _XWIDTH= 1.8*8/10/1.8/1.23
 _YWIDTH= 1.15*8/10/1.8/1.23
-#_XWIDTH= 1.8*80/100/1.8
-#_YWIDTH= 1.15*80/20/1.8
+_DEFAULTR= 1.
+_DEFAULTPHI=-1.
 def veldist_2d_Rphi(plotfilename,nx=10,ny=8,dx=_XWIDTH/20.,
                     dy=_YWIDTH/20.*1.8/1.15,
                     nsx=2,nsy=2,ngrid=101,rrange=[0.6,1.4],
@@ -134,11 +134,6 @@ def veldist_1d_Rphi(plotfilename,nx=10,ny=8,dx=_XWIDTH/20.,
                     phirange=[-m.pi/2.,m.pi/2.],
                     saveDir='../bar/1d/',normalize=True,
                     row=None):
-#def veldist_1d_Rphi(plotfilename,nx=100,ny=20,dx=_XWIDTH/20.,dy=_YWIDTH/20.,
-#                    ngrid=201,rrange=[0.7,1.3],
-#                    phirange=[-m.pi/2.,m.pi/2.],
-#                    saveDir='../bar/1dLarge/',normalize=True,
-#                    row=None):
     """
     NAME:
        veldist_1d_Rphi
@@ -251,6 +246,63 @@ def veldist_1d_Rphi(plotfilename,nx=10,ny=8,dx=_XWIDTH/20.,
     if not calcOnly:
         plot.bovy_end_print(plotfilename)
 
+def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
+                           ngrid=201,saveDir='../bar/1dvar/'):
+    """
+    NAME:
+       veldist_1d_barstrength
+    PURPOSE:
+       make a plot showing the influence of the bar strength
+    INPUT:
+       plotfilename - filename for figure
+       phi - Galactocentric azimuth
+       R - Galactocentric radius
+       ngrid - number of grid-points to calculate the los velocity distribution
+               on
+       saveDir - save pickles here
+    OUTPUT:
+       Figure in plotfilename
+    HISTORY:
+       2010-05-15 - Written - Bovy (NYU)
+    """
+    bar_strengths= [0.007,0.01,0.013]
+
[…]

### 2010-05-16 py/bar_detectability.py
groups-simulations: slope variation

@@ -11,7 +11,6 @@ from matplotlib.pyplot import contour, clabel
 _RADTODEG= 180./m.pi
 _XWIDTH= 1.8*80/100/1.8
 _YWIDTH= 1.15*80/20/1.8
-
 def bar_detectability(parser,
                       dx=_XWIDTH/20.,dy=_YWIDTH/20.,
                       nx=100,ny=20,
@@ -91,6 +90,16 @@ def bar_detectability(parser,
             detect[ii,jj]= probDistance.kullbackLeibler(vlosd,axivlosd,ddx,nan=True)
             losd[ii,jj]= m.sqrt(thisR**2.+1.-2.*thisR*m.cos(thisphi))
 
+    #Find maximum, further than 3 kpc away
+    detectformax= detect.flatten()
+    detectformax[losd.flatten() < 3./8.2]= 0.
+    x= sc.argmax(detectformax)
+    indx = sc.unravel_index(x,detect.shape)
+    maxR= (rrange[0]+(rrange[1]-rrange[0])/
+           (ny*_YWIDTH+(ny-1)*dy)*(indx[1]*(_YWIDTH+dy)+_YWIDTH/2.))
+    maxphi= (phirange[0]+(phirange[1]-phirange[0])/
+                      (nx*_XWIDTH+(nx-1)*dx)*(indx[0]*(_XWIDTH+dx)+_XWIDTH/2.))
+    print maxR, maxphi, losd[indx[0],indx[1]], detect[indx[0],indx[1]]
 
     #Now plot
     plot.bovy_print()

### 2010-05-16 py/bar_figures.py

@@ -16,8 +16,8 @@ _degtorad= m.pi/180.
 _radtodeg= 180./m.pi
 _XWIDTH= 1.8*8/10/1.8/1.23
 _YWIDTH= 1.15*8/10/1.8/1.23
-_DEFAULTR= 1.
-_DEFAULTPHI=-1.
+_DEFAULTR= 1.07517899761
+_DEFAULTPHI= -0.361455276723
 def veldist_2d_Rphi(plotfilename,nx=10,ny=8,dx=_XWIDTH/20.,
                     dy=_YWIDTH/20.*1.8/1.15,
                     nsx=2,nsy=2,ngrid=101,rrange=[0.6,1.4],
@@ -273,7 +273,7 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     vlosds= []
     basesavefilename= os.path.join(saveDir,'barstrength_')
     for bar_strength in bar_strengths:
-        savefilename= basesavefilename+'%.3f.sav' % bar_strength
+        thissavefilename= basesavefilename+'%.3f.sav' % bar_strength
         if os.path.exists(thissavefilename):
             print "Restoring los-velocity distribution at bar-strength %.3f" % bar_strength
             savefile= open(thissavefilename,'r')
@@ -289,16 +289,83 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                                pot='bar',beta=0.,
                                potparams=potparams)
             vlosd= vlosd/(sc.nansum(vlosd)*(vloss[1]-vloss[0]))
+            savefile= open(thissavefilename,'w')
+            pickle.dump(vlosd,savefile)
+            savefile.close()
         vlosds.append(vlosd)
     #Plot
     plot.bovy_print()
     plot.bovy_plot(vloss,vlosds[1],'k-',zorder=3,
                    xrange=[vloslinspace[0],vloslinspace[1]],
-                   yrange=[0.,sc.amax(vlosds.flatten())*1.1],
+                   yrange=[0.,sc.amax(sc.array(vlosds).flatten())*1.1],
+                   xlabel=r'$v_{\mathrm{los}} / v_0$')
+    plot.bovy_plot(vloss,vlosds[0],ls='-',color='0.75',
+                   overplot=True,zorder=2,lw=2.)
+    plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
[…]

### 2010-05-16 py/calc_veldist_1d.py

@@ -15,7 +15,7 @@ _EPSREL= 1.e-02
 _EPSABS= 1.e-05
 _NSIGMA=5.
 def plotVlos(vloslinspace,l=0.,d=1.,t=-4.,distCoord='GC',
-             pot='bar',beta=0.,potparams=(0.9,0.01,20.*_degtorad,.8,None),
+             pot='bar',beta=0.,potparams=(0.9,0.01,25.*_degtorad,.8,None),
              dfparams=(1./3.,1.,0.2),dftype='dehnen',correct=True,
              plotfilename='../bar/vlosdist.ps',
              savefilename='../bar/vlosdist.sav',
@@ -92,7 +92,7 @@ def plotVlos(vloslinspace,l=0.,d=1.,t=-4.,distCoord='GC',
     plot.bovy_end_print(plotfilename)
 
 def predictVlosConvolve(vloslinspace,l=0.,d=1.,t=-4.,distCoord='GC',
-                pot='bar',beta=0.,potparams=(0.9,0.01,20.*_degtorad,.8,None),
+                pot='bar',beta=0.,potparams=(0.9,0.01,25.*_degtorad,.8,None),
                 dfparams=(1./3.,1.,0.2),dftype='dehnen',correct=True,
                 convolve=0.1,nconvolve=101,sigmaconvolve=3):
     """
@@ -163,7 +163,7 @@ def predictVlosConvolve(vloslinspace,l=0.,d=1.,t=-4.,distCoord='GC',
     return vlosd/norm
 
 def predictVlos(vloslinspace,l=0.,d=1.,t=-4.,distCoord='GC',
-                pot='bar',beta=0.,potparams=(0.9,0.01,20.*_degtorad,.8,None),
+                pot='bar',beta=0.,potparams=(0.9,0.01,25.*_degtorad,.8,None),
                 dfparams=(1./3.,1.,0.2),dftype='dehnen',correct=True,
                 convolve=None,nconvolve=101,sigmaconvolve=3):
     """

### 2010-05-17 py/bar_figures.py
groups-simulations: add convolution figure

@@ -246,6 +246,75 @@ def veldist_1d_Rphi(plotfilename,nx=10,ny=8,dx=_XWIDTH/20.,
     if not calcOnly:
         plot.bovy_end_print(plotfilename)
 
+def veldist_1d_convolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
+                        ngrid=201,saveDir='../bar/1dvar/'):
+    """
+    NAME:
+       veldist_1d_convolve
+    PURPOSE:
+       make a plot showing the influence of the distance uncertainties
+    INPUT:
+       plotfilename - filename for figure
+       phi - Galactocentric azimuth
+       R - Galactocentric radius
+       ngrid - number of grid-points to calculate the los velocity distribution
+               on
+       saveDir - save pickles here
+    OUTPUT:
+       Figure in plotfilename
+    HISTORY:
+       2010-05-15 - Written - Bovy (NYU)
+    """
+    convolves= [0.,0.1,0.2]
+
+    vloslinspace= (-.9,.9,ngrid)
+    vloss= sc.linspace(*vloslinspace)
+
+    vlosds= []
+    basesavefilename= os.path.join(saveDir,'convolve_')
+    for distsig in convolves:
+        thissavefilename= basesavefilename+'%.1f.sav' % distsig
+        if os.path.exists(thissavefilename):
+            print "Restoring los-velocity distribution at distance uncertainties %.1f" % distsig
+            savefile= open(thissavefilename,'r')
+            vlosd= pickle.load(savefile)
+            savefile.close()
+        else:
+            print "Restoring los-velocity distribution at distance uncertainties %.1f" % distsig
+            potparams= (0.9,0.01,25.*_degtorad,.8,None)
+            if distsig == 0.:
+                vlosd= predictVlos(vloslinspace,
+                                   l=phi,
+                                   d=R,
+                                   distCoord='GCGC',
+                                   pot='bar',beta=0.,
+                                   potparams=potparams)
+            else:
+                vlosd= predictVlosConvolve(vloslinspace,
+                                           l=phi,
+                                           d=R,
[…]

### 2010-05-17 py/bar_figures.py
groups-simulations:typo

@@ -564,6 +564,6 @@ if __name__ == '__main__':
     elif options.df:
         veldist_1d_df(args[0],phi=options.phi,R=options.R)
     elif options.convolve:
-        veldist_1d_convovle(args[0],phi=options.phi,R=options.R)
+        veldist_1d_convolve(args[0],phi=options.phi,R=options.R)
     else:
         veldist_2d_Rphi(args[0])

### 2010-05-17 py/bar_figures.py
groups-simulations: calculating/restoring typo

@@ -280,7 +280,7 @@ def veldist_1d_convolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
             vlosd= pickle.load(savefile)
             savefile.close()
         else:
-            print "Restoring los-velocity distribution at distance uncertainties %.1f" % distsig
+            print "Calculating los-velocity distribution at distance uncertainties %.1f" % distsig
             potparams= (0.9,0.01,25.*_degtorad,.8,None)
             if distsig == 0.:
                 vlosd= predictVlos(vloslinspace,

### 2010-05-17 py/bar_figures.py
groups-simulations: import convolution prediction

@@ -11,7 +11,7 @@ import bovy_plot as plot
 from matplotlib import pyplot
 import matplotlib.ticker as ticker
 from calc_veldist_2d import calc_veldist_2d
-from calc_veldist_1d import predictVlos
+from calc_veldist_1d import predictVlos, predictVlosConvolve
 _degtorad= m.pi/180.
 _radtodeg= 180./m.pi
 _XWIDTH= 1.8*8/10/1.8/1.23

### 2010-05-17 py/bar_figures.py
groups-simulations: annotate bar strength figure

@@ -371,8 +371,12 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[0],ls='-',color='0.75',
                    overplot=True,zorder=2,lw=2.)
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
-                   overplot=True,zorder=2,lw=2.)
-    #BOVY: annotate
+                   overplot=True,zorder=2,lw=1.5)
+    plot.bovy_text(r'$\mathrm{bar\ strength}$',title=True)
+    plot.bovy_text(r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$',top_right=True)
+    pyplot.arrow(0.45,3.78,-0.41,-0.11,color='0.',lw=.75,ls='dashed')
+    pyplot.arrow(0.45,3.58,-0.4,-0.11,color='0.',lw=.75,ls='dashed')
+    pyplot.arrow(0.45,3.4,-0.4,-0.22,color='0.',lw=.75,ls='dashed')
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-05-17 py/bar_figures.py
groups-simulations: annotate slope figure

@@ -374,9 +374,9 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{bar\ strength}$',title=True)
     plot.bovy_text(r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$',top_right=True)
-    pyplot.arrow(0.45,3.78,-0.41,-0.11,color='0.',lw=.75,ls='dashed')
-    pyplot.arrow(0.45,3.58,-0.4,-0.11,color='0.',lw=.75,ls='dashed')
-    pyplot.arrow(0.45,3.4,-0.4,-0.22,color='0.',lw=.75,ls='dashed')
+    pyplot.arrow(0.45,3.78,-0.41,-0.09,color='0.',lw=.75,ls='dashed',zorder=4)
+    pyplot.arrow(0.45,3.60,-0.4,-0.13,color='0.',lw=.75,ls='dashed',zorder=4)
+    pyplot.arrow(0.45,3.4,-0.4,-0.22,color='0.',lw=.75,ls='dashed',zorder=4)
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
@@ -511,6 +511,16 @@ def veldist_1d_slope(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[4],ls='-',color='0.3',
                    overplot=True,zorder=2,lw=2.)
     #BOVY: annotate
+    plot.bovy_text(r'$\mathrm{shape\ of\ the\ rotation\ curve}$',title=True)
+    plot.bovy_text(r'$\beta = -0.2$'+'\n'+r'$\beta = -0.1$'+ '\n'+
+                   r'$\beta = \phantom{-}0.0$'+'\n'+
+                   r'$\beta= \phantom{-}0.1$'+'\n'+
+                   r'$\beta= \phantom{-}0.2$',top_right=True)
+    pyplot.arrow(0.45,3.85,-0.41,-0.09,color='0.',lw=.75,ls='dashed',zorder=4)
[…]

### 2010-05-17 py/calc_veldist_1d.py
groups-simulations: shuDF in the calc_veldist_1d code

@@ -5,7 +5,7 @@ import numpy as nu
 import scipy as sc
 import scipy.integrate as integrate
 from integrate_orbits import uvToELz
-from interpret_as_df import dehnenDF
+from interpret_as_df import dehnenDF, shuDF
 import bovy_plot as plot
 _DEBUG=True
 _degtorad= sc.pi/180.
@@ -293,8 +293,10 @@ def marginalizeAngle(vlos,alpha,R=1.,t=-4.,pot='bar',beta=0.,
     HISTORY:
        2010-04-04 - Written - Bovy (NYU)
     """
-    if dftype == 'dehnen':
+    if dftype.lower() == 'dehnen':
         thisDF= dehnenDF
+    elif dftype.lower() == 'shu':
+        thisDF= shuDF
     df= thisDF(profileParams=dfparams,beta=beta,correct=correct,
                niter=_NCORRECT)
     intLimit= _NSIGMA*m.sqrt(df.targetSigma2(R))

### 2010-05-17 py/interpret_as_df.py
groups-simulations: remove counter-rotating mass in Shu's DF

@@ -462,17 +462,18 @@ class shuDF(distFunc):
         #Calculate RL,LL, OmegaL
         if self._beta == 0.:
             xL= L
-            logECLE= sc.log(sc.log(xL)+0.5-E)
+            logECLE= sc.log(-sc.log(xL)-0.5+E)
         else: #non-flat rotation curve
             xL= L**(1./(self._beta+1.))
-            logECLE= sc.log(0.5*(1./self._beta+1.)*xL**(2.*self._beta)-E)
+            logECLE= sc.log(-0.5*(1./self._beta+1.)*xL**(2.*self._beta)+E)
+        if xL < 0.: #We must remove counter-rotating mass
+            return 0.
         if self._correct: 
             correction= self._corr.correct(xL,log=True)
         else:
             correction= sc.zeros(2)
         SRE2= self.targetSigma2(xL,log=True)+correction[1]
-        return sc.exp(logsigmaR2-SRE2+self.targetSurfacemass(xL,log=True)-logSigmaR+sc.exp(logECLE-SRE2)+correction[0])
-
+        return sc.exp(logsigmaR2-SRE2+self.targetSurfacemass(xL,log=True)-logSigmaR-sc.exp(logECLE-SRE2)+correction[0])
 
 def _surfaceIntegrand(vR,vT,R,df,logSigmaR,logsigmaR2,sigmaR1,gamma):
     """Internal function that is the integrand for the surface mass integration"""

### 2010-05-18 py/bar_figures.py
groups-simulations: annotate barstrength at the Hercules feature

@@ -373,10 +373,15 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{bar\ strength}$',title=True)
-    plot.bovy_text(r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$',top_right=True)
-    pyplot.arrow(0.45,3.78,-0.41,-0.09,color='0.',lw=.75,ls='dashed',zorder=4)
-    pyplot.arrow(0.45,3.60,-0.4,-0.13,color='0.',lw=.75,ls='dashed',zorder=4)
-    pyplot.arrow(0.45,3.4,-0.4,-0.22,color='0.',lw=.75,ls='dashed',zorder=4)
+    plot.bovy_text(0.5,1.,r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$')
+    pyplot.arrow(0.48,1.39,-0.14,-0.28,color='0.',lw=.75,zorder=4)
+    pyplot.arrow(0.48,1.22,-0.14,-0.23,color='0.',lw=.75,zorder=4)
+    pyplot.arrow(0.48,1.04,-0.14,-0.17,color='0.',lw=.75,zorder=4)
+
+    #plot.bovy_text(r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$',top_right=True)
+    #pyplot.arrow(0.45,3.78,-0.41,-0.09,color='0.',lw=.75,ls='dashed',zorder=4)
+    #pyplot.arrow(0.45,3.60,-0.4,-0.13,color='0.',lw=.75,ls='dashed',zorder=4)
+    #pyplot.arrow(0.45,3.4,-0.4,-0.22,color='0.',lw=.75,ls='dashed',zorder=4)
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-05-18 py/bar_figures.py
groups-simulations: skip the arrows for the annotation

@@ -373,10 +373,10 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{bar\ strength}$',title=True)
-    plot.bovy_text(0.5,1.,r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$')
-    pyplot.arrow(0.48,1.39,-0.14,-0.28,color='0.',lw=.75,zorder=4)
-    pyplot.arrow(0.48,1.22,-0.14,-0.23,color='0.',lw=.75,zorder=4)
-    pyplot.arrow(0.48,1.04,-0.14,-0.17,color='0.',lw=.75,zorder=4)
+    plot.bovy_text(0.5,.75,r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$')
+    #pyplot.arrow(0.48,1.39,-0.14,-0.28,color='0.',lw=.75,zorder=4)
+    #pyplot.arrow(0.48,1.22,-0.14,-0.23,color='0.',lw=.75,zorder=4)
+    #pyplot.arrow(0.48,1.04,-0.14,-0.17,color='0.',lw=.75,zorder=4)
 
     #plot.bovy_text(r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$',top_right=True)
     #pyplot.arrow(0.45,3.78,-0.41,-0.09,color='0.',lw=.75,ls='dashed',zorder=4)
@@ -517,15 +517,20 @@ def veldist_1d_slope(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    overplot=True,zorder=2,lw=2.)
     #BOVY: annotate
     plot.bovy_text(r'$\mathrm{shape\ of\ the\ rotation\ curve}$',title=True)
-    plot.bovy_text(r'$\beta = -0.2$'+'\n'+r'$\beta = -0.1$'+ '\n'+
+    plot.bovy_text(0.5,.5,r'$\beta = -0.2$'+'\n'+r'$\beta = -0.1$'+ '\n'+
                    r'$\beta = \phantom{-}0.0$'+'\n'+
[…]

### 2010-05-18 py/bar_figures.py
groups-simulations: annotate df figure

@@ -452,6 +452,16 @@ def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[4],ls='-',color='0.3',
                    overplot=True,zorder=2,lw=2.)
     #BOVY: annotate
+    plot.bovy_text(r'$\mathrm{distribution\ function}$',title=True)
+    plot.bovy_text(0.53,.3,r'$R_s = 0.25 R_0$'+'\n'
+                   +r'$R_{\sigma} = 2 R_s$'+'\n'
+                   +r'$\mathrm{fiducial}$'+'\n'
+                   +r'$\mathrm{Shu\ DF}$'+'\n'
+                   +r'$R_s = 0.4 R_0$',size=10.)
+    #pyplot.arrow(0.48,1.39,-0.14,-0.28,color='0.',lw=.75,zorder=4)
+    #pyplot.arrow(0.48,1.22,-0.14,-0.23,color='0.',lw=.75,zorder=4)
+    #pyplot.arrow(0.48,1.04,-0.14,-0.17,color='0.',lw=.75,zorder=4)
+
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_slope(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-05-18 py/bar_detectability.py
groups-simulations: titles for detectability figure

@@ -115,6 +115,12 @@ def bar_detectability(parser,
             aspect=(phirange[1]-phirange[0])*_RADTODEG/(rrange[1]-rrange[0]),
             extent=(phirange[0]*_RADTODEG,phirange[1]*_RADTODEG,
                     rrange[0],rrange[1]))
+    if options.skipCenter == 0.:
+        plot.bovy_text(r'$\mathrm{KL\ divergence\ / \ all}\ v_{\mathrm{los}}$',
+                       title=True)
+    else:
+        plot.bovy_text(r'$\mathrm{KL\ divergence\ / }\ |v_{\mathrm{los}}| \geq %.2f \ v_0$' % options.skipCenter,
+                       title=True)
     plot.bovy_end_print(args[0])
 
 def bar_detectability_convolve(parser,nconvsamples=1000,

### 2010-05-18 py/bar_figures.py

@@ -448,20 +448,15 @@ def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.60',
                    overplot=True,zorder=2,lw=2.)
     plot.bovy_plot(vloss,vlosds[3],ls='-',color='0.45',
-                   overplot=True,zorder=2,lw=2.)
+                   overplot=True,zorder=2,lw=1.5)
     plot.bovy_plot(vloss,vlosds[4],ls='-',color='0.3',
-                   overplot=True,zorder=2,lw=2.)
-    #BOVY: annotate
+                   overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{distribution\ function}$',title=True)
     plot.bovy_text(0.53,.3,r'$R_s = 0.25 R_0$'+'\n'
                    +r'$R_{\sigma} = 2 R_s$'+'\n'
                    +r'$\mathrm{fiducial}$'+'\n'
                    +r'$\mathrm{Shu\ DF}$'+'\n'
                    +r'$R_s = 0.4 R_0$',size=10.)
-    #pyplot.arrow(0.48,1.39,-0.14,-0.28,color='0.',lw=.75,zorder=4)
-    #pyplot.arrow(0.48,1.22,-0.14,-0.23,color='0.',lw=.75,zorder=4)
-    #pyplot.arrow(0.48,1.04,-0.14,-0.17,color='0.',lw=.75,zorder=4)
-
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_slope(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
@@ -522,9 +517,9 @@ def veldist_1d_slope(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[1],ls='-',color='0.60',
                    overplot=True,zorder=2,lw=2.)
     plot.bovy_plot(vloss,vlosds[3],ls='-',color='0.45',
-                   overplot=True,zorder=2,lw=2.)
+                   overplot=True,zorder=2,lw=1.5)
     plot.bovy_plot(vloss,vlosds[4],ls='-',color='0.3',
-                   overplot=True,zorder=2,lw=2.)
+                   overplot=True,zorder=2,lw=1.5)
[…]

### 2010-05-18 tex/hercules.tex
groups-simulations: start on ms

--- 2010-03-01 tex/ms.tex
@@
 
-\title{Probing Galactic structure with moving groups and radial velocity surveys}
+\title{Title}
 \author{Jo~Bovy}%\altaffilmark{1,2}}
@@
 structure of the inner Milky Way. However, at present the various
-models are hard to distinguish with observations of the local Galactic
-neighborhood. We show that future programs that survey the Milky Way
-disk's kinematics on a more global scale can distinguish between these
-models. By integrating orbits of stars in various steady-state and
-transient bar and spiral structure potentials that produce the pattern
-of moving groups observed in the Solar neighborhood, we predict the
-velocity distribution as a function of location in the Galactic
-disk. We identify the locations that distinguish most clearly between
-the various scenarios. Future infrared radial velocity surveys can
-probe the velocity distribution at these locations. We predict the
-distribution of radial velocities that can be directly compared to
-observations to establish the origin of the moving groups and to
-constrain the structural and dynamical parameters of the
-non-axisymmetric Milky Way potential.
+models are hard to confirm and distinguish with observations of the
+local Galactic neighborhood. We show that future programs that survey
+the Milky Way disk's kinematics on a more global scale can distinguish
+between these models. By integrating orbits of stars in various
+steady-state and transient bar and spiral structure potentials that
+produce the pattern of moving groups observed in the Solar
+neighborhood, we predict the velocity distribution as a function of
+location in the Galactic disk. We identify the locations that
+distinguish most clearly between the various scenarios. Future
+infrared radial velocity surveys can probe the velocity distribution
+at these locations. We predict the distribution of radial velocities
+that can be directly compared to observations to establish the origin
+of the moving groups and to constrain the structural and dynamical
+parameters of the non-axisymmetric Milky Way potential.
 \end{abstract}
@@
 
-\subsection{Steady-state spiral structure potential}
-
-\subsection{Transient spiral structure potential}
-
 \begin{thebibliography}{}
@@
 
+\clearpage
+\begin{figure}
+\includegraphics[width=\textwidth]{rphi2d.ps}
+\caption{2D}\label{fig:rphi2d}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\includegraphics[width=\textwidth]{rphi1d.ps}\\
+\includegraphics[width=\textwidth]{rphi1d2.ps}
+\caption{1D}\label{fig:rphi1d}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{detecta.ps}
+\includegraphics[width=0.5\textwidth]{detectb.ps}
+\caption{detect}\label{fig:detect}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{df.ps}\\
+\includegraphics[width=0.5\textwidth]{slope.ps}
+\includegraphics[width=0.5\textwidth]{barstrength.ps}
+\caption{1dvar}\label{fig:1dvar}
+\end{figure}
+
 \end{document}

### 2010-05-18 tex/hercules.tex
groups-simulations: add distance uncertainties panel
groups-simulations: references in document

@@
 \newcommand{\vv}{\tvector{v}}
+\newcommand{\Gaia}{\emph{Gaia}}
 
@@
 
-\title{Title}
+\title{Tracing the Hercules stream around the Galaxy}
 \author{Jo~Bovy}%\altaffilmark{1,2}}
@@
 \begin{abstract}
-Various dynamical scenarios related to the Milky Way's bar or spiral
-structure have been proposed to explain the existence of moving
-groups---clumps of co-moving stars---in the Solar neighborhood. Since
-the observed pattern of moving groups is very sensitive to the
-properties of the non-axisymmetric feature that creates them, these
-moving groups can be a powerful tool in constraining the dynamical
-structure of the inner Milky Way. However, at present the various
-models are hard to confirm and distinguish with observations of the
-local Galactic neighborhood. We show that future programs that survey
-the Milky Way disk's kinematics on a more global scale can distinguish
-between these models. By integrating orbits of stars in various
-steady-state and transient bar and spiral structure potentials that
-produce the pattern of moving groups observed in the Solar
-neighborhood, we predict the velocity distribution as a function of
-location in the Galactic disk. We identify the locations that
-distinguish most clearly between the various scenarios. Future
-infrared radial velocity surveys can probe the velocity distribution
-at these locations. We predict the distribution of radial velocities
-that can be directly compared to observations to establish the origin
-of the moving groups and to constrain the structural and dynamical
-parameters of the non-axisymmetric Milky Way potential.
+%Various dynamical scenarios related to the Milky Way's bar or spiral
+%structure have been proposed to explain the existence of moving
+%groups---clumps of co-moving stars---in the Solar neighborhood. Since
+%the observed pattern of moving groups is very sensitive to the
+%properties of the non-axisymmetric feature that creates them, these
+%moving groups can be a powerful tool in constraining the dynamical
+%structure of the inner Milky Way. However, at present the various
+%models are hard to confirm and distinguish with observations of the
+%local Galactic neighborhood. We show that future programs that survey
+%the Milky Way disk's kinematics on a more global scale can distinguish
+%between these models. By integrating orbits of stars in various
+%steady-state and transient bar and spiral structure potentials that
+%produce the pattern of moving groups observed in the Solar
+%neighborhood, we predict the velocity distribution as a function of
+%location in the Galactic disk. We identify the locations that
+%distinguish most clearly between the various scenarios. Future
+%infrared radial velocity surveys can probe the velocity distribution
+%at these locations. We predict the distribution of radial velocities
+%that can be directly compared to observations to establish the origin
+%of the moving groups and to constrain the structural and dynamical
+%parameters of the non-axisymmetric Milky Way potential.
 \end{abstract}
@@
 
+baryon acoustic feature, coincidence problem
+
 \section{Methodology}
 
-We follow Dehnen's approach \citep{dehnen00a}.
+We follow Dehnen's approach \citep{dehnen00a}. Only planar.
 
-\subsection{Bar potential}
+\section{Two-dimensional velocity distributions}
+
+\section{The line-of-sight velocity distribution}
+
+\section{Discussion}
+
+\subsection{\Gaia}
+
+\subsection{Radial velocity surveys}
+
+SDSS-III Project Description\footnote{Available at
+  \url{http://sdss3.org/collaboration/description.pdf}~.}
 
@@
 
-\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
-{Binney},~J. \& {Tremaine},~S., 2008, {Galactic Dynamics: Second Edition}
-  (Princeton University Press)
+%\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
+%{Binney},~J. \& {Tremaine},~S., 2008, {Galactic Dynamics: Second Edition}
+%  (Princeton University Press)
+
+\bibitem[Bailer-Jones(2008)]{bailerjones08a}
+  Bailer-Jones, ~C.~A.~L., 2008,
+  in IAU Symp.~254, ed.~J.~Andersen, J.~Bland-Hawthorn, and B.~Nordstr\"{o}m, (Dordrecht: Kluwer), 475
+
+\bibitem[Bensby \etal(2007)]{Bensby07a}
+  Bensby,~T., \etal, 2007,
+  \apjl, 655, L89
+
+\bibitem[Blaauw(1970)]{Blaauw70a}
+  Blaauw,~A., 1970, in IAU Symp.~38, ed.~W.~Becker \& I.~Kontopoulos (Dordrecht: Reidel), 199
+
+\bibitem[Blitz \& Spergel(1991)]{blitz91a}
+  Blitz,~L.~\& Spergel,~D.~N., 1991, \apj, 379, 631
+
+\bibitem[Binney, Gerhard, \& Spergel(1997)]{binney97a}
+  Binney,~J.~J., Gerhard,~O., \& Spergel,~D.~N., 1997,
+  \mnras, 288, 365
+
+\bibitem[Bissantz \& Gerhard(2002)]{bissantz02a}
+  Bissantz,~N.~\& Gerhard,~O., 2002,
+  \mnras, 330, 591
 
@@
 
+\bibitem[Bovy \& Hogg(2010)]{Bovy10a} Bovy,~J.~\& Hogg,~D.~W., 2010,
+  \apj, in press
+
+\bibitem[Caloi \etal(1999)]{caloi99a}
+  Caloi,~V.~, \etal, 1999,
+  \aap, 351, 925
+
+\bibitem[Cole \& Weinberg(2002)]{cole02a}
+  Cole,~A.~A.~\& Weinberg,~M.~D., 2002,
+  \apjl, 574, L43
+
+\bibitem[{{Dehnen}(1998)}]{Dehnen98b}
+Dehnen,~W., 1998, \aj, 115, 2384
+
+\bibitem[Dehnen(1999a)]{dehnen99a}
+  Dehnen,~W., 1999a, \aj, 118, 1190
+
+\bibitem[Dehnen(1999b)]{dehnen99b}
+  Dehnen,~W., 1999b, \aj, 118, 1201
+
+\bibitem[Dehnen(1999c)]{dehnen99c}
+  Dehnen,~W., 1999c, \apj, 524, L35
+
 \bibitem[Dehnen(2000)]{dehnen00a}
@@
 
+\bibitem[Fux(2001)]{fux01a}
+  Fux,~R., 2001,
+  \aap, 373, 511
+
+\bibitem[Gardner \& Flynn(2010)]{gardner10a}
+  Gardner,~E.~\& Flynn,~C., 2010,
+  \mnras, in press
+
+\bibitem[H\"{a}fner \etal(2000)]{hafner00}
+  H\"{a}fner,~R., \etal, 2000,
+  \mnras, 314, 433
+
+\bibitem[Minchev \& Famaey(2009)]{minchev09a}
+  Minchev,~I.~\& Famaey,~B., 2009,
+  \apj L, submitted
+
+\bibitem[Raboud \etal(1998)]{raboud98a}
+  Raboud,~D., \etal, 1998,
+  \aap, 335, L61
+
+\bibitem[Shen \etal(2010)]{Shen10a}
+  Shen,~J., \etal, 2010, \apj L, submitted
+
+\bibitem[Sellwood \& Binney(2002)]{sellwood02a}
+  Sellwood,~J.~A.~\& Binney,~J.J., 2002,
+  \mnras, 336, 785
+
+\bibitem[Shu(1969)]{shu69a}
+  Shu,~F.~H., 1969,
+  \apj, 158, 505
+
+\bibitem[Skuljan, Hearnshaw, \& Cottrell(1999)]{Skuljan99a}
+  Skuljan,~J., Hearnshaw,~J.~B., \& Cottrell,~P.~L., 1999, \mnras, 308, 731
+
 \end{thebibliography}
+
+\clearpage
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{veldist.eps}
+\caption{observed}\label{fig:obs}
+\end{figure}
 
@@
 \begin{figure}
+\includegraphics[width=0.5\textwidth]{distuncertain.ps}
 \includegraphics[width=0.5\textwidth]{df.ps}\\

### 2010-05-18 py/bar_detectability.py
groups-simulations: l contours + APOGEE contour

@@ -55,6 +55,7 @@ def bar_detectability(parser,
 
     detect= sc.zeros((nx,ny))
     losd= sc.zeros((nx,ny))
+    gall= sc.zeros((nx,ny))
     for ii in range(nx):
         for jj in range(ny):
             thisR= (rrange[0]+(rrange[1]-rrange[0])/
@@ -88,7 +89,14 @@ def bar_detectability(parser,
             vlosd[axivlosd_zeroindx]= 1.
             axivlosd[axivlosd_zeroindx]= 1.
             detect[ii,jj]= probDistance.kullbackLeibler(vlosd,axivlosd,ddx,nan=True)
-            losd[ii,jj]= m.sqrt(thisR**2.+1.-2.*thisR*m.cos(thisphi))
+            #los distance and Galactic longitude
+            d= m.sqrt(thisR**2.+1.-2.*thisR*m.cos(thisphi))
+            losd[ii,jj]= d
+            if 1./m.cos(thisphi) < thisR and m.cos(thisphi) > 0.:
+                l= m.pi-m.asin(thisR/d*m.sin(thisphi))
+            else:
+                l= m.asin(thisR/d*m.sin(thisphi))
+            gall[ii,jj]= l
 
     #Find maximum, further than 3 kpc away
     detectformax= detect.flatten()
@@ -99,7 +107,7 @@ def bar_detectability(parser,
            (ny*_YWIDTH+(ny-1)*dy)*(indx[1]*(_YWIDTH+dy)+_YWIDTH/2.))
     maxphi= (phirange[0]+(phirange[1]-phirange[0])/
                       (nx*_XWIDTH+(nx-1)*dx)*(indx[0]*(_XWIDTH+dx)+_XWIDTH/2.))
-    print maxR, maxphi, losd[indx[0],indx[1]], detect[indx[0],indx[1]]
+    print maxR, maxphi, losd[indx[0],indx[1]], detect[indx[0],indx[1]], gall[indx[0],indx[1]]*180./sc.pi
 
     #Now plot
     plot.bovy_print()
@@ -109,9 +117,24 @@ def bar_detectability(parser,
                      cmap='gist_yarg',xrange=sc.array(phirange)*_RADTODEG,
                      yrange=rrange,
                      aspect=(phirange[1]-phirange[0])*_RADTODEG/(rrange[1]-rrange[0]))
-    #contour the los distance
+    #contour the los distance and gall
[…]

### 2010-05-18 py/bar_figures.py

@@ -312,7 +312,8 @@ def veldist_1d_convolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    overplot=True,zorder=2,lw=2.)
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
                    overplot=True,zorder=2,lw=2.)
-    #BOVY: annotate
+    plot.bovy_text(r'$\mathrm{distance\ uncertainties}$',title=True)
+    plot.bovy_text(0.5,.65,r'$\sigma_d = 0$'+'\n'+r'$\sigma_d = 10 \%$'+ '\n'+r'$\sigma_d = 20 \%$')
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-05-19 tex/hercules.tex
groups-simulations: abstract

@@
 \begin{abstract}
-%Various dynamical scenarios related to the Milky Way's bar or spiral
-%structure have been proposed to explain the existence of moving
-%groups---clumps of co-moving stars---in the Solar neighborhood. Since
-%the observed pattern of moving groups is very sensitive to the
-%properties of the non-axisymmetric feature that creates them, these
-%moving groups can be a powerful tool in constraining the dynamical
-%structure of the inner Milky Way. However, at present the various
-%models are hard to confirm and distinguish with observations of the
-%local Galactic neighborhood. We show that future programs that survey
-%the Milky Way disk's kinematics on a more global scale can distinguish
-%between these models. By integrating orbits of stars in various
-%steady-state and transient bar and spiral structure potentials that
-%produce the pattern of moving groups observed in the Solar
-%neighborhood, we predict the velocity distribution as a function of
-%location in the Galactic disk. We identify the locations that
-%distinguish most clearly between the various scenarios. Future
-%infrared radial velocity surveys can probe the velocity distribution
-%at these locations. We predict the distribution of radial velocities
-%that can be directly compared to observations to establish the origin
-%of the moving groups and to constrain the structural and dynamical
-%parameters of the non-axisymmetric Milky Way potential.
+It has been proposed that the Hercules stream, a group of co-moving
+stars in the Solar neighborhood offset from the bulk of the velocity
+distribution, is the result of resonant interactions between stars in
+the outer disk and the Galactic bar. Only seen so far in the immediate
+Solar neighborhood, I predict the distribution of stellar velocities
+and the run of the Hercules stream as a function of location in the
+Galactic disk in a simple model for the Galaxy and the bar that
+produces the observed Hercules stream. I further show that the
+Hercules feature is strong enough that it can be unambiguously
+detected in the distribution of line-of-sight velocities in selected
+directions. I quantify the most promising lines-of-sight for the
+detection of the Hercules stream in radial velocities: The most
+promising directions to target are at $l \geq 250^{\circ}$. The
+predictions presented here are only weakly affected by distance
+uncertainties, assumptions about the distribution function of the old
+stellar disk, and the details of the Galactic potential including the
+effect of spiral structure; \Gaia\ and future spectroscopic surveys of
+the Galactic disk will be able to robustly test the origin of the
+Hercules stream and constrain the properties of the Galactic bar.
 \end{abstract}
 
-\keywords{
-  Galaxy: bulge
-  ---  
-  Galaxy: disk
-  ---
-  Galaxy: evolution
-  ---
-  Galaxy: fundamental parameters
-  ---
-  Galaxy: kinematics and dynamics
-  ---
-  Galaxy: structure
-}
+\keywords{Galaxy: bulge --- Galaxy: disk --- Galaxy: evolution ---
+  Galaxy: fundamental parameters --- Galaxy: kinematics and dynamics
+  --- Galaxy: structure }
 

### 2010-05-19 tex/hercules.tex
groups-simulations: abstract tweak

@@
 detected in the distribution of line-of-sight velocities in selected
-directions. I quantify the most promising lines-of-sight for the
-detection of the Hercules stream in radial velocities: The most
-promising directions to target are at $l \geq 250^{\circ}$. The
-predictions presented here are only weakly affected by distance
-uncertainties, assumptions about the distribution function of the old
-stellar disk, and the details of the Galactic potential including the
-effect of spiral structure; \Gaia\ and future spectroscopic surveys of
-the Galactic disk will be able to robustly test the origin of the
-Hercules stream and constrain the properties of the Galactic bar.
+directions. I quantify the most promising lines-of-sight for detection
+in radial velocities: The most promising directions to target are at
+$l \geq 250^{\circ}$. The predictions presented here are only weakly
+affected by distance uncertainties, assumptions about the distribution
+function of the old stellar disk, and the details of the Galactic
+potential including the effect of spiral structure; \Gaia\ and future
+spectroscopic surveys of the Galactic disk will be able to robustly
+test the origin of the Hercules stream and constrain the properties of
+the Galactic bar.
 \end{abstract}

### 2010-05-19 tex/hercules.tex
groups-simulations: figure captions

@@
 \newcommand{\Gaia}{\emph{Gaia}}
+\newcommand{\apogee}{APOGEE}
+\newcommand{\vR}{\ensuremath{v_R}}
+\newcommand{\vphi}{\ensuremath{v_{\phi}}}
+\newcommand{\vo}{\ensuremath{v_0}}
+\newcommand{\Ro}{\ensuremath{R_0}}
+\newcommand{\ro}{\Ro}
 
@@
 in radial velocities: The most promising directions to target are at
-$l \geq 250^{\circ}$. The predictions presented here are only weakly
-affected by distance uncertainties, assumptions about the distribution
-function of the old stellar disk, and the details of the Galactic
-potential including the effect of spiral structure; \Gaia\ and future
-spectroscopic surveys of the Galactic disk will be able to robustly
-test the origin of the Hercules stream and constrain the properties of
-the Galactic bar.
+$l \gtrsim 250^{\circ}$. The predictions presented here are only
+weakly affected by distance uncertainties, assumptions about the
+distribution function of the old stellar disk, and the details of the
+Galactic potential including the effect of spiral structure;
+\Gaia\ and future spectroscopic surveys of the Galactic disk will be
+able to robustly test the origin of the Hercules stream and constrain
+the properties of the Galactic bar.
 \end{abstract}
@@
   \url{http://sdss3.org/collaboration/description.pdf}~.}
+
+\acknowledgements David Hogg, Michael Blanton.
 
@@
 \includegraphics[width=0.5\textwidth]{veldist.eps}
-\caption{observed}\label{fig:obs}
+\caption{The velocity distribution of nearby ($\lesssim 100$ pc) stars
+  from \citet{Bovy09a}. The Hercules moving group is visible as an
+  overdensity at \vphi $\approx$ -0.2 $\vo$.}\label{fig:obs}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{rphi2d.ps}
-\caption{2D}\label{fig:rphi2d}
+\caption{Two-dimensional velocity distributions along the Solar
+  circle. The axes of each subpanel are the radial and tangential
+  velocities in Galactocentric cylindrical coordinates for the same
+  ranges as in \figurename~\ref{fig:obs}. The velocity distributions
+  on the other side of the Galaxy are obtained by shifting the
+  azimuths by 180$^{\circ}$. }\label{fig:rphi2d}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{rphi1d2.ps}
-\caption{1D}\label{fig:rphi1d}
+\caption{Line-of-sight velocity distributions along the Solar
+  circle. The range in line-of-sight velocities in each subpanel is as
+  in \figurename~\ref{fig:1dvar}. The gray curve in each panel is the
+  predicted distribution for an axisymmetric, steady-state
+  Galaxy.}\label{fig:rphi1d}
 \end{figure}
@@
 \includegraphics[width=0.5\textwidth]{detectb.ps}
-\caption{detect}\label{fig:detect}
+\caption{Kullback-Leibler divergence between the predicted
+  line-of-sight velocity distributions with and without the
+  bar. Darker values have a larger discriminatory potential. The left
+  panel uses the full line-of-sight velocity distributions, the right
+  panel excludes the main mode of the velocity distribution that is
+  likely to be affected by other dynamical and phase-space
+  effects. The concentric dashed lines indicate
+  constant-distance slices from the Sun, starting at 1 kpc and spaced
+  2 kpc apart.  The white dashed lines indicate the boundaries of the
+  four quadrants in Galactic longitude.  The white dashed-dotted line
+  shows the region of the sky that \apogee\ can see (-5$^{\circ} \leq
+  l \leq 250^{\circ}$).}\label{fig:detect}
 \end{figure}
@@
 \includegraphics[width=0.5\textwidth]{barstrength.ps}
-\caption{1dvar}\label{fig:1dvar}
+\caption{Variation of the predicted velocity distribution at R = 1.075
+  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
+  distance uncertainties, parameters and shape of the distribution
+  function, slope of the rotation curve, and bar
+  strength.}\label{fig:1dvar}
 \end{figure}

### 2010-05-19 tex/hercules.tex
groups-simulations: methodology section

@@
 \newcommand{\ro}{\Ro}
+\newcommand{\Ab}{\ensuremath{A_b}}
+\newcommand{\Rb}{\ensuremath{R_b}}
+\newcommand{\Omegab}{\ensuremath{\Omega_b}}
+\newcommand{\fdehnen}{\ensuremath{f_{\text{Dehnen}}}}
+\newcommand{\sigmaR}{\ensuremath{\sigma_R}}
+\newcommand{\rE}{\ensuremath{R_e}}
+\newcommand{\Lc}{\ensuremath{L_c}}
+\newcommand{\Rs}{\ensuremath{R_s}}
+\newcommand{\Rsigma}{\ensuremath{R_{\sigma}}}
+\newcommand{\Rolr}{\ensuremath{R_{\text{OLR}}}}
 
@@
 
-We follow Dehnen's approach \citep{dehnen00a}. Only planar.
+To simulate the effect of the bar on stellar orbits near the Solar
+circle, we follow the approach of \citet{dehnen00a}. For a fixed
+position in the Galaxy, we evaluate the velocity distribution function
+at values for the radial---toward the Galactic center---and
+tangential---in the direction of Galactic rotation---velocity by
+backward-integrating these velocities in the the Galactic potential to
+obtain the intial orbit before bar formation. We posit that at this
+epoch the old stellar disk was in a steady-state that can be described
+by a simple distribution function that is a function of the energy and
+angular momentum of the orbit alone. It follows from the collisionless
+Boltzmann equation that the value of the velocity distribution at the
+present position and velocity is equal to that of the steady-state
+distribution function before bar formation evaluated at the initial
+orbit.
+
+Since the Hercules stream is only apparent in the planar motions in
+the disk \citep[\eg][]{Bovy09a}, we only consider a two-dimensional
+model for the Galaxy. We use a simple power-law rotation curve to
+model the axisymmetric Galactic potential
+\begin{equation}
+v_c = \vo(R/\Ro)^\beta \,,
+\end{equation}
+and for most of the simulations below we specify this model further to
+a flat rotation curve ($\beta = 0$). The model for the bar is given by
+the following potential
+\begin{equation}
+\Phi_b(R,\phi) = \Ab(t) \cos [2(\phi - \Omegab t)] \times
+\left\{ \begin{array}{ll} -(\Rb/R)^3\,, & \mathrm{for}\ R \geq
+  \Rb\,,\\ (R/\Rb)^3-2\,, & \mathrm{for}\ R \leq \Rb\,. \end{array}
+\right.
+\end{equation}
+Here \Omegab\ is the pattern speed of the bar and \Rb\ is the bar
+radius, which is fixed to be 80\,percent of the bar's corotation
+radius. The bar is grown smoothly during a time $t_1$, always set to
+half of the total integration time in what follows, following the
+prescription
+\begin{equation}
+\Ab(t) = \alpha \, \frac{\vo^2}{3}\,\left(\frac{\Ro}{\Rb}\right)^3
+\times \left\{ \begin{array}{lll}
+\frac{3}{16}\xi^5 - \frac{5}{8} \xi^3 + \frac{15}{16} \xi + \frac{1}{2}\,, & \xi \equiv 2\frac{t}{t_1}-1\,, & t \leq t_1\,,\\
+1\,, & & t \geq t_1\,.\end{array} \right. 
+\end{equation}
+
+For the distribution function of the stellar disk before bar
+formation, we use a Dehnen distribution function \citep{dehnen99b}
+given by
+\begin{equation}
+\fdehnen(E,L) \propto \frac{\Sigma(\rE)}{\sigmaR^2(\rE)} \, \exp\left[ \frac{\Omega(\rE)\left[L-\Lc(E)\right]}{\sigmaR^2(\rE)}\right]\,,
+\end{equation}
+where \rE, \Lc, and $\Omega(\rE)$ are the radius, the angular
+momentum, and the angular frequency, respectively, of the circular
+orbit with energy $E$. Using the procedure given in \sectionname~3.2
+of \citet{dehnen99b}, we choose the $\Sigma(R)$ and $\sigmaR$
+functions such that they reproduce a disk with exponential surface
+density and velocity dispersion profiles 
+\begin{equation}
+\Sigma(R) = \Sigma_0 \exp\left(-R/\Rs\right)\,, \qquad 
+\sigmaR(R) = \sigma_0 \exp\left(-R/\Rsigma\right)\,,
+\end{equation}
+to an accuracy of a fraction of a percent at all radii. Our fiducial
+model for the distribution function has \Rs = \Ro /3 and \Rsigma =
+3\Rs, and $\sigma_0$ such that $\sigmaR(\Ro) = 0.2 \vo$.
+
+As shown by \citet{dehnen99c,dehnen00a}, the model described above
+that places the Sun at \Ro = 0.9\Rolr---where \Rolr\ is the radius at
+which a circular orbit is in the outer Lindblad resonance with the
+bar---at a current angle of 25$^{\circ}$ with the bar, with a bar
+strength $\alpha = 0.01$, and integrating for four bar periods,
+reproduces the Hercules feature in the local velocity distribution
+while agreeing with other photometric and kinematical observations of
+the bar. In what follows, unless indicated otherwise, we fix all of
+the parameters of the model at these fiducial values.
 
@@
 
-\bibitem[Dehnen(1999a)]{dehnen99a}
-  Dehnen,~W., 1999a, \aj, 118, 1190
+%\bibitem[Dehnen(1999a)]{dehnen99a}
+%  Dehnen,~W., 1999a, \aj, 118, 1190
 
-\bibitem[Dehnen(1999b)]{dehnen99b}
-  Dehnen,~W., 1999b, \aj, 118, 1201
+\bibitem[Dehnen(1999a)]{dehnen99b}
+  Dehnen,~W., 1999a, \aj, 118, 1201
 
-\bibitem[Dehnen(1999c)]{dehnen99c}
-  Dehnen,~W., 1999c, \apj, 524, L35
+\bibitem[Dehnen(1999b)]{dehnen99c}
+  Dehnen,~W., 1999b, \apj, 524, L35
 

### 2010-05-19 tex/hercules.tex
groups-simulations: 2d section

@@
 \newcommand{\apogee}{APOGEE}
+\newcommand{\hipparcos}{\emph{Hipparcos}}
 \newcommand{\vR}{\ensuremath{v_R}}
@@
 
-\section{Methodology}
+\section{Methodology}\label{sec:method}
 
@@
 
-\section{Two-dimensional velocity distributions}
+\section{The full planar velocity distribution in the disk}\label{sec:2d}
+
+\citet{dehnen00a} only simulated the velocity distribution at the
+position of the Sun for various combinations of the model parameters
+defined in the previous Section. Nevertheless, his results on how the
+velocity distribution at the Sun changes with the angle between the
+Sun-Galactic center line and the bar could also be interpreted as
+showing the velocity distribution on the Solar circle for different
+Galactocentric azimuths, fixing the bar angle. In this way, these
+simulations become predictions for what one might find when one
+studies the velocity distribution in locations beyond those probed by
+the \hipparcos\ mission.
+
+In \figurename~\ref{fig:rphi2d}, I present the predictions for the
+velocity distributions near the Solar circle for the full
+360$^{\circ}$ range in Galactocentric azimuths for the fiducial model
+from \sectionname~\ref{sec:method}. Only the range $-90^{\circ} \leq
+\phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
+symmetric with respect to rotations through 180$^{\circ}$, such that
+the predictions for the other side of the Galaxy can be obtained from
+those shown by shifting the azimuths by 180$^{\circ}$. The $R = \Ro$
+locus is not shown here, but it can be inferred from \figurename~2 in
+\citet{dehnen00a}.
+
+It is clear from this \figurename\ that we can trace the Hercules
+stream going around the Galaxy and when looking further in or further
+out. The effect of the bar is also not confined to a narrow range in
+radii, rather the effects of the bar and especially its outer Lindblad
+resonance are felt through a large range of radii (0.6 $\leq R/\Ro \leq
+1.4$). These effects are strong and robust predictions of the
+bar-origin model for the Hercules stream, especially in regions beyond
+the Solar radius, where the Hercules feature can be found
+significantly removed from the bulk of the velocity distribution,
+which might be severely shaped at all locations by the processes that
+create the pattern of low-velocity moving groups in the local velocity
+distribution (the main mode in \figurename~\ref{fig:obs}). As I
+discuss below, \Gaia\ will be able to easily verify these predictions.
 

### 2010-05-19 tex/hercules.tex
groups-simulations: discussion section; acks

@@
 \newcommand{\vv}{\tvector{v}}
+\newcommand{\dd}{\mathrm{d}}
 \newcommand{\Gaia}{\emph{Gaia}}
@@
 Since the Hercules stream is only apparent in the planar motions in
-the disk \citep[\eg][]{Bovy09a}, we only consider a two-dimensional
+the disk \citep[\eg,][]{Bovy09a}, we only consider a two-dimensional
 model for the Galaxy. We use a simple power-law rotation curve to
@@
 given by
-\begin{equation}
+\begin{equation}\label{eq:fdehnen}
 \fdehnen(E,L) \propto \frac{\Sigma(\rE)}{\sigmaR^2(\rE)} \, \exp\left[ \frac{\Omega(\rE)\left[L-\Lc(E)\right]}{\sigmaR^2(\rE)}\right]\,,
@@
 
+In what follows, all velocities are always considered to be with
+respect to their local standard of rest in the absence of a bar.
+
 \section{The full planar velocity distribution in the disk}\label{sec:2d}
@@
 360$^{\circ}$ range in Galactocentric azimuths for the fiducial model
-from \sectionname~\ref{sec:method}. Only the range $-90^{\circ} \leq
-\phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
-symmetric with respect to rotations through 180$^{\circ}$, such that
-the predictions for the other side of the Galaxy can be obtained from
-those shown by shifting the azimuths by 180$^{\circ}$. The $R = \Ro$
-locus is not shown here, but it can be inferred from \figurename~2 in
-\citet{dehnen00a}.
+from \sectionname~\ref{sec:method}. Only the range
+$-90^{\circ} \leq \phi \leq 90^{\circ}$ is shown here, since the
+barred Galaxy is symmetric with respect to rotations through
+180$^{\circ}$, such that the predictions for the other side of the
+Galaxy can be obtained from those shown by shifting the azimuths by
+180$^{\circ}$. The $R = \Ro$ locus is not shown here, but it can be
+inferred from \figurename~2 in \citet{dehnen00a}.
 
@@
 
-\section{The line-of-sight velocity distribution}
+\section{Line-of-sight velocity distributions}
+
+Rather than wait for a comprehensive survey of the disk of the Galaxy
+that measures the full phase-space position of a large number of stars
+by measuring proper motions, we can ask whether a targeted survey of
+line-of-sight velocities in selected regions of the Galaxy could see a
+clear signature of the Hercules feature. That this should be possible
+follows from the following argument. The Hercules feature stands out
+in the velocity distribution most in its Galactocentric tangential
+velocity $\vphi$. On the Solar circle near the Sun we see almost
+entirely this Galactocentric tangential velocity in the line-of-sight,
+so we would expect to see at least a hint of the Hercules stream in
+the line-of sight.
+
+In \figurename~\ref{fig:rphi1d} I show the predicted line-of-sight
+velocity distributions in various locations near the Solar circle all
+around the Galaxy, obtained by integrating the full two-dimensional
+velocity distribution over the tangential velocity. This
+\figurename\ is no longer symmetric with respect to shifts of
+180$^{\circ}$ degree in azimuth, since the line-of-sight direction
+breaks this symmetry. The gray curve is the line-of-sight distribution
+one would observe if the targeted stars were in a steady state in an
+axisymmetric potential with a Dehnen distribution function
+(\eqnname~[\ref{eq:fdehnen}]). Both line-of-sight velocity
+distributions are normalized to have equal mass since I assume that
+our knowledge of the Galaxy and the tracer population will not be
+great enough that we can test predicted differences in the absolute
+number of stars at a given location between the axisymmetric Galaxy
+and the barred Galaxy. These predictions are also largely a function
+of the particular form and parameters of the distribution function
+that we chose the represent the stellar disk before bar formation.
+
+We see that the prospects to see the Hercules feature are quite good
+on our side of the Galaxy. The Hercules feature is especially strong
+in the line-of-sight velocity distribution for Galactic longitudes $l
+\geq 250^{\circ}$. This is because of the fortunate coincidence that
+for those lines-of-sight we see velocities in the direction along
+which the Hercules streams separates most from the bulk of the
+velocity distribution (see \figurename~\ref{fig:rphi2d}) and that at
+these azimuths the Hercules stream is at the furthest distance from
+the rest of the velocity distribution. We see hardly any effect of the
+bar on the line-of-sight velocity distributions on the other side of
+the Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). This is due to
+the fact that on that side we see mostly the Galactocentric radial
+velocities. Even though upcoming surveys such as \apogee\ have the
+ability to probe the other side of the Galaxy, for the purposes of
+detecting the Hercules stream it is uninteresting.
+
+To further analyze the best directions to target to best see the
+Hercules feature in the line-of-sight velocities, I have calculated
+the line-of-sight velocity distribution on a 100 by 20 grid in
+Galactocentric azimuth and radius. As a measure of the difference
+between the barred Galaxy model and the steady-state, non-barred
+model, I have calculated the Kullback-Leibler divergence between these
+predicted distributions \citep[\eg,][]{mackay}
+\begin{equation}
+D_{\text{KL}}(P||Q) = \int \dd x\,p(x)\,\ln \frac{p(x)}{q(x)}\,,
+\end{equation}
+where, in this case, $x$ is the line-of-sight velocity, $p(x)$ is the
+predicted distribution for the bar-origin model of the Hercules
+stream, and $q(x)$ is the axisymmetric prediction. The
+Kullback-Leibler divergence is only one among many possible choices to
+quantify the difference between two probability distributions. It can
+be interpreted as the expected weight of evidence for the prediction
+by the barred Galaxy model over the fully axisymmetric prediction that
+can be expected from each sampled star. As such, locations with the
+maximal Kullback-Leibler divergence are those in which the Hercules
+feature stands out most prominent in the line-of-sight.
+
+The result of this analysis is shown in
+\figurename~\ref{fig:detect}. I present results both using the full
+line-of-sight velocity distributions and only considering the part of
+the distribution at $|v_{\text{los}}| \geq 0.15 \vo$. As is clear from
+\figurename~\ref{fig:obs}, the local velocity distribution contains
+many more features than the Hercules stream. These features are mostly
+contained to the main mode of the velocity distribution $|v| \lesssim
+0.15 \vo$. Observations indicate that these features are also the
+result of dynamical interactions \citep[\eg,][]{Bovy10a,sellwood10a},
+albeit with spiral arms rather than the bar. Therefore, we can expect
+these or similar features to be present in the velocity distribution
+at other locations in the Galaxy as well, and thus, if the effect of
+the Hercules stream occurs close to the center of the bulk of the
+velocity distribution it might be hard to disentangle from the other
+dynamical effects at play. However, these additional dynamical effects
+will be much less important for velocities $\gtrsim 0.15
+\vo$. \figurename~\ref{fig:detect} shows that there are many
+locations, mostly at $l \gtrsim 250^{\circ}$, where the Hercules
+stream is clearly offset from the bulk of the velocity distribution
+and should be unambiguously detectable.
+
+Finally, in \figurename~\ref{fig:1dvar} I show the influence of
+various nuisance and dynamical parameters on the predicted
+line-of-sight velocity distribution for the most promising
+line-of-sight and distance from the Sun that is more than 3 kpc
+away. We see that the effect of distance uncertainties is
+negligible. The exact shape and parameters of the stellar distribution
+function also only minimally affect the predicted line-of-sight
+velocity distribution. Changing the scalelength within its current
+uncertainty changes the prediction only in the amplitude of the
+effect: The Hercules stream is more prominent for shorter scalelengths
+since then there are relatively more stars in the inner Galaxy that
+are brought to the Solar circle by the interaction with the bar. Using
+a Shu distribution function \citep{shu69a} instead of the Dehnen
+distribution function also gives very similar predictions. Changing
+the axisymmetric potential by changing the shape of the rotation curve
+shifts the peak as expected \citep[see][]{dehnen00a}, but the Hercules
+feature nevertheless remains detectable.
 
@@
 
+The astrometric \Gaia\ mission will measure the parallaxes and proper
+motions of up to one billion stars, most of which will be disk
+stars. With parallax and proper motion accuracies down to 10
+microarcseconds, \Gaia\ will be able to probe the kinematics of the
+disk out to several kpc in all directions
+\citep{bailerjones08a}. Crucially, it will provide large samples of
+stars in many of the promising regions in
+\figurename~\ref{fig:rphi2d}, which will be able to verify the
+predictions laid out in this paper and will allow the dynamical
+properties of the bar to be tightly constrained, something that will
+be hard for \Gaia\ to do using direct observations of the bulge due to
+the large extinction toward the Galactic center.
+
 \subsection{Radial velocity surveys}
 
-SDSS-III Project Description\footnote{Available at
-  \url{http://sdss3.org/collaboration/description.pdf}~.}
+Since the extinction in the optical bandpasses is rather large in the
+directions identified as promising in \figurename~\ref{fig:detect}
+\citep[$A_V \approx 8$ mag toward $l \approx
+  270^{\circ}$;][]{marshall06a}, spectroscopic surveys in the
+near-infrared such as \apogee\ are the best bet for seeing the
+Hercules feature in the line-of-sight direction. The range of Galactic
+longitudes that \apogee\ can target along the disk is $-5^{\circ} \leq
+l \leq 250^{\circ}$ (SDSS-III Project Description\footnote{Available
+  at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
+\figurename~\ref{fig:detect} this limit seems to be right on the edge
+of where one might hope to detect the Hercules feature. Since
+\apogee\ is expected to start taking data in 2011 and to finish
+operations before the first \Gaia\ data release, \apogee\ could trace
+the Hercules stream around the Galaxy \Gaia. Since \apogee\ will also
+obtain detailed chemical abundances for its target stars, this would
+also allow for a more detailed analysis of the abundances of stars in
+the Hercules stream \citep[following, \eg,][]{Bensby07a,Bovy10a},
+which could shed more light on radial mixing in the Galactic disk.
 
-\acknowledgements David Hogg, Michael Blanton.
+\acknowledgements It is a pleasure to thank Mike Blanton and David
+Hogg for helpful discussions and assistance.  Financial support for
+this project was provided by the National Aeronautics and Space
+Administration (grant NNX08AJ48G) and the National Science Foundation
+(grant AST-0908357).
 
@@
 
+\bibitem[Mackay(2003)]{mackay}
+  Mackay,~D.~J.~C., 2003,
+  Information theory, inference, and learning algorithms (Cambridge University Press)
+
+\bibitem[Marshall \etal(2006)]{marshall06a}
+  Marshall,~D.~J.~, \etal, 2006,
+  \aap, 453, 635
+
 \bibitem[Minchev \& Famaey(2009)]{minchev09a}
@@
   Shen,~J., \etal, 2010, \apj L, submitted
+
+\bibitem[Sellwood(2010)]{sellwood10a}
+  Sellwood,~J.~A., 2010,
+  \mnras, submitted, 
+  arXiv:1001.5197v1 [astro-ph.GA]
 
@@
   on the other side of the Galaxy are obtained by shifting the
-  azimuths by 180$^{\circ}$. }\label{fig:rphi2d}
+  azimuths by 180$^{\circ}$.}\label{fig:rphi2d}
 \end{figure}

### 2010-05-20 tex/hercules.tex
groups-simulations: rearrange the reference

@@
 be hard for \Gaia\ to do using direct observations of the bulge due to
-the large extinction toward the Galactic center.
+the large extinction toward the Galactic center \citep{robin05a}.
 
@@
 
-\bibitem[Blaauw(1970)]{Blaauw70a}
-  Blaauw,~A., 1970, in IAU Symp.~38, ed.~W.~Becker \& I.~Kontopoulos (Dordrecht: Reidel), 199
-
-\bibitem[Blitz \& Spergel(1991)]{blitz91a}
-  Blitz,~L.~\& Spergel,~D.~N., 1991, \apj, 379, 631
-
 \bibitem[Binney, Gerhard, \& Spergel(1997)]{binney97a}
@@
   \mnras, 330, 591
+
+\bibitem[Blaauw(1970)]{Blaauw70a}
+  Blaauw,~A., 1970, in IAU Symp.~38, ed.~W.~Becker \& I.~Kontopoulos (Dordrecht: Reidel), 199
+
+\bibitem[Blitz \& Spergel(1991)]{blitz91a}
+  Blitz,~L.~\& Spergel,~D.~N., 1991, \apj, 379, 631
 
@@
 
-\bibitem[Shen \etal(2010)]{Shen10a}
-  Shen,~J., \etal, 2010, \apj L, submitted
+\bibitem[Robin \etal(2005)]{robin05a}
+  Robin,~A.~C., \etal, 2005,
+  \aap, 430, 129
 
@@
   \mnras, 336, 785
+
+\bibitem[Shen \etal(2010)]{Shen10a}
+  Shen,~J., \etal, 2010, \apj L, submitted
 

### 2010-05-20 tex/hercules.tex
groups-simulations: intro+references

@@
 
-baryon acoustic feature, coincidence problem
+The velocity distribution of nearby stars contains various large
+overdensities of co-moving stars that cannot be explained as the
+vestiges of the clumpy nature of star formation
+\citep[\eg,][]{Bovy10a}. In particular, the Hercules moving group, a
+group of stars significantly offset from the bulk of the observed
+velocity distribution \citep[\eg,][see
+  \figurename~\ref{fig:obs}]{Dehnen98b,Bovy10a}, displays a wide range
+of ages and metallicities
+\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
+that has it originating through resonant interactions of stars in the
+outer disk with the bar in the central region of the Galaxy has been
+put forward \citep{dehnen00a,fux01a}, but any confirmation of this
+picture beyond the locallly observed kinematics of stars remains
+lacking. An investigation of the Hercules members' color--magnitude
+diagram found no evidence for any significant metallicity anomaly with
+respect to other local stars \citep{Bovy10a}, even though such a
+difference should be expected as the Hercules stars originate from a
+few kiloparsec toward the Galactic center where the average
+metallicity is higher than in the Solar neighborhood. The absence of
+the predicted metallicity anomaly could potentially be explained
+through, \eg, radial mixing \citep{sellwood02a}, but it is clear that
+chemodynamical modeling of the Galactic disk is currently not up to
+the task of separating out these effects.
+
+In this Letter I propose that a robust and unambiguous test of the
+bar-resonance model for the Hercules stream is the distinct pattern it
+predicts as we trace the velocity distribution of stars near the Solar
+radius around the Galaxy. \Gaia\ is in a good position to have the
+final word on this, but I also show that the Hercules feature can be
+detected in the line-of-sight velocity distribution in selected
+directions on the sky. Clear signatures should exist in regions a few
+kiloparsec from the Sun, allowing for a clean verification of the
+bar-resonance model for the Hercules stream.
 
@@
 \end{equation}
+A comparison of the local velocity distribution obtained by using this
+simple bar potential with that resulting from the more realistic bar
+potential of \citet{gardner10a} shows that this simple model is
+sufficient to model the bar's influence on stellar orbits in the outer
+disk.
 
@@
 while agreeing with other photometric and kinematical observations of
-the bar. In what follows, unless indicated otherwise, we fix all of
-the parameters of the model at these fiducial values.
+the bar \citep[\eg,][]{binney97a,bissantz02a,cole02a,Shen10a}. In what
+follows, unless indicated otherwise, we fix all of the parameters of
+the model at these fiducial values.
 
@@
 great enough that we can test predicted differences in the absolute
-number of stars at a given location between the axisymmetric Galaxy
-and the barred Galaxy. These predictions are also largely a function
-of the particular form and parameters of the distribution function
-that we chose the represent the stellar disk before bar formation.
+number of stars at a given location between the axisymmetric and the
+bar-resonance predictions. These predictions are also largely a
+function of the particular form and parameters of the distribution
+function that we chose the represent the stellar disk before bar
+formation.
 
@@
 Galactocentric azimuth and radius. As a measure of the difference
-between the barred Galaxy model and the steady-state, non-barred
+between the bar-resonance model and the steady-state, axisymmetric
 model, I have calculated the Kullback-Leibler divergence between these
@@
 be interpreted as the expected weight of evidence for the prediction
-by the barred Galaxy model over the fully axisymmetric prediction that
+by the bar-resonance model over the fully axisymmetric prediction that
 can be expected from each sampled star. As such, locations with the
@@
 
-%\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
-%{Binney},~J. \& {Tremaine},~S., 2008, {Galactic Dynamics: Second Edition}
-%  (Princeton University Press)
-
 \bibitem[Bailer-Jones(2008)]{bailerjones08a}
@@
 
-\bibitem[Blitz \& Spergel(1991)]{blitz91a}
-  Blitz,~L.~\& Spergel,~D.~N., 1991, \apj, 379, 631
-
 \bibitem[Bovy, Hogg, \& Roweis(2009)]{Bovy09a} Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T., 2009,
@@
 Dehnen,~W., 1998, \aj, 115, 2384
-
-%\bibitem[Dehnen(1999a)]{dehnen99a}
-%  Dehnen,~W., 1999a, \aj, 118, 1190
 
@@
 
-\bibitem[H\"{a}fner \etal(2000)]{hafner00}
-  H\"{a}fner,~R., \etal, 2000,
-  \mnras, 314, 433
-
 \bibitem[Mackay(2003)]{mackay}
@@
 
-\bibitem[Minchev \& Famaey(2009)]{minchev09a}
-  Minchev,~I.~\& Famaey,~B., 2009,
-  \apj L, submitted
-
 \bibitem[Raboud \etal(1998)]{raboud98a}
@@
   \aap, 430, 129
-
-\bibitem[Sellwood(2010)]{sellwood10a}
-  Sellwood,~J.~A., 2010,
-  \mnras, submitted, 
-  arXiv:1001.5197v1 [astro-ph.GA]
 
@@
 
-\bibitem[Skuljan, Hearnshaw, \& Cottrell(1999)]{Skuljan99a}
-  Skuljan,~J., Hearnshaw,~J.~B., \& Cottrell,~P.~L., 1999, \mnras, 308, 731
-
 \end{thebibliography}
@@
 \caption{The velocity distribution of nearby ($\lesssim 100$ pc) stars
-  from \citet{Bovy09a}. The Hercules moving group is visible as an
-  overdensity at \vphi $\approx$ -0.2 $\vo$.}\label{fig:obs}
+  from \citet{Bovy09a}. \vR\ and \vphi\ are the velocities toward the
+  Galactic center and in the direction of Galactic rotation,
+  respectively. The Hercules moving group is visible as an overdensity
+  at \vphi $\approx$ -0.2 $\vo$.}\label{fig:obs}
 \end{figure}

### 2010-05-20 tex/hercules.tex
groups-simulations: small tweaks

@@
 directions. I quantify the most promising lines-of-sight for detection
-in radial velocities: The most promising directions to target are at
-$l \gtrsim 250^{\circ}$. The predictions presented here are only
-weakly affected by distance uncertainties, assumptions about the
-distribution function of the old stellar disk, and the details of the
-Galactic potential including the effect of spiral structure;
-\Gaia\ and future spectroscopic surveys of the Galactic disk will be
-able to robustly test the origin of the Hercules stream and constrain
-the properties of the Galactic bar.
+in radial velocities; These are at $l \gtrsim 250^{\circ}$. The
+predictions presented here are only weakly affected by distance
+uncertainties, assumptions about the distribution function of the
+stellar disk, and the details of the Galactic potential including the
+effect of spiral structure; \Gaia\ and future spectroscopic surveys of
+the Galactic disk will be able to robustly test the origin of the
+Hercules stream and constrain the properties of the Galactic bar.
 \end{abstract}
@@
 put forward \citep{dehnen00a,fux01a}, but any confirmation of this
-picture beyond the locallly observed kinematics of stars remains
+picture beyond the locallly observed stellar kinematics remains
 lacking. An investigation of the Hercules members' color--magnitude
@@
 tangential---in the direction of Galactic rotation---velocity by
-backward-integrating these velocities in the the Galactic potential to
+backward-integrating these velocities in the Galactic potential to
 obtain the intial orbit before bar formation. We posit that at this
@@
 \begin{equation}
-v_c = \vo(R/\Ro)^\beta \,,
+v_c(R) = \vo(R/\Ro)^\beta \,,
 \end{equation}
@@
 \end{equation}
-where \rE, \Lc, and $\Omega(\rE)$ are the radius, the angular
-momentum, and the angular frequency, respectively, of the circular
-orbit with energy $E$. Using the procedure given in \sectionname~3.2
-of \citet{dehnen99b}, we choose the $\Sigma(R)$ and $\sigmaR$
+where \rE, \Lc, and $\Omega(\rE)$ are the radius, angular momentum,
+and angular frequency, respectively, of the circular orbit with energy
+$E$. Using the procedure given in \sectionname~3.2 of
+\citet{dehnen99b}, we choose the $\Sigma(R)$ and $\sigmaR(R)$
 functions such that they reproduce a disk with exponential surface
@@
 As shown by \citet{dehnen99c,dehnen00a}, the model described above
-that places the Sun at \Ro = 0.9\Rolr---where \Rolr\ is the radius at
-which a circular orbit is in the outer Lindblad resonance with the
-bar---at a current angle of 25$^{\circ}$ with the bar, with a bar
-strength $\alpha = 0.01$, and integrating for four bar periods,
-reproduces the Hercules feature in the local velocity distribution
-while agreeing with other photometric and kinematical observations of
-the bar \citep[\eg,][]{binney97a,bissantz02a,cole02a,Shen10a}. In what
+with the Sun at \Ro = 0.9\Rolr---where \Rolr\ is the radius at which a
+circular orbit is in the outer Lindblad resonance with the bar---at a
+current angle of 25$^{\circ}$ with the bar, with a bar strength
+$\alpha = 0.01$, and integrating for four bar periods, reproduces the
+Hercules feature in the local velocity distribution while agreeing
+with other photometric and kinematical observations of the bar
+\citep[\eg,][]{binney97a,bissantz02a,cole02a,Shen10a}. In what
 follows, unless indicated otherwise, we fix all of the parameters of
-the model at these fiducial values.
-
-In what follows, all velocities are always considered to be with
-respect to their local standard of rest in the absence of a bar.
+the model at these fiducial values. All velocities are always
+considered to be with respect to their local standard of rest in the
+absence of a bar.
 
@@
 simulations become predictions for what one might find when one
-studies the velocity distribution in locations beyond those probed by
+studies the velocity distribution in regions beyond those probed by
 the \hipparcos\ mission.
@@
 velocity distributions near the Solar circle for the full
-360$^{\circ}$ range in Galactocentric azimuths for the fiducial model
-from \sectionname~\ref{sec:method}. Only the range
-$-90^{\circ} \leq \phi \leq 90^{\circ}$ is shown here, since the
-barred Galaxy is symmetric with respect to rotations through
-180$^{\circ}$, such that the predictions for the other side of the
-Galaxy can be obtained from those shown by shifting the azimuths by
-180$^{\circ}$. The $R = \Ro$ locus is not shown here, but it can be
-inferred from \figurename~2 in \citet{dehnen00a}.
+360$^{\circ}$ range in Galactocentric azimuth for the fiducial model
+from \sectionname~\ref{sec:method}. Only the range $-90^{\circ} \leq
+\phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
+symmetric with respect to rotations through 180$^{\circ}$, such that
+the predictions for the other side of the Galaxy can be obtained from
+those shown by shifting the azimuths by 180$^{\circ}$. The $R = \Ro$
+locus is not shown here, but it can be inferred from \figurename~2 in
+\citet{dehnen00a}.
 
 It is clear from this \figurename\ that we can trace the Hercules
-stream going around the Galaxy and when looking further in or further
-out. The effect of the bar is also not confined to a narrow range in
-radii, rather the effects of the bar and especially its outer Lindblad
-resonance are felt through a large range of radii (0.6 $\leq R/\Ro \leq
-1.4$). These effects are strong and robust predictions of the
-bar-origin model for the Hercules stream, especially in regions beyond
-the Solar radius, where the Hercules feature can be found
-significantly removed from the bulk of the velocity distribution,
-which might be severely shaped at all locations by the processes that
-create the pattern of low-velocity moving groups in the local velocity
-distribution (the main mode in \figurename~\ref{fig:obs}). As I
-discuss below, \Gaia\ will be able to easily verify these predictions.
+stream going around the Galaxy and when looking further in toward the
+Galactic center or further out. The effect of the bar is also not
+confined to a narrow range in radii, rather the effects of the bar and
+especially its outer Lindblad resonance are felt through a large range
+of radii (0.6 $\leq R/\Ro \leq 1.4$). These effects are strong and
+robust predictions of the bar-origin model for the Hercules stream,
+especially in regions beyond the Solar radius, where the Hercules
+feature can be found significantly removed from the bulk of the
+velocity distribution, which might be severely shaped at all locations
+by the processes that create the pattern of low-velocity moving groups
+in the local velocity distribution (the main mode in
+\figurename~\ref{fig:obs}). As I discuss below, \Gaia\ will be able to
+verify these predictions.
 
@@
 
-Rather than wait for a comprehensive survey of the disk of the Galaxy
-that measures the full phase-space position of a large number of stars
-by measuring proper motions, we can ask whether a targeted survey of
+Rather than wait for a comprehensive survey of the Galactic disk that
+measures the full phase-space position of a large number of stars by
+measuring proper motions, we can ask whether a targeted survey of
 line-of-sight velocities in selected regions of the Galaxy could see a
@@
 follows from the following argument. The Hercules feature stands out
-in the velocity distribution most in its Galactocentric tangential
-velocity $\vphi$. On the Solar circle near the Sun we see almost
-entirely this Galactocentric tangential velocity in the line-of-sight,
-so we would expect to see at least a hint of the Hercules stream in
-the line-of sight.
+in the velocity distribution most in its tangential velocity
+$\vphi$. On the Solar circle near the Sun we see almost entirely this
+Galactocentric tangential velocity in the line-of-sight, so we would
+expect to see at least a hint of the Hercules stream in the line-of
+sight.
 
@@
 number of stars at a given location between the axisymmetric and the
-bar-resonance predictions. These predictions are also largely a
+bar-resonance predictions. These differences are also largely a
 function of the particular form and parameters of the distribution
-function that we chose the represent the stellar disk before bar
+function that we choose the represent the stellar disk before bar
 formation.
@@
 the fact that on that side we see mostly the Galactocentric radial
-velocities. Even though upcoming surveys such as \apogee\ have the
-ability to probe the other side of the Galaxy, for the purposes of
-detecting the Hercules stream it is uninteresting.
+velocities, in which the Hercules feature does not separate clearly
+from the bulk of the stars. Even though upcoming surveys such as
+\apogee\ have the ability to probe the other side of the Galaxy, for
+the purposes of detecting the Hercules stream this is uninteresting.
 
@@
 maximal Kullback-Leibler divergence are those in which the Hercules
-feature stands out most prominent in the line-of-sight.
+feature stands out most prominently in the line-of-sight.
 
@@
 operations before the first \Gaia\ data release, \apogee\ could trace
-the Hercules stream around the Galaxy \Gaia. Since \apogee\ will also
-obtain detailed chemical abundances for its target stars, this would
-also allow for a more detailed analysis of the abundances of stars in
-the Hercules stream \citep[following, \eg,][]{Bensby07a,Bovy10a},
-which could shed more light on radial mixing in the Galactic disk.
+the Hercules stream around the Galaxy before \Gaia\ does. Since
+\apogee\ will also obtain detailed chemical abundances for its target
+stars, this would also allow for a more detailed analysis of the
+abundances of stars in the Hercules stream \citep[following,
+  \eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
+mixing in the Galactic disk.
 
@@
 \bibitem[Caloi \etal(1999)]{caloi99a}
-  Caloi,~V.~, \etal, 1999,
+  Caloi,~V., \etal, 1999,
   \aap, 351, 925
@@
 \bibitem[Marshall \etal(2006)]{marshall06a}
-  Marshall,~D.~J.~, \etal, 2006,
+  Marshall,~D.~J., \etal, 2006,
   \aap, 453, 635
@@
   \aap, 430, 129
+
+\bibitem[Sellwood(2010)]{sellwood10a}
+  Sellwood,~J.~A., 2010,
+  \mnras, submitted, 
+  arXiv:1001.5197v1 [astro-ph.GA]
 
@@
   Galactic center and in the direction of Galactic rotation,
-  respectively. The Hercules moving group is visible as an overdensity
-  at \vphi $\approx$ -0.2 $\vo$.}\label{fig:obs}
+  respectively, corrected for the Solar motion. The Hercules moving
+  group is visible as an overdensity at \vphi $\approx$ -0.2
+  $\vo$.}\label{fig:obs}
 \end{figure}
@@
 \caption{Two-dimensional velocity distributions along the Solar
-  circle. The axes of each subpanel are the radial and tangential
-  velocities in Galactocentric cylindrical coordinates for the same
-  ranges as in \figurename~\ref{fig:obs}. The velocity distributions
-  on the other side of the Galaxy are obtained by shifting the
-  azimuths by 180$^{\circ}$.}\label{fig:rphi2d}
+  circle. The axes of each subpanel are the same as in
+  \figurename~\ref{fig:obs}. The velocity distributions on the other
+  side of the Galaxy are obtained by shifting the azimuths by
+  180$^{\circ}$.}\label{fig:rphi2d}
 \end{figure}
@@
   distance uncertainties, parameters and shape of the distribution
-  function, slope of the rotation curve, and bar
+  function, shape of the rotation curve, and bar
   strength.}\label{fig:1dvar}

### 2010-05-21 tex/hercules.tex
groups-simulations: minor tweaks

@@
 function of the particular form and parameters of the distribution
-function that we choose the represent the stellar disk before bar
+function that we choose to represent the stellar disk before bar
 formation.
 
-We see that the prospects to see the Hercules feature are quite good
-on our side of the Galaxy. The Hercules feature is especially strong
-in the line-of-sight velocity distribution for Galactic longitudes $l
-\geq 250^{\circ}$. This is because of the fortunate coincidence that
-for those lines-of-sight we see velocities in the direction along
-which the Hercules streams separates most from the bulk of the
-velocity distribution (see \figurename~\ref{fig:rphi2d}) and that at
-these azimuths the Hercules stream is at the furthest distance from
-the rest of the velocity distribution. We see hardly any effect of the
-bar on the line-of-sight velocity distributions on the other side of
-the Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). This is due to
-the fact that on that side we see mostly the Galactocentric radial
-velocities, in which the Hercules feature does not separate clearly
-from the bulk of the stars. Even though upcoming surveys such as
+The prospects to detect the Hercules feature are quite good on our
+side of the Galaxy. The Hercules feature is especially strong in the
+line-of-sight velocity distribution for Galactic longitudes $l \geq
+250^{\circ}$. This is because of the fortunate coincidence that for
+those lines-of-sight we see velocities in the direction along which
+the Hercules streams separates most from the bulk of the velocity
+distribution (see \figurename~\ref{fig:rphi2d}) and that at these
+azimuths the Hercules stream is at the furthest distance from the rest
+of the velocity distribution. We see hardly any effect of the bar on
+the line-of-sight velocity distributions on the other side of the
+Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). This is due to the
+fact that on that side we mostly see the Galactocentric radial
+velocity, in which the Hercules feature does not separate clearly from
+the bulk of the stars. Even though upcoming surveys such as
 \apogee\ have the ability to probe the other side of the Galaxy, for
@@
 
-To further analyze the best directions to target to best see the
-Hercules feature in the line-of-sight velocities, I have calculated
-the line-of-sight velocity distribution on a 100 by 20 grid in
+To further analyze the directions to target to best see the Hercules
+feature in the line-of-sight velocities, I have calculated the
+line-of-sight velocity distribution on a 100 by 20 grid in
 Galactocentric azimuth and radius. As a measure of the difference
@@
 the Hercules stream occurs close to the center of the bulk of the
-velocity distribution it might be hard to disentangle from the other
+velocity distribution, it might be hard to disentangle from the other
 dynamical effects at play. However, these additional dynamical effects
@@
 function also only minimally affect the predicted line-of-sight
-velocity distribution. Changing the scalelength within its current
+velocity distribution. Changing the scale-length within its current
 uncertainty changes the prediction only in the amplitude of the
-effect: The Hercules stream is more prominent for shorter scalelengths
-since then there are relatively more stars in the inner Galaxy that
-are brought to the Solar circle by the interaction with the bar. Using
-a Shu distribution function \citep{shu69a} instead of the Dehnen
-distribution function also gives very similar predictions. Changing
-the axisymmetric potential by changing the shape of the rotation curve
-shifts the peak as expected \citep[see][]{dehnen00a}, but the Hercules
-feature nevertheless remains detectable.
+effect: The Hercules stream is more prominent for shorter
+scale-lengths since then there are relatively more stars in the inner
+Galaxy that are brought to the Solar circle by the interaction with
+the bar. Using a Shu distribution function \citep{shu69a} instead of
+the Dehnen distribution function also gives very similar
+predictions. Changing the axisymmetric potential by adjusting the
+shape of the rotation curve shifts the peak as expected
+\citep[see][]{dehnen00a}, but the Hercules feature nevertheless
+remains detectable.
 
@@
 \figurename~\ref{fig:rphi2d}, which will be able to verify the
-predictions laid out in this paper and will allow the dynamical
+predictions laid out in this paper. This will allow the dynamical
 properties of the bar to be tightly constrained, something that will
@@
   270^{\circ}$;][]{marshall06a}, spectroscopic surveys in the
-near-infrared such as \apogee\ are the best bet for seeing the
+near-infrared such as \apogee\ are the best option for seeing the
 Hercules feature in the line-of-sight direction. The range of Galactic
@@
   at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
-\figurename~\ref{fig:detect} this limit seems to be right on the edge
-of where one might hope to detect the Hercules feature. Since
-\apogee\ is expected to start taking data in 2011 and to finish
-operations before the first \Gaia\ data release, \apogee\ could trace
-the Hercules stream around the Galaxy before \Gaia\ does. Since
-\apogee\ will also obtain detailed chemical abundances for its target
-stars, this would also allow for a more detailed analysis of the
-abundances of stars in the Hercules stream \citep[following,
-  \eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
-mixing in the Galactic disk.
+\figurename~\ref{fig:detect} this limit appears right on the edge of
+where one might hope to detect the Hercules feature. Since \apogee\ is
+expected to start taking data in 2011 and to finish operations before
+the first \Gaia\ data release, \apogee\ could trace the Hercules
+stream around the Galaxy before \Gaia\ does. Since \apogee\ will also
+obtain detailed chemical abundances for its target stars, this would
+also allow for a more detailed analysis of the abundances of stars in
+the Hercules stream \citep[following, \eg,][]{Bensby07a,Bovy10a},
+which could shed more light on radial mixing in the Galactic disk.
 
@@
   likely to be affected by other dynamical and phase-space
-  effects. The concentric dashed lines indicate
-  constant-distance slices from the Sun, starting at 1 kpc and spaced
-  2 kpc apart.  The white dashed lines indicate the boundaries of the
-  four quadrants in Galactic longitude.  The white dashed-dotted line
-  shows the region of the sky that \apogee\ can see (-5$^{\circ} \leq
-  l \leq 250^{\circ}$).}\label{fig:detect}
+  effects. The concentric dashed lines indicate constant-distance
+  slices from the Sun, starting at 1 kpc and spaced 2 kpc apart.  The
+  white dashed lines indicate the boundaries of the four quadrants in
+  Galactic longitude.  The white dashed-dotted line shows the region
+  of the sky that \apogee\ can observe (-5$^{\circ} \leq l \leq
+  250^{\circ}$).}\label{fig:detect}
 \end{figure}
@@
   \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
-  distance uncertainties, parameters and shape of the distribution
-  function, shape of the rotation curve, and bar
+  distance uncertainties; parameters and shape of the distribution
+  function; shape of the rotation curve; and bar
   strength.}\label{fig:1dvar}

### 2010-05-21 tex/hercules.tex
groups-simulations: abstract and intro tweaks

@@
 \email{jb2777@nyu.edu}
-%\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
-%  New York University, 4 Washington Place, New York, NY 10003, USA}
-%\altaffiltext{2}{Correspondence should be addressed to jo.bovy@nyu.edu~.}
 
@@
 distribution, is the result of resonant interactions between stars in
-the outer disk and the Galactic bar. Only seen so far in the immediate
-Solar neighborhood, I predict the distribution of stellar velocities
-and the run of the Hercules stream as a function of location in the
-Galactic disk in a simple model for the Galaxy and the bar that
-produces the observed Hercules stream. I further show that the
-Hercules feature is strong enough that it can be unambiguously
-detected in the distribution of line-of-sight velocities in selected
-directions. I quantify the most promising lines-of-sight for detection
-in radial velocities; These are at $l \gtrsim 250^{\circ}$. The
+the outer disk and the Galactic bar. So far it has only been seen in
+the immediate Solar neighborhood. I predict the distribution of
+stellar velocities and the run of the Hercules stream as a function of
+location in the Galactic disk in a simple model for the Galaxy and the
+bar that produces the observed Hercules stream. I further show that
+the Hercules feature is strong enough to be unambiguously detected in
+the distribution of line-of-sight velocities in selected directions. I
+quantify the most promising lines-of-sight for detection in radial
+velocities; these directions are all at $l \gtrsim 250^{\circ}$. The
 predictions presented here are only weakly affected by distance
@@
 stellar disk, and the details of the Galactic potential including the
-effect of spiral structure; \Gaia\ and future spectroscopic surveys of
+effect of spiral structure. \Gaia\ and future spectroscopic surveys of
 the Galactic disk will be able to robustly test the origin of the
@@
 \citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
-that has it originating through resonant interactions of stars in the
-outer disk with the bar in the central region of the Galaxy has been
-put forward \citep{dehnen00a,fux01a}, but any confirmation of this
-picture beyond the locallly observed stellar kinematics remains
-lacking. An investigation of the Hercules members' color--magnitude
-diagram found no evidence for any significant metallicity anomaly with
-respect to other local stars \citep{Bovy10a}, even though such a
-difference should be expected as the Hercules stars originate from a
-few kiloparsec toward the Galactic center where the average
-metallicity is higher than in the Solar neighborhood. The absence of
-the predicted metallicity anomaly could potentially be explained
-through, \eg, radial mixing \citep{sellwood02a}, but it is clear that
-chemodynamical modeling of the Galactic disk is currently not up to
-the task of separating out these effects.
+that has the Hercules stream originating through resonant interactions
+of stars in the outer disk with the bar in the central region of the
+Galaxy has been put forward \citep{dehnen00a,fux01a}, but any
+confirmation of this picture beyond the locally observed stellar
+kinematics remains lacking. An investigation of the Hercules members'
+color--magnitude diagram found no evidence for any significant
+metallicity anomaly with respect to other local stars \citep{Bovy10a},
+even though such a difference should be expected as the Hercules stars
+originate from a few kiloparsec toward the Galactic center where the
+average metallicity is higher than in the Solar neighborhood. The
+absence of the predicted metallicity anomaly could potentially be
+explained, \eg, through radial mixing \citep{sellwood02a}, but it is
+clear that chemodynamical modeling of the Galactic disk is currently
+not up to the task of separating out these effects.
 

### 2010-05-26 tex/hercules.tex
groups-simulations:most of Hogg's comments

@@
+%TO DO:
+% - l-range of good directions
+% - radial and tangential confusion
+% - Hercules in Fig 2.
 \documentclass[12pt,preprint]{aastex}
@@
 the outer disk and the Galactic bar. So far it has only been seen in
-the immediate Solar neighborhood. I predict the distribution of
-stellar velocities and the run of the Hercules stream as a function of
-location in the Galactic disk in a simple model for the Galaxy and the
-bar that produces the observed Hercules stream. I further show that
-the Hercules feature is strong enough to be unambiguously detected in
-the distribution of line-of-sight velocities in selected directions. I
-quantify the most promising lines-of-sight for detection in radial
-velocities; these directions are all at $l \gtrsim 250^{\circ}$. The
-predictions presented here are only weakly affected by distance
-uncertainties, assumptions about the distribution function of the
-stellar disk, and the details of the Galactic potential including the
-effect of spiral structure. \Gaia\ and future spectroscopic surveys of
-the Galactic disk will be able to robustly test the origin of the
-Hercules stream and constrain the properties of the Galactic bar.
+the immediate Solar neighborhood, but any resonance model make a
+prediction over a large fraction of the Galactic disk. I predict the
+distribution of stellar velocities and changing Hercules signature as
+a function of location in the Galactic disk in a simple model for the
+Galaxy and the bar that produces the observed Hercules stream. The
+Hercules feature is expected to be strong enough to be unambiguously
+detected in the distribution of line-of-sight velocities in selected
+directions. I identify quantitatively the most promising lines of
+sight for detection in line-of-sight velocities using the
+Kullback-Leibler divergence between the predictions from the resonance
+model and an axisymmetric model; these directions are all at $l
+\gtrsim 250^{\circ}$. The predictions presented here are only weakly
+affected by distance uncertainties, assumptions about the distribution
+function in the stellar disk, and the details of the Galactic
+potential including the effect of spiral structure. \Gaia\ and future
+spectroscopic surveys of the Galactic disk will be able to robustly
+test the origin of the Hercules stream and constrain the properties of
+the Galactic bar.
 \end{abstract}
@@
 overdensities of co-moving stars that cannot be explained as the
-vestiges of the clumpy nature of star formation
-\citep[\eg,][]{Bovy10a}. In particular, the Hercules moving group, a
-group of stars significantly offset from the bulk of the observed
-velocity distribution \citep[\eg,][see
-  \figurename~\ref{fig:obs}]{Dehnen98b,Bovy10a}, displays a wide range
-of ages and metallicities
-\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
-that has the Hercules stream originating through resonant interactions
-of stars in the outer disk with the bar in the central region of the
-Galaxy has been put forward \citep{dehnen00a,fux01a}, but any
+vestiges of clumpy star formation \citep[\eg,][]{Bovy10a}. In
+particular, the Hercules moving group, a group of stars significantly
+offset from the bulk of the observed velocity distribution
+\citep[\eg,][see \figurename~\ref{fig:obs}]{Dehnen98b,Bovy09a},
+displays a wide range of ages and metallicities
+\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a,Bovy10a}. A compelling
+model has the Hercules stream originating through resonant
+interactions of stars in the outer disk with the bar in the central
+region of the Galaxy \citep{dehnen00a,fux01a}, but there is no
 confirmation of this picture beyond the locally observed stellar
-kinematics remains lacking. An investigation of the Hercules members'
-color--magnitude diagram found no evidence for any significant
-metallicity anomaly with respect to other local stars \citep{Bovy10a},
-even though such a difference should be expected as the Hercules stars
-originate from a few kiloparsec toward the Galactic center where the
-average metallicity is higher than in the Solar neighborhood. The
-absence of the predicted metallicity anomaly could potentially be
-explained, \eg, through radial mixing \citep{sellwood02a}, but it is
-clear that chemodynamical modeling of the Galactic disk is currently
-not up to the task of separating out these effects.
+kinematics that it was proposed to explain. An investigation of the
+Hercules members' color--magnitude diagram found no evidence for any
+significant metallicity anomaly with respect to other local stars
+\citep{Bovy10a}, even though such a difference should be expected as
+the Hercules stars originate from a few kiloparsec toward the Galactic
+center where the average metallicity is higher than in the Solar
+neighborhood. The absence of the predicted metallicity anomaly could
+potentially be explained, \eg, through radial mixing
+\citep{sellwood02a}, but it is clear that chemodynamical modeling of
+the Galactic disk is not currently up to the task of separating out
+these effects.
 
 In this Letter I propose that a robust and unambiguous test of the
-bar-resonance model for the Hercules stream is the distinct pattern it
-predicts as we trace the velocity distribution of stars near the Solar
-radius around the Galaxy. \Gaia\ is in a good position to have the
-final word on this, but I also show that the Hercules feature can be
-detected in the line-of-sight velocity distribution in selected
-directions on the sky. Clear signatures should exist in regions a few
-kiloparsec from the Sun, allowing for a clean verification of the
-bar-resonance model for the Hercules stream.
+bar-resonance model for the Hercules stream is a search for the
+distinct pattern it predicts as we trace the velocity distribution of
+stars near the Solar radius around the Galaxy. \Gaia\ is in a good
+position to have the final word on this, but I also show that the
+Hercules feature can be detected in the line-of-sight velocity
+distribution in selected directions on the sky. Clear signatures are
+predicted to be detectable in regions a few kiloparsec from the Sun,
+allowing for a clean verification of the bar-resonance model for the
+Hercules stream.
 
@@
 To simulate the effect of the bar on stellar orbits near the Solar
-circle, we follow the approach of \citet{dehnen00a}. For a fixed
+circle, we follow the approach of \citet{dehnen00a}. For each fixed
 position in the Galaxy, we evaluate the velocity distribution function
@@
 \end{equation}
-and for most of the simulations below we specify this model further to
-a flat rotation curve ($\beta = 0$). The model for the bar is given by
-the following potential
+where \Ro\ is the distance from the Sun to the Galactic center. For
+most of the simulations below we specify this model further to a flat
+rotation curve ($\beta = 0$). The model for the bar is given by the
+following potential
 \begin{equation}
@@
 radius. The bar is grown smoothly during a time $t_1$, always set to
-half of the total integration time in what follows, following the
+half of the total integration time in what follows, using the
 prescription
@@
 A comparison of the local velocity distribution obtained by using this
-simple bar potential with that resulting from the more realistic bar
-potential of \citet{gardner10a} shows that this simple model is
-sufficient to model the bar's influence on stellar orbits in the outer
-disk.
+simple bar potential with that resulting from the more realistic one
+of \citet{gardner10a} shows that this simple potential is sufficient
+to model the bar's influence on stellar orbits in the outer disk.
 
@@
 Galactocentric azimuths, fixing the bar angle. In this way, these
-simulations become predictions for what one might find when one
-studies the velocity distribution in regions beyond those probed by
-the \hipparcos\ mission.
+simulations can make predictions for what we might find when we study
+the velocity distribution in regions beyond those probed by the
+\hipparcos\ mission.
 
 In \figurename~\ref{fig:rphi2d}, I present the predictions for the
-velocity distributions near the Solar circle for the full
-360$^{\circ}$ range in Galactocentric azimuth for the fiducial model
-from \sectionname~\ref{sec:method}. Only the range $-90^{\circ} \leq
-\phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
-symmetric with respect to rotations through 180$^{\circ}$, such that
-the predictions for the other side of the Galaxy can be obtained from
-those shown by shifting the azimuths by 180$^{\circ}$. The $R = \Ro$
-locus is not shown here, but it can be inferred from \figurename~2 in
-\citet{dehnen00a}.
+two-dimensional velocity distributions near the Solar circle for the
+full 360$^{\circ}$ range in Galactocentric azimuth for the fiducial
+model from \sectionname~\ref{sec:method}. Only the range $-90^{\circ}
+\leq \phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
+symmetric with respect to rotations through 180$^{\circ}$; the
+predictions for the other side of the Galaxy can be obtained from
+those shown by shifting the azimuths by 180$^{\circ}$. The predicted
+velocity distributions on the Solar circle are not shown here, but
+they can be inferred from \figurename~2 in \citet{dehnen00a}.
 
@@
 stream going around the Galaxy and when looking further in toward the
-Galactic center or further out. The effect of the bar is also not
-confined to a narrow range in radii, rather the effects of the bar and
-especially its outer Lindblad resonance are felt through a large range
-of radii (0.6 $\leq R/\Ro \leq 1.4$). These effects are strong and
-robust predictions of the bar-origin model for the Hercules stream,
-especially in regions beyond the Solar radius, where the Hercules
-feature can be found significantly removed from the bulk of the
-velocity distribution, which might be severely shaped at all locations
-by the processes that create the pattern of low-velocity moving groups
-in the local velocity distribution (the main mode in
-\figurename~\ref{fig:obs}). As I discuss below, \Gaia\ will be able to
-verify these predictions.
+Galactic center or further out. Furthermore, the effect of the bar is
+not confined to a narrow range in radii, but rather the effects of the
+bar and especially its outer Lindblad resonance are felt through a
+large range of radii (0.6 $\leq R/\Ro \leq 1.4$). These effects are
+strong and robust predictions of the bar-origin model for the Hercules
+stream. This is especially the case in regions beyond the Solar
+radius. There the Hercules feature can be found significantly removed
+from the bulk of the velocity distribution, which might be severely
+shaped at all locations by the processes that create the pattern of
+low-velocity moving groups in the local velocity distribution (the
+main mode in \figurename~\ref{fig:obs}). As I discuss below, \Gaia\
+will be able to verify these predictions.
 
@@
 measuring proper motions, we can ask whether a targeted survey of
-line-of-sight velocities in selected regions of the Galaxy could see a
-clear signature of the Hercules feature. That this should be possible
-follows from the following argument. The Hercules feature stands out
-in the velocity distribution most in its tangential velocity
-$\vphi$. On the Solar circle near the Sun we see almost entirely this
-Galactocentric tangential velocity in the line-of-sight, so we would
-expect to see at least a hint of the Hercules stream in the line-of
-sight.
+line-of-sight velocities in selected regions of the Galaxy could
+detect a clear signature of the Hercules feature? The Hercules feature
+stands out in the velocity distribution most in its tangential
+velocity $\vphi$. On the Solar circle near the Sun the line-of-sight
+velocity is almost entirely this Galactocentric tangential velocity,
+so we would expect to see at least a hint of the Hercules stream in
+the line-of sight.
 
@@
 around the Galaxy, obtained by integrating the full two-dimensional
-velocity distribution over the tangential velocity. This
-\figurename\ is no longer symmetric with respect to shifts of
-180$^{\circ}$ degree in azimuth, since the line-of-sight direction
-breaks this symmetry. The gray curve is the line-of-sight distribution
-one would observe if the targeted stars were in a steady state in an
+velocity distribution over the tangential velocity. This \figurename\
+is no longer symmetric with respect to shifts of 180$^{\circ}$ degree
+in azimuth, since the line-of-sight direction breaks this
+symmetry. The gray curve is the line-of-sight distribution we would
+observe if the targeted stars were in a steady state in an
 axisymmetric potential with a Dehnen distribution function
@@
 the line-of-sight velocity distributions on the other side of the
-Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). This is due to the
-fact that on that side we mostly see the Galactocentric radial
-velocity, in which the Hercules feature does not separate clearly from
-the bulk of the stars. Even though upcoming surveys such as
-\apogee\ have the ability to probe the other side of the Galaxy, for
-the purposes of detecting the Hercules stream this is uninteresting.
+Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). On that side we
+mostly see the Galactocentric radial velocity in the line-of-sight, in
+which the Hercules feature does not separate clearly from the bulk of
+the stars. Even though upcoming surveys such as \apogee\ have the
+ability to probe the other side of the Galaxy, for the purposes of
+detecting the Hercules stream this is uninteresting.
 
@@
 various nuisance and dynamical parameters on the predicted
-line-of-sight velocity distribution for the most promising
-line-of-sight and distance from the Sun that is more than 3 kpc
-away. We see that the effect of distance uncertainties is
-negligible. The exact shape and parameters of the stellar distribution
-function also only minimally affect the predicted line-of-sight
-velocity distribution. Changing the scale-length within its current
-uncertainty changes the prediction only in the amplitude of the
-effect: The Hercules stream is more prominent for shorter
-scale-lengths since then there are relatively more stars in the inner
-Galaxy that are brought to the Solar circle by the interaction with
-the bar. Using a Shu distribution function \citep{shu69a} instead of
-the Dehnen distribution function also gives very similar
-predictions. Changing the axisymmetric potential by adjusting the
-shape of the rotation curve shifts the peak as expected
+line-of-sight velocity distribution for the most promising region that
+is more than 3 kpc away. We see in this \figurename\ that the effect
+of distance uncertainties is negligible. The exact shape and
+parameters of the stellar distribution function also only minimally
+affect the predicted line-of-sight velocity distribution. Changing the
+scale-length within its current uncertainty changes the prediction
+only in the amplitude of the effect: The Hercules stream is more
+prominent for shorter scale-lengths since then there are relatively
+more stars in the inner Galaxy that are brought to the Solar circle by
+the interaction with the bar. Using a Shu distribution function
+\citep{shu69a} instead of the Dehnen distribution function also gives
+very similar predictions. Changing the axisymmetric potential by
+adjusting the shape of the rotation curve shifts the peak as expected
 \citep[see][]{dehnen00a}, but the Hercules feature nevertheless
@@
 \figurename~\ref{fig:rphi2d}, which will be able to verify the
-predictions laid out in this paper. This will allow the dynamical
+predictions laid out in this paper. This will also allow the dynamical
 properties of the bar to be tightly constrained, something that will
@@
 \figurename~\ref{fig:detect} this limit appears right on the edge of
-where one might hope to detect the Hercules feature. Since \apogee\ is
+where we might hope to detect the Hercules feature. Since \apogee\ is
 expected to start taking data in 2011 and to finish operations before
@@
 \acknowledgements It is a pleasure to thank Mike Blanton and David
-Hogg for helpful discussions and assistance.  Financial support for
+W.~Hogg for helpful discussions and assistance.  Financial support for
 this project was provided by the National Aeronautics and Space
@@
 \caption{Two-dimensional velocity distributions along the Solar
-  circle. The axes of each subpanel are the same as in
+  circle. The $x$-axis in each subplot is the Galactocentric radial
+  velocity and the $y$-axis is the Galactocentric tangential velocity;
+  the axis-ranges of each subplot are the same as in
   \figurename~\ref{fig:obs}. The velocity distributions on the other

### 2010-05-28 py/bar_figures.py
groups-simulations: different distance uncertainties

@@ -265,7 +265,7 @@ def veldist_1d_convolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     HISTORY:
        2010-05-15 - Written - Bovy (NYU)
     """
-    convolves= [0.,0.1,0.2]
+    convolves= [0.,0.2,0.3]
 
     vloslinspace= (-.9,.9,ngrid)
     vloss= sc.linspace(*vloslinspace)
@@ -310,10 +310,12 @@ def veldist_1d_convolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    xlabel=r'$v_{\mathrm{los}} / v_0$')
     plot.bovy_plot(vloss,vlosds[1],ls='-',color='0.75',
                    overplot=True,zorder=2,lw=2.)
-    plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
+    plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.6',
                    overplot=True,zorder=2,lw=2.)
+    #plot.bovy_plot(vloss,vlosds[3],ls='-',color='0.45',
+    #               overplot=True,zorder=2,lw=2.)
     plot.bovy_text(r'$\mathrm{distance\ uncertainties}$',title=True)
-    plot.bovy_text(0.5,.65,r'$\sigma_d = 0$'+'\n'+r'$\sigma_d = 10 \%$'+ '\n'+r'$\sigma_d = 20 \%$')
+    plot.bovy_text(0.5,.65,r'$\sigma_d = 0$'+'\n'+r'$\sigma_d = 20 \%$'+'\n'+r'$\sigma_d = 30 \%$')
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-05-28 py/calc_veldist_1d.py

@@ -222,6 +222,7 @@ def predictVlos(vloslinspace,l=0.,d=1.,t=-4.,distCoord='GC',
         print "Theta: ", theta
         print "alpha: ", alphalos
         print "R: ", R
+        print "d: ", d
     potparams= (Rolr,alpha,phi,chi,t1)
     return marginalizeAngleGrid(vloslinspace,alphalos,R=R,t=t,pot=pot,
                                 beta=beta,potparams=potparams,

### 2010-05-28 py/bar_figures.py
groups-simulations: apogee figure

@@ -541,6 +541,72 @@ def veldist_1d_slope(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     #pyplot.arrow(0.45,3.12,-0.40,-0.03,color='0.',lw=.75,ls='dashed',zorder=4)
     plot.bovy_end_print(plotfilename)
 
+def veldist_1d_apogee(plotfilename,l=250./180.*m.pi,d=0.25,
+                      ngrid=201,saveDir='../bar/apogee/'):
+    """
+    NAME:
+       veldist_1d_apogee
+    PURPOSE:
+       make a plot showing a 1d velocity distribution in an apogee los
+    INPUT:
+       plotfilename - filename for figure
+       l - Galactic longitude
+       d - distance from the Sun
+       ngrid - number of grid-points to calculate the los velocity distribution
+               on
+       saveDir - save pickles here
+    OUTPUT:
+       Figure in plotfilename
+    HISTORY:
+       2010-05-28 - Written - Bovy (NYU)
+    """
+    vloslinspace= (-.9,.9,ngrid)
+    vloss= sc.linspace(*vloslinspace)
+
+    basesavefilename= os.path.join(saveDir,'apogee_')
+    thissavefilename= basesavefilename+'%.2f_%.2f.sav' % (d,l)
+    if os.path.exists(thissavefilename):
+        print "Restoring apogee los-velocity distribution at d,l = %.2f,%.2f" % (d,l)
+        savefile= open(thissavefilename,'r')
+        vlosd= pickle.load(savefile)
+        axivlosd= pickle.load(savefile)
+        savefile.close()
+    else:
+        print "Calculating apogee los-velocity distribution at d,l = %.2f,%.2f" % (d,l)
+        potparams= (0.9,0.01,25.*_degtorad,.8,None)
+        vlosd= predictVlos(vloslinspace,
+                           l=l,
+                           d=d,
+                           distCoord='sun',
+                           pot='bar',beta=0.,
+                           potparams=potparams)
+        vlosd= vlosd/(sc.nansum(vlosd)*(vloss[1]-vloss[0]))
+        potparams= (0.9,0.00,25.*_degtorad,.8,None)
+        axivlosd= predictVlos(vloslinspace,
+                              l=l,
+                              d=d,
+                              t=0.01,
+                              distCoord='sun',
+                              pot='bar',beta=0.,
+                              potparams=potparams)
[…]

### 2010-05-28 tex/hercules.tex
groups-simulations: to do item
groups-simulations: one more to-do item
groups-simulations: add apogee more clearly, some minor edits

@@
-%TO DO:
-% - l-range of good directions
-% - radial and tangential confusion
-% - Hercules in Fig 2.
 \documentclass[12pt,preprint]{aastex}
@@
 Kullback-Leibler divergence between the predictions from the resonance
-model and an axisymmetric model; these directions are all at $l
-\gtrsim 250^{\circ}$. The predictions presented here are only weakly
-affected by distance uncertainties, assumptions about the distribution
-function in the stellar disk, and the details of the Galactic
-potential including the effect of spiral structure. \Gaia\ and future
-spectroscopic surveys of the Galactic disk will be able to robustly
-test the origin of the Hercules stream and constrain the properties of
-the Galactic bar.
+model and an axisymmetric model; these directions are at $250^{\circ}
+\lesssim l \lesssim 290^{\circ}$. The predictions presented here are
+only weakly affected by distance uncertainties, assumptions about the
+distribution function in the stellar disk, and the details of the
+Galactic potential including the effect of spiral structure. \Gaia\
+and future spectroscopic surveys of the Galactic disk such as \apogee\
+will be able to robustly test the origin of the Hercules stream and
+constrain the properties of the Galactic bar.
 \end{abstract}
@@
 around the Galaxy, obtained by integrating the full two-dimensional
-velocity distribution over the tangential velocity. This \figurename\
-is no longer symmetric with respect to shifts of 180$^{\circ}$ degree
-in azimuth, since the line-of-sight direction breaks this
-symmetry. The gray curve is the line-of-sight distribution we would
-observe if the targeted stars were in a steady state in an
+velocity distribution over the velocity perpendicular to the line of
+sight. This \figurename\ is no longer symmetric with respect to shifts
+of 180$^{\circ}$ degree in azimuth, since the line-of-sight direction
+breaks this symmetry. The gray curve is the line-of-sight distribution
+we would observe if the targeted stars were in a steady state in an
 axisymmetric potential with a Dehnen distribution function
@@
 side of the Galaxy. The Hercules feature is especially strong in the
-line-of-sight velocity distribution for Galactic longitudes $l \geq
-250^{\circ}$. This is because of the fortunate coincidence that for
-those lines-of-sight we see velocities in the direction along which
-the Hercules streams separates most from the bulk of the velocity
-distribution (see \figurename~\ref{fig:rphi2d}) and that at these
-azimuths the Hercules stream is at the furthest distance from the rest
-of the velocity distribution. We see hardly any effect of the bar on
-the line-of-sight velocity distributions on the other side of the
-Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). On that side we
-mostly see the Galactocentric radial velocity in the line-of-sight, in
-which the Hercules feature does not separate clearly from the bulk of
-the stars. Even though upcoming surveys such as \apogee\ have the
-ability to probe the other side of the Galaxy, for the purposes of
-detecting the Hercules stream this is uninteresting.
+line-of-sight velocity distribution for Galactic longitudes
+$250^{\circ} \lesssim l \lesssim 290^{\circ}$. This is because of the
+fortunate coincidence that for those lines-of-sight we see velocities
+in the direction along which the Hercules streams separates most from
+the bulk of the velocity distribution (see
+\figurename~\ref{fig:rphi2d}) and that at these azimuths the Hercules
+stream is at the furthest distance from the rest of the velocity
+distribution. We see hardly any effect of the bar on the line-of-sight
+velocity distributions on the other side of the Galaxy ($90^{\circ}
+\leq \phi \leq 270^{\circ}$). On that side we mostly see the
+Galactocentric radial velocity in the line-of-sight, in which the
+Hercules feature does not separate clearly from the bulk of the
+stars. Even though upcoming surveys such as \apogee\ have the ability
+to probe the other side of the Galaxy, for the purposes of detecting
+the Hercules stream this is uninteresting.
 
@@
 \vo$. \figurename~\ref{fig:detect} shows that there are many
-locations, mostly at $l \gtrsim 250^{\circ}$, where the Hercules
-stream is clearly offset from the bulk of the velocity distribution
-and should be unambiguously detectable.
+locations, mostly at $250^{\circ} \lesssim l \lesssim 290^{\circ}$,
+where the Hercules stream is clearly offset from the bulk of the
+velocity distribution and should be unambiguously detectable.
 
@@
 at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
-\figurename~\ref{fig:detect} this limit appears right on the edge of
-where we might hope to detect the Hercules feature. Since \apogee\ is
-expected to start taking data in 2011 and to finish operations before
-the first \Gaia\ data release, \apogee\ could trace the Hercules
-stream around the Galaxy before \Gaia\ does. Since \apogee\ will also
-obtain detailed chemical abundances for its target stars, this would
-also allow for a more detailed analysis of the abundances of stars in
-the Hercules stream \citep[following, \eg,][]{Bensby07a,Bovy10a},
-which could shed more light on radial mixing in the Galactic disk.
+\figurename~\ref{fig:detect} this limit is right on the edge of where
+we might hope to detect the Hercules feature. The hercules bump in the
+direction $l = 250^{\circ}$ is about half as strong as in
+\figurename~\ref{fig:1dvar}, but remains as a clear signature of the
+bar. Since \apogee\ is expected to start taking data in 2011 and to
+finish operations before the first \Gaia\ data release, \apogee\ could
+trace the Hercules stream around the Galaxy before \Gaia\ does. Since
+\apogee\ will also obtain detailed chemical abundances for its target
+stars, this would also allow for a more detailed analysis of the
+abundances of stars in the Hercules stream \citep[following,
+\eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
+mixing in the Galactic disk.
 
@@
 \includegraphics[width=\textwidth]{rphi2d.ps}
-\caption{Two-dimensional velocity distributions along the Solar
-  circle. The $x$-axis in each subplot is the Galactocentric radial
-  velocity and the $y$-axis is the Galactocentric tangential velocity;
-  the axis-ranges of each subplot are the same as in
-  \figurename~\ref{fig:obs}. The velocity distributions on the other
-  side of the Galaxy are obtained by shifting the azimuths by
-  180$^{\circ}$.}\label{fig:rphi2d}
+\caption{Two-dimensional velocity distributions along the Solar circle
+in the bar-resonance model for the Hercules stream. The $x$-axis in
+each subplot is the Galactocentric radial velocity and the $y$-axis is
+the Galactocentric tangential velocity; the axis-ranges of each
+subplot are the same as in \figurename~\ref{fig:obs}. The velocity
+distributions on the other side of the Galaxy are obtained by shifting
+the azimuths by 180$^{\circ}$.}\label{fig:rphi2d}
 \end{figure}

### 2010-05-28 py/bar_figures.py
groups-simulations: get rid of some commented-out code

@@ -377,14 +377,6 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{bar\ strength}$',title=True)
     plot.bovy_text(0.5,.75,r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$')
-    #pyplot.arrow(0.48,1.39,-0.14,-0.28,color='0.',lw=.75,zorder=4)
-    #pyplot.arrow(0.48,1.22,-0.14,-0.23,color='0.',lw=.75,zorder=4)
-    #pyplot.arrow(0.48,1.04,-0.14,-0.17,color='0.',lw=.75,zorder=4)
-
-    #plot.bovy_text(r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$',top_right=True)
-    #pyplot.arrow(0.45,3.78,-0.41,-0.09,color='0.',lw=.75,ls='dashed',zorder=4)
-    #pyplot.arrow(0.45,3.60,-0.4,-0.13,color='0.',lw=.75,ls='dashed',zorder=4)
-    #pyplot.arrow(0.45,3.4,-0.4,-0.22,color='0.',lw=.75,ls='dashed',zorder=4)
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
@@ -523,22 +515,11 @@ def veldist_1d_slope(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_plot(vloss,vlosds[4],ls='-',color='0.3',
                    overplot=True,zorder=2,lw=1.5)
-    #BOVY: annotate
     plot.bovy_text(r'$\mathrm{shape\ of\ the\ rotation\ curve}$',title=True)
     plot.bovy_text(0.5,.5,r'$\beta = -0.2$'+'\n'+r'$\beta = -0.1$'+ '\n'+
                    r'$\beta = \phantom{-}0.0$'+'\n'+
                    r'$\beta= \phantom{-}0.1$'+'\n'+
                    r'$\beta= \phantom{-}0.2$')
[…]

### 2010-05-28 tex/hercules.tex
groups-simulations: arXiv reference for Shen paper
groups-simulations: small tweaks

@@
 the outer disk and the Galactic bar. So far it has only been seen in
-the immediate Solar neighborhood, but any resonance model make a
+the immediate Solar neighborhood, but the resonance model makes a
 prediction over a large fraction of the Galactic disk. I predict the
-distribution of stellar velocities and changing Hercules signature as
-a function of location in the Galactic disk in a simple model for the
-Galaxy and the bar that produces the observed Hercules stream. The
-Hercules feature is expected to be strong enough to be unambiguously
-detected in the distribution of line-of-sight velocities in selected
-directions. I identify quantitatively the most promising lines of
-sight for detection in line-of-sight velocities using the
-Kullback-Leibler divergence between the predictions from the resonance
-model and an axisymmetric model; these directions are at $250^{\circ}
-\lesssim l \lesssim 290^{\circ}$. The predictions presented here are
-only weakly affected by distance uncertainties, assumptions about the
-distribution function in the stellar disk, and the details of the
-Galactic potential including the effect of spiral structure. \Gaia\
-and future spectroscopic surveys of the Galactic disk such as \apogee\
-will be able to robustly test the origin of the Hercules stream and
-constrain the properties of the Galactic bar.
+distribution of stellar velocities and the changing Hercules feature
+in this distribution as a function of location in the Galactic disk in
+a simple model for the Galaxy and the bar that produces the observed
+Hercules stream. The Hercules feature is expected to be strong enough
+to be unambiguously detected in the distribution of line-of-sight
+velocities in selected directions. I identify quantitatively the most
+promising lines of sight for detection in line-of-sight velocities
+using the Kullback-Leibler divergence between the predictions of the
+resonance model and an axisymmetric model; these directions are at
+$250^{\circ} \lesssim l \lesssim 290^{\circ}$. The predictions
+presented here are only weakly affected by distance uncertainties,
+assumptions about the distribution function in the stellar disk, and
+the details of the Galactic potential including the effect of spiral
+structure. \Gaia\ and future spectroscopic surveys of the Galactic
+disk such as \apogee\ will be able to robustly test the origin of the
+Hercules stream and constrain the properties of the Galactic bar.
 \end{abstract}
@@
 stands out in the velocity distribution most in its tangential
-velocity $\vphi$. On the Solar circle near the Sun the line-of-sight
-velocity is almost entirely this Galactocentric tangential velocity,
-so we would expect to see at least a hint of the Hercules stream in
-the line-of sight.
+velocity. On the Solar circle near the Sun the line-of-sight velocity
+is almost entirely this Galactocentric tangential velocity, so we
+would expect to see at least a hint of the Hercules stream in the
+line-of sight.
 
@@
 around the Galaxy, obtained by integrating the full two-dimensional
-velocity distribution over the velocity perpendicular to the line of
-sight. This \figurename\ is no longer symmetric with respect to shifts
-of 180$^{\circ}$ degree in azimuth, since the line-of-sight direction
-breaks this symmetry. The gray curve is the line-of-sight distribution
-we would observe if the targeted stars were in a steady state in an
-axisymmetric potential with a Dehnen distribution function
+velocity distribution over the velocity component perpendicular to the
+line of sight. This \figurename\ is no longer symmetric with respect
+to shifts of 180$^{\circ}$ degree in azimuth, since the line-of-sight
+direction breaks this symmetry. The gray curve is the line-of-sight
+distribution we would observe if the targeted stars were in a steady
+state in an axisymmetric potential with a Dehnen distribution function
 (\eqnname~[\ref{eq:fdehnen}]). Both line-of-sight velocity
 distributions are normalized to have equal mass since I assume that
-our knowledge of the Galaxy and the tracer population will not be
-great enough that we can test predicted differences in the absolute
-number of stars at a given location between the axisymmetric and the
+our knowledge of the Galaxy and the tracer population will not be good
+enough that we can test predicted differences in the absolute number
+of stars at a given location between the axisymmetric and the
 bar-resonance predictions. These differences are also largely a
@@
 \bibitem[Shen \etal(2010)]{Shen10a}
-  Shen,~J., \etal, 2010, \apj L, submitted
+  Shen,~J., \etal, 2010, \apj L, submitted, arXiv:1005.0385v1 [astro-ph.CO]
 

### 2010-06-01 tex/hercules.tex
groups-simulations: not a Letter anymore

@@
 
-In this Letter I propose that a robust and unambiguous test of the
-bar-resonance model for the Hercules stream is a search for the
-distinct pattern it predicts as we trace the velocity distribution of
-stars near the Solar radius around the Galaxy. \Gaia\ is in a good
-position to have the final word on this, but I also show that the
-Hercules feature can be detected in the line-of-sight velocity
-distribution in selected directions on the sky. Clear signatures are
-predicted to be detectable in regions a few kiloparsec from the Sun,
-allowing for a clean verification of the bar-resonance model for the
-Hercules stream.
+Here I propose that a robust and unambiguous test of the bar-resonance
+model for the Hercules stream is a search for the distinct pattern it
+predicts as we trace the velocity distribution of stars near the Solar
+radius around the Galaxy. \Gaia\ is in a good position to have the
+final word on this, but I also show that the Hercules feature can be
+detected in the line-of-sight velocity distribution in selected
+directions on the sky. Clear signatures are predicted to be detectable
+in regions a few kiloparsec from the Sun, allowing for a clean
+verification of the bar-resonance model for the Hercules stream.
 

### 2010-09-09 tex/response.txt
hercules: first round of response

Dear Editor,

Please find below the response to the referees' comments for the
submission "Tracing the Hercules Stream around the Galaxy" by myself
(MS# ApJ80617).

I have made changes following the referee's suggestions ...

A detailed list of the main changes is:

Please contact me by the information given below with any questions
regarding this re-submission.

With kind regards,

Jo Bovy

-
Jo Bovy
Ctr. for Cosmology & Particle Physics, 
Dept. of Physics
New York University
4 Washington Place
New York, NY 10003
jb2777@nyu.edu

In what follows the referee's comments are indented.

==================================================================

	Referee Report
	 Reviewer's Comments:

	 The paper titled Tracing the Hercules stream around the Galaxy by Jo
	 Bovy is concerned with the origin and detectability of this
	 elusive stream. Overall the paper is well written and
	 clear. I have a few major concerns and some minor concerns
	 that should be addressed before the paper can be accepted for
	 publication.

	 Major concerns:

	 1. Sect 4, page 6, last paragraph: The author discusses in
	 some detail, and shows plots to illustrate, the possible
	 implications of e.g.  errors in the distance determination
	 etc. The one error that is not address explicitly in the text
	 or plots is the error on the measured radial velocities - how
	 well do we need to determine these velocities in order to be
	 able to unambiguously detect the Hercules stream? 1, 5, or 10
	 km/s? Or some other number? And what happens when V0 is
	 absolutely known or if it also has an error? Perhaps the
	 result is that the errors do no matter, but in that case this
	 should be shown and arguments to why it does not matter
	 should be given. Also, will either the RVS on board Gaia or
	 APOGEE deliver radial velocities with small enough errors to
	 detect the Hercules stream?

	 2. A brief discussion about which tracers, e.g. dwarf stars,
	 red clump, red giant stars, to be used to test the model
	 should also be included. For example, is it necessary to
	 reach as far away from the solar neighbourhood as possible
	 (eg giants), or is the major test to get as high volume
	 density of tracers as possible (eg cool dwarfs). Such a
	 discussion would be highly useful for anyone wanting to
	 construct surveys for the very purpose of detecting
	 Hercules. This discussion does not need to be long.

	 3. I find Discussion in Sect 5 (for both Gaia and APOGEE)
	 somewhat sketchy. For example, no mention of the fact that
	 Gaia will provide radial velocities for stars down to roughly
	 17 magnitude is mentioned.  Should the reader conclude that
	 Hercules could be detected with the tangential velocities
	 alone? In that case, how does that work with the tests
	 presented in this article? Again, here the intrinsic
	 brightness of the tracers is of some interest.

	 4. The discussion could in general be deeper, and also connect to
	 the abstract (see below).

	 Additional concerns:

	 1. On page 2, last paragraph of the introduction it is stated
	 that "propose and ... test of the bar resonance model for the
	 Hercules stream". However, this statement is not really
	 carried through in the text. It might be a good idea to
	 connect back to this statement in e.g. Sect 5 and tell the
	 reader how "testable" this model is once all the
	 uncertainties of data and shortcomings of the missions has
	 been discussed. The test and properties of the bar are also
	 mentioned in the abstract, however, there is no discussion in
	 the paper of how the properties of the bar are constrained by
	 the proposed observational test. Add this discussion, which
	 would be very interesting.

	 2. Page 2, statement at the end of the first paragraph "..up
	 to the task of separating out these effects." This statement
	 should be supported with a few relevant references.

	 3. On page 5, sect 4, third paragraph, last sentence. I do
	 not understand what the last sentence means - could the
	 author please re-phrase it? The sentence somehow, does not
	 connect to the previous text.

	 Minor items:

	 1. Page 2, sect 2, end of the last sentence. "evaluated at
	  the initial orbit".  This is, for a person with English as a
	  second language, really hard to understand. It is not the
	  words that are hard but how it follows from the first part
	  of the sentence. For better clarity to all readers I suggest
	  that the author re-writes the last sentence, perhaps into
	  two.

This statement was just supposed to state the collisionless Boltzmann
equation, but I wrote "initial orbit" rather than "initial position
and velocity", since the phases or angles of the initial orbit are
irrelevant. This is because the initial distribution function is
time-independent, and thus only a function of the integrals of the
motion. I changed the wording to "initial position and velocity" but
added a comment about the time-indepence of the DF.

	  2. Page 4, sect 3, first paragraph. Hipparcos mission (at
	   the end of the first paragraph) should be
	   referenced. Perryman et al. 1997, A&A, 323, L49, and/or the
	   catalogue references.

I have added the catalog reference.

	   3. Page 5, sect 4, end of third paragraph. The last
	    sentence is unclear.  APOGEE is able to probe "the other
	    side of the Galaxy" with reference to which side of the
	    Galaxy should the "other" be taken?  Other to 90 - 270
	    degrees? A more explicit statement would be helpful to the
	    reader.

The "other" side of the Galaxy is defined two sentences before this
sentence as the region 90 < \phi < 270, where \phi is the
Galactocentric azimuth. I don't think it is necessary to repeat the
definition two sentences later.

	    4. Page 7, sect 5.1, first sentence, Gaia should have a
	    reference. A relevant reference is Perryman, M. A. C., et
	    al., A&A 369, 339 (2001).

That's quite right. I have added this reference.

	    5. Page 8, sect 5.2, first paragraph
	      Line 6 from top: hercules --> Hercules
	      Line 9 from top: APOGEE could trace --> APOGEE could 
	         potentially trace
	      	 (his is more in line with the other comments in the article
		 that APOGEE is jut on the edge of the suitable region etc)
	      Line 4 from the bottom: 
	         chemical abundances --> elemental abundances

All fixed!

### 2010-09-09 tex/hercules.tex
groups-simulations: Gardner reference update
groups-simulations: groups paper reference
hercules: first round of response

@@
 distribution function before bar formation evaluated at the initial
-orbit.
+position and velocity---or, equivalently, at the initial orbit since
+the initial distribution function is time-independent.
 
@@
 the velocity distribution in regions beyond those probed by the
-\hipparcos\ mission.
+\hipparcos\ mission \citep{ESA97a}.
 
@@
 The astrometric \Gaia\ mission will measure the parallaxes and proper
-motions of up to one billion stars, most of which will be disk
-stars. With parallax and proper motion accuracies down to 10
-microarcseconds, \Gaia\ will be able to probe the kinematics of the
-disk out to several kpc in all directions
+motions of up to one billion stars, most of which will be disk stars
+\citep{Perryman01a}. With parallax and proper motion accuracies down
+to 10 microarcseconds, \Gaia\ will be able to probe the kinematics of
+the disk out to several kpc in all directions
 \citep{bailerjones08a}. Crucially, it will provide large samples of
@@
 \figurename~\ref{fig:detect} this limit is right on the edge of where
-we might hope to detect the Hercules feature. The hercules bump in the
+we might hope to detect the Hercules feature. The Hercules bump in the
 direction $l = 250^{\circ}$ is about half as strong as in
@@
 finish operations before the first \Gaia\ data release, \apogee\ could
-trace the Hercules stream around the Galaxy before \Gaia\ does. Since
-\apogee\ will also obtain detailed chemical abundances for its target
-stars, this would also allow for a more detailed analysis of the
-abundances of stars in the Hercules stream \citep[following,
-\eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
-mixing in the Galactic disk.
+potentially trace the Hercules stream around the Galaxy before
+\Gaia\ does. Since \apogee\ will also obtain detailed elemental
+abundances for its target stars, this would also allow for a more
+detailed analysis of the abundances of stars in the Hercules stream
+\citep[following, \eg,][]{Bensby07a,Bovy10a}, which could shed more
+light on radial mixing in the Galactic disk.
 
-\acknowledgements It is a pleasure to thank Mike Blanton and David
-W.~Hogg for helpful discussions and assistance.  Financial support for
-this project was provided by the National Aeronautics and Space
-Administration (grant NNX08AJ48G) and the National Science Foundation
-(grant AST-0908357).
+\acknowledgements It is a pleasure to thank the anonymous referee,
+Mike Blanton, and David W.~Hogg for helpful discussions and
+assistance.  Financial support for this project was provided by the
+National Aeronautics and Space Administration (grant NNX08AJ48G) and
+the National Science Foundation (grant AST-0908357).
 
@@
 \bibitem[Bovy \& Hogg(2010)]{Bovy10a} Bovy,~J.~\& Hogg,~D.~W., 2010,
-  \apj, in press
+  \apj, 717, 617
 
@@
 
+\bibitem[{{ESA}(1997)}]{ESA97a}
+{ESA}, 1997, {The \emph{Hipparcos} and Tycho Catalogues} (ESA SP-1200; Noordwijk: ESA)
+
 \bibitem[Fux(2001)]{fux01a}
@@
   Gardner,~E.~\& Flynn,~C., 2010,
-  \mnras, in press
+  \mnras, 405, 545
 
@@
   \aap, 453, 635
+
+\bibitem[Perryman \etal(2001)]{Perryman01a}
+  Perryman,~M.~A.~C., \etal, 2001,
+  \aap, 369, 339
 

### 2010-09-09 tex/response.txt
hercules: one more comment

@@
 
+I thought this followed clearly from the sentences before it, but I
+have changed this sentence to
+
+Even though upcoming surveys such as APOGEE have the ability to probe
+the other side of the Galaxy, for the purposes of detecting the
+Hercules stream this is uninteresting as it does not show up in the
+line-of-sight velocity distribution there.
+
+to make the conclusion clearer.
+
 	 Minor items:

### 2010-09-09 tex/hercules.tex
hercules: one more comment

@@
 to probe the other side of the Galaxy, for the purposes of detecting
-the Hercules stream this is uninteresting.
+the Hercules stream this is uninteresting as it does not show up in
+the line-of-sight velocity distribution there.
 

### 2010-09-11 py/bar_figures.py
groups-simulations: vlos uncertainties

@@ -6,6 +6,7 @@ import sys
 import cPickle as pickle
 import math as m
 import scipy as sc
+from scipy import signal
 from optparse import OptionParser
 import bovy_plot as plot
 from matplotlib import pyplot
@@ -318,6 +319,60 @@ def veldist_1d_convolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_text(0.5,.65,r'$\sigma_d = 0$'+'\n'+r'$\sigma_d = 20 \%$'+'\n'+r'$\sigma_d = 30 \%$')
     plot.bovy_end_print(plotfilename)
 
+def veldist_1d_vrconvolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
+                        ngrid=201,saveDir='../bar/1dvar/'):
+    """
+    NAME:
+       veldist_1d_vrconvolve
+    PURPOSE:
+       make a plot showing the influence of the los velocity uncertainties
+    INPUT:
+       plotfilename - filename for figure
+       phi - Galactocentric azimuth
+       R - Galactocentric radius
+       ngrid - number of grid-points to calculate the los velocity distribution
+               on
+       saveDir - save pickles here
+    OUTPUT:
+       Figure in plotfilename
+    HISTORY:
+       2010-09-11 - Written - Bovy (NYU)
+    """
+    convolves= [0.,0.02,0.04,0.08]#0, 5, 10, 20 km/s
+
+    vloslinspace= (-.9,.9,ngrid)
+    vloss= sc.linspace(*vloslinspace)
+
+    vlosds= []
+    basesavefilename= os.path.join(saveDir,'vrconvolve_')
+    for distsig in convolves:
+        thissavefilename= os.path.join(saveDir,'convolve_')+'%.1f.sav' % 0.
+        print "Restoring los-velocity distribution at distance uncertainties %.1f" % 0.
+        savefile= open(thissavefilename,'r')
+        vlosd= pickle.load(savefile)
+        savefile.close()
+        #Create Gaussian
+        gauss= sc.exp(-0.5*vloss**2./distsig)
+        gauss= gauss/sc.sum(gauss)/(vloss[1]-vloss[0])
+        vlosd= signal.convolve(vlosd,gauss,mode='same')
+        vlosds.append(vlosd)
+    #Plot
+    plot.bovy_print()
+    plot.bovy_plot(vloss,vlosds[0],'k-',zorder=3,
+                   xrange=[vloslinspace[0],vloslinspace[1]],
[…]

### 2010-09-11 py/bar_figures.py
groups-simulations: vlos uncertainties

@@ -338,12 +338,18 @@ def veldist_1d_vrconvolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     HISTORY:
        2010-09-11 - Written - Bovy (NYU)
     """
-    convolves= [0.,0.02,0.04,0.08]#0, 5, 10, 20 km/s
+    convolves= [0.02,0.04,0.08]#0, 5, 10, 20 km/s
 
     vloslinspace= (-.9,.9,ngrid)
     vloss= sc.linspace(*vloslinspace)
 
     vlosds= []
+    thissavefilename= os.path.join(saveDir,'convolve_')+'%.1f.sav' % 0.
+    print "Restoring los-velocity distribution at distance uncertainties %.1f" % 0.
+    savefile= open(thissavefilename,'r')
+    vlosd= pickle.load(savefile)
+    savefile.close()
+    vlosds.append(vlosd)
     basesavefilename= os.path.join(saveDir,'vrconvolve_')
     for distsig in convolves:
         thissavefilename= os.path.join(saveDir,'convolve_')+'%.1f.sav' % 0.
@@ -352,9 +358,10 @@ def veldist_1d_vrconvolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
         vlosd= pickle.load(savefile)
         savefile.close()
         #Create Gaussian
-        gauss= sc.exp(-0.5*vloss**2./distsig)
-        gauss= gauss/sc.sum(gauss)/(vloss[1]-vloss[0])
+        gauss= sc.exp(-0.5*vloss**2./distsig**2.)
+        #gauss= gauss/sc.sum(gauss)/(vloss[1]-vloss[0])
         vlosd= signal.convolve(vlosd,gauss,mode='same')
+        vlosd= vlosd/sc.sum(vlosd)/(vloss[1]-vloss[0])
         vlosds.append(vlosd)
     #Plot
     plot.bovy_print()
@@ -368,9 +375,9 @@ def veldist_1d_vrconvolve(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    overplot=True,zorder=2,lw=2.)
     plot.bovy_plot(vloss,vlosds[3],ls='-',color='0.45',
                    overplot=True,zorder=2,lw=2.)
[…]

### 2010-09-11 tex/hercules.tex
groups-simulations: add vlos uncertainty figure

@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{distuncertain.ps}
-\includegraphics[width=0.5\textwidth]{df.ps}\\
-\includegraphics[width=0.5\textwidth]{slope.ps}
-\includegraphics[width=0.5\textwidth]{barstrength.ps}
+\includegraphics[width=0.35\textwidth]{distuncertain.ps}
+\includegraphics[width=0.35\textwidth]{vlosuncertain.ps}\\
+\includegraphics[width=0.35\textwidth]{df.ps}
+\includegraphics[width=0.35\textwidth]{slope.ps}\\
+\includegraphics[width=0.35\textwidth]{barstrength.ps}
 \caption{Variation of the predicted velocity distribution at R = 1.075
   \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
-  distance uncertainties; parameters and shape of the distribution
-  function; shape of the rotation curve; and bar
-  strength.}\label{fig:1dvar}
+  distance uncertainties; line-of-sight velocity uncertainties;
+  parameters and shape of the distribution function; shape of the
+  rotation curve; and bar strength.}\label{fig:1dvar}
 \end{figure}

### 2010-09-12 py/bar_figures.py
groups-simulations: Rolr figure

@@ -441,6 +441,67 @@ def veldist_1d_barstrength(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_text(0.5,.75,r'$\alpha = 0.013$'+'\n'+r'$\alpha = 0.010$'+ '\n'+r'$\alpha = 0.007$')
     plot.bovy_end_print(plotfilename)
 
+def veldist_1d_rolr(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
+                    ngrid=201,saveDir='../bar/1dvar/'):
+    """
+    NAME:
+       veldist_1d_rolr
+    PURPOSE:
+       make a plot showing the influence of the bar R_OLR
+    INPUT:
+       plotfilename - filename for figure
+       phi - Galactocentric azimuth
+       R - Galactocentric radius
+       ngrid - number of grid-points to calculate the los velocity distribution
+               on
+       saveDir - save pickles here
+    OUTPUT:
+       Figure in plotfilename
+    HISTORY:
+       2010-09-11 - Written - Bovy (NYU)
+    """
+    rolrs= [0.8,0.9,1.0]
+
+    vloslinspace= (-.9,.9,ngrid)
+    vloss= sc.linspace(*vloslinspace)
+
+    vlosds= []
+    basesavefilename= os.path.join(saveDir,'rolr_')
+    for rolr in rolrs:
+        thissavefilename= basesavefilename+'%.3f.sav' % rolr
+        if os.path.exists(thissavefilename):
+            print "Restoring los-velocity distribution at R_OLR %.1f" % rolr
+            savefile= open(thissavefilename,'r')
+            vlosd= pickle.load(savefile)
+            savefile.close()
+        else:
+            print "Calculating los-velocity distribution at R_OLR %.1f" % rolr
+            potparams= (rolr,0.01,25.*_degtorad,.8,None)
+            vlosd= predictVlos(vloslinspace,
+                               l=phi,
+                               d=R,
+                               distCoord='GCGC',
+                               pot='bar',beta=0.,
+                               potparams=potparams)
+            vlosd= vlosd/(sc.nansum(vlosd)*(vloss[1]-vloss[0]))
+            savefile= open(thissavefilename,'w')
+            pickle.dump(vlosd,savefile)
+            savefile.close()
+        vlosds.append(vlosd)
+    #Plot
+    plot.bovy_print()
[…]

### 2010-09-12 py/bar_figures.py
groups-simulations: bar olr figure

@@ -498,7 +498,7 @@ def veldist_1d_rolr(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
                    overplot=True,zorder=2,lw=2.)
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
                    overplot=True,zorder=2,lw=1.5)
-    plot.bovy_text(r'$\mathrm{bar\ strength}$',title=True)
+    plot.bovy_text(r'$\mathrm{bar}\ R_{\mathrm{OLR}}$',title=True)
     plot.bovy_text(0.5,.75,r'$R_{\mathrm{OLR}} = 1.0$'+'\n'+r'$R_{\mathrm{OLR}} = 0.9$'+ '\n'+r'$R_{\mathrm{OLR}} = 0.8$')
     plot.bovy_end_print(plotfilename)
 

### 2010-09-12 py/bar_figures.py
groups-simulations: different Rolr values

@@ -460,7 +460,7 @@ def veldist_1d_rolr(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     HISTORY:
        2010-09-11 - Written - Bovy (NYU)
     """
-    rolrs= [0.8,0.9,1.0]
+    rolrs= [0.85,0.9,0.95]
 
     vloslinspace= (-.9,.9,ngrid)
     vloss= sc.linspace(*vloslinspace)
@@ -499,7 +499,7 @@ def veldist_1d_rolr(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{bar}\ R_{\mathrm{OLR}}$',title=True)
-    plot.bovy_text(0.5,.75,r'$R_{\mathrm{OLR}} = 1.0$'+'\n'+r'$R_{\mathrm{OLR}} = 0.9$'+ '\n'+r'$R_{\mathrm{OLR}} = 0.8$')
+    plot.bovy_text(0.5,.75,r'$R_{\mathrm{OLR}} = 0.95$'+'\n'+r'$R_{\mathrm{OLR}} = 0.90$'+ '\n'+r'$R_{\mathrm{OLR}} = 0.85$')
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-09-12 py/bar_figures.py
groups-simulations: tweak

@@ -499,7 +499,7 @@ def veldist_1d_rolr(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{bar}\ R_{\mathrm{OLR}}$',title=True)
-    plot.bovy_text(0.5,.75,r'$R_{\mathrm{OLR}} = 0.95$'+'\n'+r'$R_{\mathrm{OLR}} = 0.90$'+ '\n'+r'$R_{\mathrm{OLR}} = 0.85$')
+    plot.bovy_text(0.4,.75,r'$R_{\mathrm{OLR}} = 0.95\ R_0$'+'\n'+r'$R_{\mathrm{OLR}} = 0.90\ R_0$'+ '\n'+r'$R_{\mathrm{OLR}} = 0.85\ R_0$')
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-09-12 py/bar_figures.py
groups-simulations: info tweak

@@ -470,12 +470,12 @@ def veldist_1d_rolr(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     for rolr in rolrs:
         thissavefilename= basesavefilename+'%.3f.sav' % rolr
         if os.path.exists(thissavefilename):
-            print "Restoring los-velocity distribution at R_OLR %.1f" % rolr
+            print "Restoring los-velocity distribution at R_OLR %.2f" % rolr
             savefile= open(thissavefilename,'r')
             vlosd= pickle.load(savefile)
             savefile.close()
         else:
-            print "Calculating los-velocity distribution at R_OLR %.1f" % rolr
+            print "Calculating los-velocity distribution at R_OLR %.2f" % rolr
             potparams= (rolr,0.01,25.*_degtorad,.8,None)
             vlosd= predictVlos(vloslinspace,
                                l=phi,

### 2010-09-12 tex/response.txt
hercules: some comments-to-self in the response
groups-simulations: further Gaia discussion

@@
 
+Gaia errors: see below; APOGEE errors: better than 0.5 km/s.
+
 	 2. A brief discussion about which tracers, e.g. dwarf stars,
@@
 
+> 3 kpc: giants (dwarfs in GAIA few kpc, giants go far enough and
+  bright enough to have RVs)
+
 	 3. I find Discussion in Sect 5 (for both Gaia and APOGEE)
@@
 	 brightness of the tracers is of some interest.
+
+I added a reference to the line-of-sight velocity accuracies expected
+for Gaia, and an explicit comment that the velocities of giants will
+be measured accurately enough to verify the predictions of this paper.
 

### 2010-09-12 tex/hercules.tex
groups-simulations: add HERMES reference
groups-simulations: bar olr figure
groups-simulations: further Gaia discussion

@@
 \citep{Perryman01a}. With parallax and proper motion accuracies down
-to 10 microarcseconds, \Gaia\ will be able to probe the kinematics of
-the disk out to several kpc in all directions
-\citep{bailerjones08a}. Crucially, it will provide large samples of
-stars in many of the promising regions in
-\figurename~\ref{fig:rphi2d}, which will be able to verify the
-predictions laid out in this paper. This will also allow the dynamical
-properties of the bar to be tightly constrained, something that will
-be hard for \Gaia\ to do using direct observations of the bulge due to
-the large extinction toward the Galactic center \citep{robin05a}.
+to 10 microarcseconds and mission-averaged line-of-sight velocity
+uncertainties smaller than 10 km s$^{-1}$ for many stars to 17th
+magnitude, \Gaia\ will be able to probe the kinematics of the disk out
+to several kpc in all directions
+\citep{Katz04a,bailerjones08a}. Crucially, it will provide large
+samples of giants in many of the promising regions in
+\figurename~\ref{fig:rphi2d}, with velocity measurements accurate
+enough to verify the predictions laid out in this paper. This will
+also allow the dynamical properties of the bar to be tightly
+constrained, something that will be hard for \Gaia\ to do using direct
+observations of the bulge due to the large extinction toward the
+Galactic center \citep{robin05a}.
 
@@
 
-\acknowledgements It is a pleasure to thank the anonymous referee,
-Mike Blanton, and David W.~Hogg for helpful discussions and
-assistance.  Financial support for this project was provided by the
-National Aeronautics and Space Administration (grant NNX08AJ48G) and
-the National Science Foundation (grant AST-0908357).
+\citep{Freeman10a}.
+
+\acknowledgements It is a pleasure to thank the anonymous referee for
+valuable comments and Mike Blanton and David W.~Hogg for helpful
+discussions and assistance.  Financial support for this project was
+provided by the National Aeronautics and Space Administration (grant
+NNX08AJ48G) and the National Science Foundation (grant AST-0908357).
 
@@
 
+\bibitem[Freeman \etal(2010)]{Freeman10a}
+  Freeman,~K., Bland-Hawthorn,~J., \& Barden,~S., 2010, 
+  AAO Newsletter, 117, 9
+
 \bibitem[Fux(2001)]{fux01a}
@@
   \mnras, 405, 545
+
+\bibitem[Katz \etal(2004)]{Katz04a}
+Katz,~D., \etal, 2004,
+\mnras, 354, 1223
 
@@
 \includegraphics[width=0.35\textwidth]{slope.ps}\\
+\includegraphics[width=0.35\textwidth]{rolr.ps}
 \includegraphics[width=0.35\textwidth]{barstrength.ps}

### 2010-09-12 py/bar_figures.py
groups-simulations: rolr figure tweak

@@ -499,7 +499,7 @@ def veldist_1d_rolr(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,
     plot.bovy_plot(vloss,vlosds[2],ls='-',color='0.5',
                    overplot=True,zorder=2,lw=1.5)
     plot.bovy_text(r'$\mathrm{bar}\ R_{\mathrm{OLR}}$',title=True)
-    plot.bovy_text(0.4,.75,r'$R_{\mathrm{OLR}} = 0.95\ R_0$'+'\n'+r'$R_{\mathrm{OLR}} = 0.90\ R_0$'+ '\n'+r'$R_{\mathrm{OLR}} = 0.85\ R_0$')
+    plot.bovy_text(0.36,.75,r'$R_{\mathrm{OLR}} = 0.95\ R_0$'+'\n'+r'$R_{\mathrm{OLR}} = 0.90\ R_0$'+ '\n'+r'$R_{\mathrm{OLR}} = 0.85\ R_0$')
     plot.bovy_end_print(plotfilename)
 
 def veldist_1d_df(plotfilename,phi=_DEFAULTPHI,R=_DEFAULTR,

### 2010-09-12 py/sample_vd.py
groups-simulations: sample from a velocity distribution

@@ -1,0 +1,28 @@
+import numpy as nu
+from scipy import stats
+from scipy import interpolate
+
+def sample_vd(vloslinspace,vd,nsample=1):
+    """
+    NAME:
+       sample_vd
+    PURPOSE:
+       sample points from a velocity distribution
+    INPUT:
+       vloslinspace - argument for linspace
+       vd - velocity distribution over this range
+       nsample - number of samples
+    OUTPUT:
+       list of samples
+    HISTORY:
+       2010-09-12 - Written - Bovy (NYU)
+    """
+    vloss= nu.linspace(*vloslinspace)
+    vdmax= nu.amax(vd)
+    vd= interpolate.InterpolatedUnivariateSpline(vloss,vd)
+    out= []
+    while len(out) < nsample:
+        prop= stats.uniform.rvs()*(vloss[-1]-vloss[0])+vloss[0]
+        if stats.uniform.rvs()*vdmax < vd(prop):
+            out.append(prop)
+    return out

### 2010-09-12 tex/response.txt
hercules: add general discussion

@@
 
-Gaia errors: see below; APOGEE errors: better than 0.5 km/s.
+I added a panel to Figure 5 showing the influence of realistic
+line-of-sight velocity uncertainties. The expected uncertainties for
+APOGEE and HERMES are below 1 km/s; for such small uncertainties the
+uncertainty-free prediction and the uncertainty-convolved prediction
+are indistinguishable.
+
+I added explicit references to the Gaia, APOGEE, and HERMES
+uncertainties in the sections describing these surveys.
+
+BOVY: V0
 
@@
 
+I hope that by adding the general discussion and the greater detail in
+discussing the specific surveys that could detect the Hercules
+feature, the discussion now has the required depth.
+
 	 Additional concerns:
@@
 I thought this followed clearly from the sentences before it, but I
-have changed this sentence to
+have changed this sentence to:
 

### 2010-09-12 tex/hercules.tex
groups-simulations: note about APOGEE
hercules: add general discussion

@@
 \newcommand{\apogee}{APOGEE}
+\newcommand{\hermes}{HERMES}
 \newcommand{\hipparcos}{\emph{Hipparcos}}
@@
 structure. \Gaia\ and future spectroscopic surveys of the Galactic
-disk such as \apogee\ will be able to robustly test the origin of the
-Hercules stream and constrain the properties of the Galactic bar.
+disk such as \apogee\ and \hermes\ will be able to robustly test the
+origin of the Hercules stream and constrain the properties of the
+Galactic bar.
 \end{abstract}
@@
 is more than 3 kpc away. We see in this \figurename\ that the effect
-of distance uncertainties is negligible. The exact shape and
-parameters of the stellar distribution function also only minimally
-affect the predicted line-of-sight velocity distribution. Changing the
-scale-length within its current uncertainty changes the prediction
-only in the amplitude of the effect: The Hercules stream is more
-prominent for shorter scale-lengths since then there are relatively
-more stars in the inner Galaxy that are brought to the Solar circle by
-the interaction with the bar. Using a Shu distribution function
-\citep{shu69a} instead of the Dehnen distribution function also gives
-very similar predictions. Changing the axisymmetric potential by
-adjusting the shape of the rotation curve shifts the peak as expected
+of distance and realistic line-of-sight velocity uncertainties is
+negligible. The exact shape and parameters of the stellar distribution
+function also only minimally affect the predicted line-of-sight
+velocity distribution. Changing the scale-length within its current
+uncertainty changes the prediction only in the amplitude of the
+effect: The Hercules stream is more prominent for shorter
+scale-lengths since then there are relatively more stars in the inner
+Galaxy that are brought to the Solar circle by the interaction with
+the bar. Using a Shu distribution function \citep{shu69a} instead of
+the Dehnen distribution function also gives very similar
+predictions. Changing the axisymmetric potential by adjusting the
+shape of the rotation curve shifts the peak as expected
 \citep[see][]{dehnen00a}, but the Hercules feature nevertheless
-remains detectable.
+remains detectable. The properties of the bar itself, \eg, its pattern
+speed through \Rolr\ or its strength $\alpha$, change the predicted
+pattern, but they do not significantly alter the prospects for
+detection.
 
 \section{Discussion}
+
+The discussion in the previous sections shows that the bar-origin
+paradigm for the Hercules moving group makes strong predictions about
+how the Hercules feature should show up when we probe the velocity
+distribution in parts of the Galaxy far removed from the Solar
+neighborhood. In order to most strongly test the bar-origin for the
+Hercules moving group it is best to extrapolate this picture far from
+the Solar neighborhood and look at regions further than a few kpc
+removed. \Gaia 's parallaxes, proper motions, and mission-averaged
+line-of-sight velocities for dwarfs nearby and giants at distances
+greater than a few kpc will allow the full two-dimensional velocity
+distribution to be confronted with the predictions of
+\figurename~\ref{fig:rphi2d} (see below). However, the final
+\Gaia\ catalog will only be available around 2020.
+
+The predicted line-of-sight velocity distributions in the previous
+section show that the bar-origin theory for the Hercules moving group
+can be tested with line-of-sight velocities alone. At distances
+greater than a few kpc this will mean observing giants to trace the
+velocity distribution. Large, planned spectroscopic surveys with
+line-of-sight velocity uncertainties less than 1 km s$^{-1}$ can test
+the predictions made in this paper in the next few years (see
+below). These surveys will not need to observe many stars to test the
+predictions laid out here: Drawing samples from the predicted velocity
+distribution in \figurename~\ref{fig:1dvar} shows that about 500 stars
+suffice to distinguish the bar-origin model from the axisymmetric
+prediction (based on a KS test with $p < 0.01$). Since large distance
+uncertainties do not affect the prediction very much (see
+\figurename~\ref{fig:1dvar}), these stars can even be sampled
+throughout a volume of 500 pc or more.
+
+The bottom two panels of \figurename~\ref{fig:1dvar} show that the
+predicted line-of-sight velocity distribution depends strongly on the
+dynamical parameters of the bar. Changing the pattern speed of the bar
+through \Rolr\ shifts the predicted distribution and in particular the
+location of the Hercules feature. While this shift at a particular
+location is degenerate with the shift due to the uncertainty in the
+local circular velocity, results from various locations can be
+combined to break this degeneracy and the local circular velocity is
+also already quite well constrained \citep[\eg,][]{Bovy09b}. Changing
+the strength of the bar changes the relative heights of the two main
+peaks in the predicted velocity distribution. However, the main peak
+of the velocity distribution might not be as simple as the one
+predicted here due to other dynamical effects. Integrated measures,
+such as the number of stars in the main peak and the number of stars
+in the Hercules feature, could mitigate this somewhat. While
+predicting actual constraints would involve detailed simulations of
+the expected data, it is clear that constraints on the bar can be
+derived from the experiment proposed in this paper.
 
@@
 
+BOVY: 500 stars is enough for APOGEE to distinguish from axisymmetric
+background.
+
 \citep{Freeman10a}.
@@
   Blaauw,~A., 1970, in IAU Symp.~38, ed.~W.~Becker \& I.~Kontopoulos (Dordrecht: Reidel), 199
+
+\bibitem[Bovy, Hogg, \& Rix(2009)]{Bovy09b}
+  Bovy,~J., Hogg,~D.~W., \& Rix, H.-W., 2009,
+  \apj, 704, 1704
 
@@
   Sellwood,~J.~A., 2010,
-  \mnras, submitted, 
+  \mnras, in press,
   arXiv:1001.5197v1 [astro-ph.GA]
@@
 \bibitem[Shen \etal(2010)]{Shen10a}
-  Shen,~J., \etal, 2010, \apj L, submitted, arXiv:1005.0385v1 [astro-ph.CO]
+  Shen,~J., \etal, 2010, \apj, 720L, 72
 
@@
   parameters and shape of the distribution function; shape of the
-  rotation curve; and bar strength.}\label{fig:1dvar}
+  rotation curve; bar \Rolr; and bar strength.}\label{fig:1dvar}
 \end{figure}

### 2010-09-12 tex/response.txt
groups-simulations: finish referee edits

@@
 
-BOVY: V0
+Uncertainty in V0 only changes the scale of the predicted velocity
+distribution (since it is only used to go from vlos/V0 to the observed
+vlos). This is somewhat degenerate with the dynamical R_{olr}
+parameter, which is discussed at the end of the general discussion in
+Section 5.
 
@@
 
-> 3 kpc: giants (dwarfs in GAIA few kpc, giants go far enough and
-  bright enough to have RVs)
+This discussion is part of the new "general" discussion at the start
+of section 5.
 
@@
 
+The "testibility" of the model is discussed at greater length, but I
+think the whole paper makes it clear that this experiment is not a
+stretch for the spectroscopic surveys or Gaia. The line-of-sight
+velocity uncertainties of the spectroscopic surveys are much better
+than necessary to detect the feature, and not that many stars need to
+be observed to detect the feature (about 500, added to the
+discussion).
+
+The discussion at the end of Section 4 also makes it clear that none
+of the nuisance or dynamical parameters change how well the model can
+be tested.
+
+I added some discussion of how the prediction could be used to
+constrain the properties of the bar: The prediction depends on the
+exact properties of the bar, and while some of these dependencies are
+degenerate with dependecies on nuisance parameters, observing the
+velocity distribution at various locations will break some of these
+degeneracies. To actually calculate expected constraints would involve
+simulating the data, which are very unknown for most of the proposed
+surveys, but it is clear that constraints can be achieved using this
+experiment.
+
 	 2. Page 2, statement at the end of the first paragraph "..up
@@
 	 should be supported with a few relevant references.
+
+I added a few relevant references.
 

### 2010-09-12 tex/hercules.tex
groups-simulations: finish referee edits

@@
 the Galactic disk is not currently up to the task of separating out
-these effects.
+these effects \citep[\eg,][]{samland03a,roskar08a}.
 
@@
 distribution in parts of the Galaxy far removed from the Solar
-neighborhood. In order to most strongly test the bar-origin for the
+neighborhood. In order to most strongly test the bar origin for the
 Hercules moving group it is best to extrapolate this picture far from
@@
 suffice to distinguish the bar-origin model from the axisymmetric
-prediction (based on a KS test with $p < 0.01$). Since large distance
+prediction (based on a KS test with $P < 0.01$). Since large distance
 uncertainties do not affect the prediction very much (see
 \figurename~\ref{fig:1dvar}), these stars can even be sampled
-throughout a volume of 500 pc or more.
+throughout a volume of 500 pc or more. These spectroscopic surveys
+will also obtain detailed elemental abundances for their target stars,
+so they would also allow for a more detailed analysis of the
+abundances of stars in the Hercules stream \citep[following,
+  \eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
+mixing in the Galactic disk.
 
-The bottom two panels of \figurename~\ref{fig:1dvar} show that the
+The last two panels of \figurename~\ref{fig:1dvar} show that the
 predicted line-of-sight velocity distribution depends strongly on the
@@
 predicting actual constraints would involve detailed simulations of
-the expected data, it is clear that constraints on the bar can be
-derived from the experiment proposed in this paper.
+the expected data, it is clear that constraints on the dynamical
+properties of the bar can be derived from the experiment proposed in
+this paper.
 
@@
 longitudes that \apogee\ can target along the disk is $-5^{\circ} \leq
-l \leq 250^{\circ}$ (SDSS-III Project Description\footnote{Available
+l \leq 250^{\circ}$ and its expected line-of-sight uncertainties are
+about 0.5 km s$^{-1}$ (SDSS-III Project Description\footnote{Available
   at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
-\figurename~\ref{fig:detect} this limit is right on the edge of where
-we might hope to detect the Hercules feature. The Hercules bump in the
-direction $l = 250^{\circ}$ is about half as strong as in
+\figurename~\ref{fig:detect} this longitude limit is right on the edge
+of where we might hope to detect the Hercules feature. The Hercules
+bump in the direction $l = 250^{\circ}$ is about half as strong as in
 \figurename~\ref{fig:1dvar}, but remains as a clear signature of the
-bar. Since \apogee\ is expected to start taking data in 2011 and to
-finish operations before the first \Gaia\ data release, \apogee\ could
-potentially trace the Hercules stream around the Galaxy before
-\Gaia\ does. Since \apogee\ will also obtain detailed elemental
-abundances for its target stars, this would also allow for a more
-detailed analysis of the abundances of stars in the Hercules stream
-\citep[following, \eg,][]{Bensby07a,Bovy10a}, which could shed more
-light on radial mixing in the Galactic disk.
+bar. As before, about 500 stars a few kpc away are enough to detect
+the Hercules feature. Since \apogee\ is expected to start taking data
+in 2011 and to finish operations before the first \Gaia\ data release,
+\apogee\ could potentially trace the Hercules stream around the Galaxy
+before \Gaia\ does.
 
-BOVY: 500 stars is enough for APOGEE to distinguish from axisymmetric
-background.
-
-\citep{Freeman10a}.
+Another planned spectroscopic survey is
+\hermes\ \citep{Freeman10a}. While operating in the optical, it
+expects to survey hundreds of thousands of giants in the inner Galaxy,
+many as far as 6 kpc away, with line-of-sight velocity uncertainties
+less than 1 km s$^{-1}$ (K. Freeman, private communication). Since
+\hermes\ operates from the Southern hemisphere it can observe the most
+promising directions in \figurename~\ref{fig:detect} and, like
+\apogee, detect the Hercules feature before \Gaia.
 
@@
 
+\bibitem[Ro{\v s}kar \etal(2008)]{roskar08a}
+  Ro{\v s}kar,~R., \etal, 2008,
+  \apjl, 684, L79
+
+\bibitem[Samland \& Gerhard(2003)]{samland03a}
+  Samland,~M.~\& Gerhard,~O.~E., 2003,
+  \aap, 399, 961
+
 \bibitem[Sellwood(2010)]{sellwood10a}
@@
 \begin{figure}
-\includegraphics[width=0.35\textwidth]{distuncertain.ps}
-\includegraphics[width=0.35\textwidth]{vlosuncertain.ps}\\
-\includegraphics[width=0.35\textwidth]{df.ps}
-\includegraphics[width=0.35\textwidth]{slope.ps}\\
-\includegraphics[width=0.35\textwidth]{rolr.ps}
-\includegraphics[width=0.35\textwidth]{barstrength.ps}
+\includegraphics[width=0.32\textwidth]{distuncertain.ps}
+\includegraphics[width=0.32\textwidth]{vlosuncertain.ps}
+\includegraphics[width=0.32\textwidth]{df.ps}\\
+\includegraphics[width=0.32\textwidth]{slope.ps}
+\includegraphics[width=0.32\textwidth]{rolr.ps}
+\includegraphics[width=0.32\textwidth]{barstrength.ps}
 \caption{Variation of the predicted velocity distribution at R = 1.075

### 2010-09-12 tex/response.txt
groups-simulations: finish response

@@
 
-I have made changes following the referee's suggestions ...
-
-A detailed list of the main changes is:
+I have made changes following the referee's suggestions, in particular
+adding a general discussion to section 5 which discusses the kind of
+tracers needed and the number of stars needed to be observed to detect
+the Hercules feature. I also added two panels to figure 5 to address
+some of the referee's comments. All other changes were minor and in
+response to the referee.
 
@@
 uncertainty-free prediction and the uncertainty-convolved prediction
-are indistinguishable.
+are indistinguishable (which is clear from the new panel).
 
@@
 for Gaia, and an explicit comment that the velocities of giants will
-be measured accurately enough to verify the predictions of this paper.
+be measured accurately enough to verify the predictions of this
+paper. While the Hercules feature could certainly be detected in
+tangential velocities alone, Gaia will have plenty of line-of-sight
+velocities to make this prediction unnecessary.
 
@@
 
-All fixed!
+All fixed, thanks!

### 2010-11-10 tex/hercules.tex
hercules: proof changes

@@
 line of sight. This \figurename\ is no longer symmetric with respect
-to shifts of 180$^{\circ}$ degree in azimuth, since the line-of-sight
+to shifts of 180$^{\circ}$ in azimuth, since the line-of-sight
 direction breaks this symmetry. The gray curve is the line-of-sight
@@
 maximal Kullback-Leibler divergence are those in which the Hercules
-feature stands out most prominently in the line-of-sight.
+feature stands out most prominently in the line of sight.
 
@@
 function also only minimally affect the predicted line-of-sight
-velocity distribution. Changing the scale-length within its current
+velocity distribution. Changing the scale length within its current
 uncertainty changes the prediction only in the amplitude of the
-effect: The Hercules stream is more prominent for shorter
-scale-lengths since then there are relatively more stars in the inner
-Galaxy that are brought to the Solar circle by the interaction with
-the bar. Using a Shu distribution function \citep{shu69a} instead of
-the Dehnen distribution function also gives very similar
+effect: The Hercules stream is more prominent for shorter scale
+lengths since then there are relatively more stars in the inner Galaxy
+that are brought to the Solar circle by the interaction with the
+bar. Using a Shu distribution function \citep{shu69a} instead of the
+Dehnen distribution function also gives very similar
 predictions. Changing the axisymmetric potential by adjusting the
@@
 \citep{Perryman01a}. With parallax and proper motion accuracies down
-to 10 microarcseconds and mission-averaged line-of-sight velocity
+to 10 $\mu$as and mission-averaged line-of-sight velocity
 uncertainties smaller than 10 km s$^{-1}$ for many stars to 17th
@@
   Bailer-Jones, ~C.~A.~L., 2008,
-  in IAU Symp.~254, ed.~J.~Andersen, J.~Bland-Hawthorn, and B.~Nordstr\"{o}m, (Dordrecht: Kluwer), 475
+  in IAU Symp.~254, The Galaxy Disk in Cosmological Context, ed.~J.~Andersen, J.~Bland-Hawthorn, \& B.~Nordstr\"{o}m, (Dordrecht: Kluwer), 475
 
@@
 
-\bibitem[Bovy, Hogg, \& Rix(2009)]{Bovy09b}
-  Bovy,~J., Hogg,~D.~W., \& Rix, H.-W., 2009,
+\bibitem[Bovy \& Hogg(2010)]{Bovy10a} Bovy,~J.~\& Hogg,~D.~W., 2010,
+  \apj, 717, 617
+
+\bibitem[Bovy \etal(2009a)]{Bovy09b}
+  Bovy,~J., Hogg,~D.~W., \& Rix, H.-W., 2009a,
   \apj, 704, 1704
 
-\bibitem[Bovy, Hogg, \& Roweis(2009)]{Bovy09a} Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T., 2009,
+\bibitem[Bovy \etal(2009b)]{Bovy09a} Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T., 2009b,
   \apj, 700, 1794
-
-\bibitem[Bovy \& Hogg(2010)]{Bovy10a} Bovy,~J.~\& Hogg,~D.~W., 2010,
-  \apj, 717, 617
 
@@
 \bibitem[{{ESA}(1997)}]{ESA97a}
-{ESA}, 1997, {The \emph{Hipparcos} and Tycho Catalogues} (ESA SP-1200; Noordwijk: ESA)
+{ESA}, 1997, {The \emph{Hipparcos} and \emph{Tycho} Catalogues} (ESA SP-1200; Noordwijk: ESA)
 
@@
   Sellwood,~J.~A., 2010,
-  \mnras, in press,
-  arXiv:1001.5197v1 [astro-ph.GA]
+  \mnras, 409, 145
 
@@
 \bibitem[Shen \etal(2010)]{Shen10a}
-  Shen,~J., \etal, 2010, \apj, 720L, 72
+  Shen,~J., \etal, 2010, \apj, 720, L72
 
@@
 \includegraphics[width=0.5\textwidth]{veldist.eps}
-\caption{The velocity distribution of nearby ($\lesssim 100$ pc) stars
+\caption{Velocity distribution of nearby ($\lesssim 100$ pc) stars
   from \citet{Bovy09a}. \vR\ and \vphi\ are the velocities toward the
@@
 \includegraphics[width=0.32\textwidth]{barstrength.ps}
-\caption{Variation of the predicted velocity distribution at R = 1.075
-  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
-  distance uncertainties; line-of-sight velocity uncertainties;
+\caption{Variation of the predicted velocity distribution at $R$ =
+  1.075 \Ro , $\phi = -21^{\circ}$ ($d$= 0.4 \Ro, $l$ = 270$^{\circ}$)
+  with distance uncertainties; line-of-sight velocity uncertainties;
   parameters and shape of the distribution function; shape of the
